### 2024-01-25 products/manuscript/Manuscript.qmd
Setting up final project
I explored the different folders and added our names to the main readme file and as authors in the manuscript.qmd file.

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Taylor Glass and Aryln Santiago
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2024-01-31 products/manuscript/Manuscript.qmd
Added an .rds file to data folder
Added the Covid.rds file to the data folder to begin forming a research question.
Added a new dataset to data folder
Added the AIU All Women Dataset from the Guttmacher Institute which is fully publicly available
Project Part 1
I uploaded the codebook for the AIU dataset and added information on the data and data source. I began exploring the data and formulated a potential research question.

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
+title: "Global Distribution of Unintended Pregnancies and Contraceptive Usage"
 subtitle: ""
@@
 ```{r, echo=FALSE, message=FALSE}
+xcode-select --install
 # load a few R packages
@@
 ```
+## Project Part 1  
+# Data Source 
+The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world.
+The data used here is from their Adding It Up project.
+The goals of this project include estimating the need for, impact of, and costs associated with providing sexual and reproductive health services.
+Low and middle income countries are the target audience for this research, specifically women of reproductive age (15-49).
+Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted.
+
+# Load and describe the data 
+```{r}
+library(here)
+AIUdata <- read_csv(here("data", "raw-data", "AIU All Women Dataset.csv"))
+dim(AIUdata)
+colnames(AIUdata)
+```
+There are 132 observations, one row for each low- to middle- income country, of the 399 variable included in the Adding It Up All Women Dataset.
+Some examples of measurements included are percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+
+# Potential Research Questions
+1) What is the relationship between rates of unintended pregnancies and usage of contraceptive methods?
+  -outcome variables: 
+  unintended pregnancies, total number (pct_upreg) 
+  unintended pregnancies, total number [subregional]      (upreg_subreg)
+
+  -predictor variables:
+  number of women 25-34 using modern methods (modmmethods_2534)
+  number of women 25-34 using sterilization (ster_2534)
+  number of women using long-acting reversible methods (larc_2534)
+  number of women 25-34 using short-acting methods (sarc_2534)
+  number of women 25-34 using traditional methods (tradmethods_2534)
+  number of women 25-34 using no method (nomethod_2534)
+  *These variables exist for two other age ranges (20-24 and 35-49), so we could analyze contraceptive usage across age categories too
+
+2) How do the rates of safe abortions fluctuate with available methods for safe abortions?
+
+# How to analyze 
+All of the variables used in this analysis are numeric, so we can build linear regression models to analyze how each method of contraception affects the percentage of unintended pregnancies.
+We can group countries together based on the subregion variable to compare how these relationships vary across the world.
+There are other interesting variables that can be used for grouping for analysis such as wealth, urban/rural location, and marriage status.
 

### 2024-02-01 products/manuscript/Manuscript.qmd
Project Part 1 Updates
I added a summary of the structure of the dataset along with a crude visualization of unintended pregnancy rates by country. I also added information about the surveys and organizations that provide the information included in this dataset.

@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
 This uses MS Word as output format.
@@
 # load a few R packages
+install.packages("dplyr")
+install.packages("naniar")
 library(here)
+library(readr)
 library(knitr)
+library(dplyr)
+library(naniar)
+library(ggplot2)
 ```
@@
 Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted.
+Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others.
+The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.
 
@@
 dim(AIUdata)
-colnames(AIUdata)
+str(AIUdata)
 ```
 There are 132 observations, one row for each low- to middle- income country, of the 399 variable included in the Adding It Up All Women Dataset.
-Some examples of measurements included are percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+Some examples of numeric variables included are percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+There are three important character variables that can be used to group the numeric variables: country name, region, and sub region.
+
+# Summarize the data 
+```{r}
+table(AIUdata$region)
+table(AIUdata$subregion)
+summary(AIUdata$pct_upreg)
+ggplot(AIUdata, aes(x = country, y = pct_upreg, fill = region)) +
+  geom_col(position = "dodge") +
+  labs(title = "Percentage Unintended Pregnancies by Country",
+       x = "Country",
+       y = "Percentage")
+```
+The dataset includes 53 African countries, 36 Asian countries, 11 European Countries, 24 Latin American/Caribbean countries, and 8 Oceanic countries.There are 16 subregions that provide better detail about each country's location.
+The average percentage of unintended pregnancies at the country level is 46.79%.
+with a maximum of 75.65%.
+This variable has 24 missing values, which will need to be cleaned.
+A simple visualization shows that African and Latin American/Carribean countries have the highest percentages of unintended pregnancies.
 
@@
   -predictor variables:
-  number of women 25-34 using modern methods (modmmethods_2534)
-  number of women 25-34 using sterilization (ster_2534)
-  number of women using long-acting reversible methods (larc_2534)
-  number of women 25-34 using short-acting methods (sarc_2534)
-  number of women 25-34 using traditional methods (tradmethods_2534)
-  number of women 25-34 using no method (nomethod_2534)
-  *These variables exist for two other age ranges (20-24 and 35-49), so we could analyze contraceptive usage across age categories too
+  number of women 35-49 using modern methods (modmmethods_3549)
+  number of women 35-49 using sterilization (ster_3549)
+  number of women 35-49 using long-acting reversible methods (larc_3549)
+  number of women 35-49 using short-acting methods (sarc_3549)
+  number of women 35-49 using traditional methods (tradmethods_3549)
+  number of women 35-49 using no method (nomethod_3549)
 
-2) How do the rates of safe abortions fluctuate with available methods for safe abortions?
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies to investigate the effectiveness of each method.
+We hope to find relationships between subregions, rates of unintended pregnancies, and contraceptive methods to determine possible interventions that will increase the effectiveness of contraceptive methods in each subregion.
+Each of the predictor variables exist for two other age ranges (20-24 and 25-34), so we could analyze contraceptive usage across age categories within each subregion as well.
+
+Backup question: How do the rates of safe abortions fluctuate with available methods for safe abortions?
 
@@
 All of the variables used in this analysis are numeric, so we can build linear regression models to analyze how each method of contraception affects the percentage of unintended pregnancies.
-We can group countries together based on the subregion variable to compare how these relationships vary across the world.
+We can group countries together based on the subregion variable to compare how these relationships vary across the world as shown in the crude visualization above.
 There are other interesting variables that can be used for grouping for analysis such as wealth, urban/rural location, and marriage status.

### 2024-02-09 R/processing-code/processingfile-v1.qmd
Part 2 Updates
Started working in the processing-code file with loading the data and codebook. Also deleted an extra version of the codebook that was unnecessary.

---
title: "Cleaning Script"
author: "Taylor Glass and Arlyn Santiago"
date: "2024-02-09"
output: html_document
---

# Processing script

# Setup

Load the necessary packages for data loading and processing.

```{r}
library(readr) # for loading csv files
library(readxl) # for loading excel filess (codebook)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

```{r}
# path to data using here() function
data_location <- here("data","raw-data","AIU All Women Dataset.csv")
rawdata <- read_csv(data_location)
```

# Check data

We will load and examine the codebook first.

```{r}
# write file path to the codebook
codebook_location <- here("data", "AIU All Women Dataset Codebook.xlsx")

# specify the 3 sheets of the codebook we need 
sheets <- c("Demographics", "Contraceptive need & use", "Pregnancy-related & newborn")

# load codebook
codebook <- lapply(sheets, function(sheet) {
  readxl::read_excel(codebook_location, sheet = sheet)
})
print(codebook)
```

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

First, there is an entry for height which says "sixty" instead of a number. 
Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.

```{r}
d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
                  dplyr::mutate(Height = as.numeric(Height))
skimr::skim(d1)
hist(d1$Height)
```

Now we see that there is one person with a height of 6. That could be a typo, or someone mistakenly entered their height in feet. Since we unfortunately don't know, we might need to remove this person, which we'll do here.

```{r}
d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
skimr::skim(d2)
```

Height values seem ok now.

Now let's look at the `Weight` variable. There is a person with weight of 7000, which is impossible, and one person with missing weight.
To be able to analyze the data, we'll remove those individuals as well.

```{r}
d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
skimr::skim(d3)
```

Now checking the `Gender` variable. Gender should be a categorical/factor variable but is loaded as character. We can fix that with simple base R code to mix things up.

```{r}
d3$Gender <- as.factor(d3$Gender)  
skimr::skim(d3)
```

Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
Well proceed here by removing that individual with that NA entry. Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.

```{r}
d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
skimr::skim(d4)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- d4
```

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-02-11 code/processing-code/processingfile-v1.qmd
Part 2 Updates
Worked on processing the data by converting variable types and exploring missingness in the variables we selected to answer our question.  Also edited read.me files to be updated to our project.

---
title: "Processing Script"
author: "Taylor Glass and Arlyn Santiago"
date: "2024-02-09"
output: html_document
---

# Processing script

# Setup

Load the necessary packages for data loading and processing.

```{r}
library(readr) # for loading csv files
library(readxl) # for loading excel files (codebook)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(naniar) #to explore data missingness
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

```{r}
# path to data using here() function
data_location <- here("data","raw-data","AIU All Women Dataset.csv")
rawdata <- read_csv(data_location)
```

# Check data

We will load and examine the codebook first.

```{r}
# write file path to the codebook
codebook_location <- here("data", "AIU All Women Dataset Codebook.xlsx")

# specify the 3 sheets of the codebook we need 
sheets <- c("Demographics", "Contraceptive need & use", "Pregnancy-related & newborn")

# load codebook
codebook <- lapply(sheets, function(sheet) {
  readxl::read_excel(codebook_location, sheet = sheet)
})
print(codebook)
```

# Glimpse at the data

We will take a glimpse at our data to better understand the variables we are working with. There are four character variables: `country`, `data_source_year`, `region`, and `subregion`. The remainder of the variables are numeric and describe percentage of birth outcomes, maternal deaths due to birth outcomes, contraceptive methods divided by age groups, wealth levels in relation to contraceptive methods, various abortion procedure rates, current cost of pregnancy outcomes and STI treatment.

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```
# Identifying variables to be cleaned

```{r}
rawdata2 <- rawdata %>% 
                select(country, region, subregion, pct_upreg, upreg_subreg, modmethods_3549, ster_3549, larc_3549, sarc_3549, tradmethods_3549, nomethod_3549)

skimr::skim(rawdata2)
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

First, the `region` variable is a character variable, but it will be more useful later as a factor. We can easily mutate it from a character variable to factor variable.  
```{r}
# convert character variable to factor variable
rawdata2 <- rawdata2 %>% 
              mutate(region = as.factor(region))

# check that the conversion worked 
class(rawdata2$region)
```

Second, there are 24 entries missing for the percentage of all unintended pregnancies by country, which is only 18.18% of the observations. After isolating the observations with this missing variable, we see that the majority of countries in this category are located in Asia. This does not make sense because the majority of countries in the raw dataset are located in Africa, which suggests that the data are not missing at random. 
```{r}
# find missing percentage for each variable
miss_var_summary(rawdata2)

# isolate the percentage of all unintended pregnancies by country to discover trends in missingness 
variable_with_missing <- "pct_upreg"
missing_obs <- rawdata2[is.na(rawdata2[[variable_with_missing]]), ]
other_variables <- colnames(rawdata2)[!colnames(rawdata2) == variable_with_missing]
pct_upregmissing <- missing_obs[, other_variables]
print(pct_upregmissing)

# compare dataset with the missing variable to the full dataset
table(pct_upregmissing$region)
table(rawdata2$region)
```

We can impute the missing data here because only 18.18% of the data is missing, so the mean value of the variable is fairly reliable (find a source for this?) After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.  
```{r}
summary(rawdata2$pct_upreg)
```

Third, we need to evaluate the 108 missing entries of the percentage of unintended pregnancies subregion. Since 81.81% of this variable is missing, it does not make sense to include it in the analyis because we do not have enough information to impute the missing data. 
```{r}
# isolate the percentage of all unintended pregnancies by subregion to discover trends in missingness
variable_with_missing <- "upreg_subreg"
missing_obs2 <- rawdata2[is.na(rawdata2[[variable_with_missing]]), ]
other_variables2 <- colnames(rawdata2)[!colnames(rawdata2) == variable_with_missing]
pct_subregmissing <- missing_obs2[, other_variables2]
print(pct_subregmissing)
```

Lastly, we will assign the clean data to a final variable, which makes it easier to add further cleaning steps above if necessary.

```{r}
processeddata <- d4
```

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-02-16 code/processing-code/processingfile-v1.qmd
Part 2 Updates
I processed several variables that we selected to answer our research question, updated readme files, and deleted extra files that are not related to our project. I found variables that seem to answer our first question already, so I formulated a second question in the manuscript.qmd file.

@@
 author: "Taylor Glass and Arlyn Santiago"
-date: "2024-02-09"
+date: "2024-02-16"
 output: html_document
@@
 library(here) #to set paths
+library(tidymodels) #to impute missing data
+library(recipes) #to impute missing data
 ```
@@
 ```
+Note: should we include other covariates like marriage percentages, number of live births, number of induced abortions
 
@@
 
-First, the `region` variable is a character variable, but it will be more useful later as a factor. We can easily mutate it from a character variable to factor variable.  
+First, the `region` variable is a character variable, but it will be more useful later as a factor.
+We can easily mutate it from a character variable to factor variable.
+All the remaining variables are numeric as they should be.
 ```{r}
@@
 
-Second, there are 24 entries missing for the percentage of all unintended pregnancies by country, which is only 18.18% of the observations.
+Second, there are 24 entries missing for the percentage of all unintended pregnancies by country `pct_upreg`, which accounts for 18.18% of the observations.
 After isolating the observations with this missing variable, we see that the majority of countries in this category are located in Asia.
@@
 
-We can impute the missing data here because only 18.18% of the data is missing, so the mean value of the variable is fairly reliable (find a source for this?) After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.
+We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is the best outcome variable to answer the research question.
+After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.
 ```{r}
+# find mean to use for imputed data
 summary(rawdata2$pct_upreg)
+
+# create a recipe to impute data 
+rec <- recipe(pct_upreg ~ ., data = rawdata2) %>%
+  step_naomit(all_predictors(), -all_outcomes(), skip = TRUE) %>%
+  step_impute_mean(all_predictors()) #which predictors would I use?
+
+# prepare the raw data using the recipe
+data_prep <- prep(rec, training = rawdata2)
+
+# impute the data 
+data_imputed <- bake(data_prep, new_data = NULL) #all pct_upreg are NaN values
+
+# identify rows with missing values in the original dataset
+missing_rows <- which(is.na(rawdata2$pct_upreg))
+
+# replace missing values in the original dataset with imputed values
+rawdata2$pct_upreg[missing_rows] <- data_imputed$pct_upreg[missing_rows]
+
+```
+Once the missing values have been imputed, we should further clean the `pct_upreg` variable to have a typical percentage format. 
+```{r}
+rawdata2 <- rawdata2 %>% 
+              mutate(pct_upreg = pct_upreg * 100)
 ```
 
-Third, we need to evaluate the 108 missing entries of the percentage of unintended pregnancies subregion.
-Since 81.81% of this variable is missing, it does not make sense to include it in the analyis because we do not have enough information to impute the missing data.
+We also need to evaluate the 108 missing entries of the percentage of unintended pregnancies by subregion `upreg_subreg`.
+Since 81.81% of this variable is missing, it does not make sense to include it in the analysis because we do not have enough information to impute the missing data.
+We will remove this variable for the sake of this analysis.
 ```{r}
@@
 print(pct_subregmissing)
+
+# remove unintended pregnancies by subregion variable due to missingness
+rawdata2 <- rawdata2 %>% 
+              select(-upreg_subreg)
+```
+
+Each of the variables about birth control methods, `mod_methods3549`, `ster_3549`, `larc_3549`, `sarc_3549`, `tradmethods_3549`, `nomethod_3549`, are not very intuitive, so they should be renamed.
+Since we are only studying one age group, 35 to 49 years, it is unnecessary for those numbers to be included in each variable name.
+It also makes sense to rename `pct_upreg` to be more descriptive since we already know it is percentage format.
+```{r}
+# rename predictor variables
+rawdata2 <- rawdata2 %>% rename(modern = modmethods_3549,
+                                sterile = ster_3549, 
+                                longactrev = larc_3549,
+                                shortact = sarc_3549,
+                                tradition = tradmethods_3549,
+                                nomethod = nomethod_3549,
+                                unint_preg = pct_upreg)
+# check if renaming worked
+head(rawdata2)
+```
+To answer our research question about how unintended pregnancies are related to birth control methods, it makes sense to have a variable with the total number of women using birth control in each country.
+This will service as the main predictor variable, and each type of birth control can serve as a covariate.
+```{r}
+# create new variable with total number on birth control of any form
+rawdata2 <- rawdata2 %>% 
+              mutate(totalbc = modern + sterile + longactrev + shortact + tradition + nomethod)
+
+# check new variable is created 
+head(rawdata2)
 ```
@@
 ```{r}
-processeddata <- d4
+processeddata <- rawdata2
 ```
@@
 
-Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
-This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data:
-http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
+Finally, we will save the clean data as RDS file to preserve coding like factors, characters, numeric, etc.
 
 ```{r}
-save_data_location <- here::here("data","processed_data","processeddata.rds")
+save_data_location <- here("data","processed-data","processeddata.rds")
 saveRDS(processeddata, file = save_data_location)
 ```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).

### 2024-02-16 products/manuscript/Manuscript.qmd
Testing to see if I can push changes.
Added an extra backup question.
References
Part 2 Updates
I processed several variables that we selected to answer our research question, updated readme files, and deleted extra files that are not related to our project. I found variables that seem to answer our first question already, so I formulated a second question in the manuscript.qmd file.

@@
 
-Backup question: How do the rates of safe abortions fluctuate with available methods for safe abortions?
+2) How does marriage status affect percentage of unintended pregnancies across different types of birth control?
+  -outcome variables:
+      unintended pregnancies, total number (pct_upreg)
+      unintended pregnancies, using larc and sterilization (pct_upreg_larcster)
+      unintended pregnancies, using short term reversible (pct_upreg_sarc)
+  -predictor variables:
+      Number wanting to avoid pregnancy, all married women (inneed_married)
+      Number using sterilization, all married women	(ster_married) **combine these two in processing
+      Number using long-acting reversible methods, all married women	(larc_married)
+      Number using short-acting methods, all married women	(sarc_married)
+      Number wanting to avoid pregnancy, formerly married women 	(inneed_formerlymar)
+      Number using sterilization, formerly married women	(ster_formerlymar)
+      Number using long-acting reversible methods, formerly married women	(larc_formerlymar)
+      Number using short-acting methods, formerly married women	(sarc_formerlymar)
+      Number wanting to avoid pregnancy, never married women 	(inneed_nevermar)
+      Number using all modern methods, never married women	(modmethods_nevermar)
+      Number using sterilization, never married women	(ster_nevermar)
+      Number using long-acting reversible methods, never married women	(larc_nevermar)
+      Number using short-acting methods, never married women	(sarc_nevermar)
 
@@
 # References
+
+https://pubmed.ncbi.nlm.nih.gov/32981858/
+https://pubmed.ncbi.nlm.nih.gov/30687521/ #This study is set in Senegal 
+https://www.sciencedirect.com/science/article/pii/S001078241730478X

### 2024-02-18 code/eda-code/eda.qmd
Part 2 Updates
I found new variables to use as covariates for our updated research question. I added those variables to the raw data that is being processed, explored them, and created a new variable. I also made notes on how to start exploring the data.

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

explore how many people use each method
explore how many countries are in each category for each method using new factor variable
explore percentage of unintended pregnancies across methods

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

Height first.

```{r}
p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
plot(p1)
figure_file = here("results","height_distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Now weights.

```{r}
p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
plot(p2)
figure_file = here("results","weight_distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now height as function of weight.

```{r}
p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","height_weight.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

```{r}
p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","height_weight_stratified.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

### 2024-02-18 code/processing-code/processingfile-v1.qmd
Part 2 Updates
I found new variables to use as covariates for our updated research question. I added those variables to the raw data that is being processed, explored them, and created a new variable. I also made notes on how to start exploring the data.

@@
 ```{r}
+#| message: false
 library(readr) # for loading csv files
@@
 ```{r}
+#| message: false
 # path to data using here() function
@@
 ```{r}
+#| message: false
 # write file path to the codebook
@@
 ```{r}
+#| message: false
 dplyr::glimpse(rawdata)
@@
 # Identifying variables to be cleaned 
+We want to explore the relationship between unintended pregnancies and birth control methods across different age groups.
+We will use three demographic variable: `country`, `region`, and `subregion`.
+The variable for the total percentage of unintended pregnancies, `pct_upreg` will be the main outcome variable.
+We could also use `upreg_subreg` to explore unintended pregnancies by subregion.
+The predictor variables will include percentage of unintended pregnancies across three types of birth control: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`).
+No method variables can serve as control groups for statistical tests.
+Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49.
+-Number of women 15-19 using long-acting reversible methods `larc_1519`
+-Number of women 15-19 using short-acting methods `sarc_1519`
+-Number of women 15-19 using traditional methods `tradmethods_1519`
+-Number of women 15-19 using no method `nomethod_1519`
+-Number of women 20-24 using long-acting reversible methods `larc_2024`
+-Number of women 20-24 using short-acting methods	`sarc_2024`
+-Number of women 20-24 using traditional methods	`tradmethods_2024`
+-Number of women 20-24 using no method	`nomethod_2024`
+-Number of women 25-34 using long-acting reversible methods	`larc_2534`
+-Number of women 25-34 using short-acting methods	`sarc_2534`
+-Number of women 25-34 using traditional methods	`tradmethods_2534`
+-Number of women 25-34 using no method	`nomethod_2534`
+-Number of women 35-49 using long-acting reversible method	`larc_3549`
+-Number of women 35-49 using short-acting method	`sarc_3549`
+-Number of women 35-49 using all traditional methods	`tradmethods_3549`
+-Number of women 35-49 using no method	`nomethod_3549`
+
+code for low, medium, and high use per country using case_when()
 
 ```{r}
+#| message: false
+# select the outcome variable, predictor variables, and covariates  
 rawdata2 <- rawdata %>% 
-                select(country, region, subregion, pct_upreg, upreg_subreg, modmethods_3549, ster_3549, larc_3549, sarc_3549, tradmethods_3549, nomethod_3549)
+                select(country, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534,  sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, sarc_3549, tradmethods_3549, nomethod_3549)
 
-skimr::skim(rawdata2)
+# take a glimpse at the selected variables 
+skim(rawdata2)
+head(rawdata2)
 ```
@@
 We can easily mutate it from a character variable to factor variable.
-All the remaining variables are numeric as they should be.
+All the remaining predictor and covariate variables are numeric as they should be.
 ```{r}
+#| message: false
 # convert character variable to factor variable
@@
 ```{r}
+#| message: false
 # find missing percentage for each variable
@@
 
-We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is the best outcome variable to answer the research question.
+We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is essential for answering the research question.
 After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.
 ```{r}
+#| message: false
 # find mean to use for imputed data
@@
 ```{r}
+#| message: false
+# convert percentage variable to typical percentage format
 rawdata2 <- rawdata2 %>% 
               mutate(pct_upreg = pct_upreg * 100)
+
+# ensure mutation worked correctly
+summary(rawdata2$pct_upreg)
 ```
@@
 ```{r}
+#| message: false
 # isolate the percentage of all unintended pregnancies by subregion to discover trends in missingness
@@
 
-Each of the variables about birth control methods, `mod_methods3549`, `ster_3549`, `larc_3549`, `sarc_3549`, `tradmethods_3549`, `nomethod_3549`, are not very intuitive, so they should be renamed.
-Since we are only studying one age group, 35 to 49 years, it is unnecessary for those numbers to be included in each variable name.
-It also makes sense to rename `pct_upreg` to be more descriptive since we already know it is percentage format.
+Fortunately, the rest of the variables required for this analysis are complete.
+Each of the four variables describing the percentage of unintended pregnancies based on birth control type also need to be converted to traditional percentage format.
 ```{r}
-# rename predictor variables
-rawdata2 <- rawdata2 %>% rename(modern = modmethods_3549,
-                                sterile = ster_3549, 
-                                longactrev = larc_3549,
-                                shortact = sarc_3549,
-                                tradition = tradmethods_3549,
-                                nomethod = nomethod_3549,
-                                unint_preg = pct_upreg)
-# check if renaming worked
-head(rawdata2)
+#| message: false
+# convert percentage variable to typical percentage format
+rawdata2 <- rawdata2 %>% 
+              mutate(pct_upreg_larcster = pct_upreg_larcster * 100, 
+                     pct_upreg_sarc = pct_upreg_sarc * 100, 
+                     pct_upreg_trad = pct_upreg_trad * 100, 
+                     pct_upreg_nouse = pct_upreg_nouse * 100)
+
+# ensure mutations worked correctly
+summary(rawdata2$pct_upreg_larcster)
+summary(rawdata2$pct_upreg_sarc)
+summary(rawdata2$pct_upreg_trad)
+summary(rawdata2$pct_upreg_nouse)
 ```
-To answer our research question about how unintended pregnancies are related to birth control methods, it makes sense to have a variable with the total number of women using birth control in each country.
-This will service as the main predictor variable, and each type of birth control can serve as a covariate.
+
+To make the covariate variables more useful, it makes sense to create a new factor variable denoting low, medium, and high usage of each birth control method by country.
+First, we need to determine the boundaries for each level of usage by exploring how many women use each method across age groups.
+We need to create a variable showing the total usage of each birth control method to remove the effect of age.
 ```{r}
-# create new variable with total number on birth control of any form
+#| message: false
+# create new variables showing total usage of each method including all age groups
 rawdata2 <- rawdata2 %>% 
-              mutate(totalbc = modern + sterile + longactrev + shortact + tradition + nomethod)
+              mutate(total_larc = larc_1519 + larc_2024 + larc_2534 + larc_3549,
+                     total_sarc = sarc_1519 + sarc_2024 + sarc_2534 + sarc_3549,
+                     total_trad = tradmethods_1519 + tradmethods_2024 + tradmethods_2534 + tradmethods_3549,
+                     total_none = nomethod_1519 + nomethod_2024 + nomethod_2534 + nomethod_3549)
 
-# check new variable is created 
-head(rawdata2)
+# explore summary of each total to set boundaries for each level of usage
+rawdata2 %>% skim(total_larc, total_sarc, total_trad, total_none)
+```
+After finding the distribution of number of women using each type of birth control, we can use that information to create a new factor variable denoting low, medium, and high usage for each method.
+The "low" category will include observations below the 25th percentile.
+The "medium" category will include observations between the 25th and 75th percentiles.
+The "high" category will include observations above the 75th percentile.
+```{r}
+#| message: false
+# create usage variable for long acting reversible methods 
+rawdata2 <- rawdata2 %>% 
+              mutate(larc_usage = as.factor(case_when(
+                                  total_larc < 11.5 ~ "Low",
+                                  total_larc >= 11.5 & total_larc < 407.4633 ~ "Medium",
+                                  total_larc >= 407.4633 ~ "High")))
+
+# check if new variable was created correctly 
+class(rawdata2$larc_usage)
+
+# create usage variable for short acting reversible methods 
+rawdata2 <- rawdata2 %>% 
+              mutate(sarc_usage = as.factor(case_when(
+                                  total_sarc < 114.33113 ~ "Low",
+                                  total_sarc >= 114.33113 & total_larc < 1554.7609 ~ "Medium",
+                                  total_sarc >= 1554.7609	 ~ "High")))
+
+# check if new variable was created correctly 
+class(rawdata2$sarc_usage)
+
+# create usage variable for traditional methods 
+rawdata2 <- rawdata2 %>% 
+              mutate(trad_usage = as.factor(case_when(
+                                  total_trad < 15.05000	 ~ "Low",
+                                  total_trad >= 15.05000 & total_larc < 332.4750 ~ "Medium",
+                                  total_trad >= 332.4750	 ~ "High")))
+
+# check if new variable was created correctly 
+class(rawdata2$trad_usage)
+
+# create usage variable for no methods
+rawdata2 <- rawdata2 %>% 
+              mutate(none_usage = as.factor(case_when(
+                                  total_none < 80.07500	 ~ "Low",
+                                  total_none >= 80.07500 & total_larc < 875.2750 ~ "Medium",
+                                  total_none >= 875.2750	 ~ "High")))
+
+# check if new variable was created correctly 
+class(rawdata2$none_usage)
 ```
@@
 ```{r}
+#| message: false
 processeddata <- rawdata2
@@
 ```{r}
+#| message: false
 save_data_location <- here("data","processed-data","processeddata.rds")

### 2024-02-18 products/manuscript/Manuscript.qmd
Part 2 Updates
I found new variables to use as covariates for our updated research question. I added those variables to the raw data that is being processed, explored them, and created a new variable. I also made notes on how to start exploring the data.

@@
 
-# Potential Research Questions
-1) What is the relationship between rates of unintended pregnancies and usage of contraceptive methods?
-  -outcome variables: 
-  unintended pregnancies, total number (pct_upreg) 
-  unintended pregnancies, total number [subregional]      (upreg_subreg)
+# Research Questions
+1) What is the relationship between rates of unintended pregnancies and usage of contraceptive methods across various age groups?
+The main outcome variables will be total percentage of unintended pregnancies (`pct_upreg`). 
+The predictor variables will include percentage of unintended pregnancies across three types of birth control: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`).
+No method variables can serve as control groups for statistical tests.
+Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49. 
+-Number of women 15-19 using long-acting reversible methods `larc_1519`
+-Number of women 15-19 using short-acting methods `sarc_1519`
+-Number of women 15-19 using traditional methods `tradmethods_1519`
+-Number of women 15-19 using no method `nomethod_1519`
+-Number of women 20-24 using long-acting reversible methods `larc_2024`
+-Number of women 20-24 using short-acting methods	`sarc_2024`
+-Number of women 20-24 using traditional methods	`tradmethods_2024`
+-Number of women 20-24 using no method	`nomethod_2024`
+-Number of women 25-34 using long-acting reversible methods	`larc_2534`
+-Number of women 25-34 using short-acting methods	`sarc_2534`
+-Number of women 25-34 using traditional methods	`tradmethods_2534`
+-Number of women 25-34 using no method	`nomethod_2534`
+-Number of women 35-49 using long-acting reversible method	`larc_3549`
+-Number of women 35-49 using short-acting method	`sarc_3549`
+-Number of women 35-49 using all traditional methods	`tradmethods_3549`
+-Number of women 35-49 using no method	`nomethod_3549`
 
-  -predictor variables:
-  number of women 35-49 using modern methods (modmmethods_3549)
-  number of women 35-49 using sterilization (ster_3549)
-  number of women 35-49 using long-acting reversible methods (larc_3549)
-  number of women 35-49 using short-acting methods (sarc_3549)
-  number of women 35-49 using traditional methods (tradmethods_3549)
-  number of women 35-49 using no method (nomethod_3549)
-
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies to investigate the effectiveness of each method.
-We hope to find relationships between subregions, rates of unintended pregnancies, and contraceptive methods to determine possible interventions that will increase the effectiveness of contraceptive methods in each subregion.
-Each of the predictor variables exist for two other age ranges (20-24 and 25-34), so we could analyze contraceptive usage across age categories within each subregion as well.
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies across age groups to investigate the effectiveness of each method.
+We hope to find relationships between subregions, rates of unintended pregnancies, and contraceptive methods to determine possible interventions that will increase the effectiveness of contraceptive methods.
 

### 2024-02-20 products/manuscript/Manuscript.qmd
Introduction

@@
 
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies across age groups (20-24) and (35-49) to investigate the effectiveness of each contraceptive method.Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies.
+According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group.
+Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study.
+The table provided above under `Summarize the data` shows that African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
+This could be due to the current legal status of contraception within Latin American countries.
+Free emergency contraception access exists, but regulations differ.
+Chile, Colombia, and Ecuador legally recognize access, while Nicaragua and Bolivia rely on Ministerial protocols.
+Argentina and Brazil lack legal recognition, but offer protocols and guides.
+Mexico requires provision to victims of sexual and domestic violence (Hevia, 2012).
+
 ## Description of data and data source
@@
 
+This data originates from the Guttmacher Institute's Adding It Up project.
+The data contains numeric variables: percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+There are only three  character variables that can be used to group the numeric variables: country name, region, and sub region.
+So far we have dvided the observations into "low", "medium", and "high" category of birth control usage.
+
 ## Questions/Hypotheses to be addressed
 _State the research questions you plan to answer with this analysis._
+
+1. How does marriage status affect percentage of unintended pregnancies across different types of birth control?
+
+2. Is there a significant difference in the percentage of unplanned pregnancies between reversible contraceptives (LARCs?) among the age groups (20-24) and (35-49)?
 
@@
 https://www.sciencedirect.com/science/article/pii/S001078241730478X 
+https://www.sciencedirect.com/science/article/abs/pii/S0020729211005406 #legal status article

### 2024-02-20 code/eda-code/eda.qmd
Part 2 Updates
We fixed an error in one of the variables we created, total_larcster. We began our EDA with histograms and scatter plots with numeric variables.

@@
 ```{r}
-#Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+#Path to data
+data_location <- here("data","processed-data","processeddata.rds")
 #load data
@@
 # Data exploration through tables
-
-Showing a bit of code to produce and save a summary table.
+We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information
 
 ```{r}
-summary_df = skimr::skim(mydata)
+summary_df = skim(mydata)
 print(summary_df)
 # save to file
-summarytable_file = here("results", "summarytable.rds")
+summarytable_file = here("results", "tables", "summarytable.rds")
 saveRDS(summary_df, file = summarytable_file)
@@
 
-We are saving the results to the `results` folder.
-Depending on how many tables/figures you have, it might make sense to have separate folders for each.
-And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
-Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
-
 # Data exploration through figures
 
-Histogram plots for the continuous outcomes.
+We will make histogram plots for the continuous outcomes in our dataset.
 
-Height first.
+Percent of unintended pregnancies is the main outcome of interest, so we will explore this variable first.
 
 ```{r}
-p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
+p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + geom_histogram() 
 plot(p1)
-figure_file = here("results","height_distribution.png")
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_distribution.png")
 ggsave(filename = figure_file, plot=p1) 
@@
 
-Now weights.
+It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
+We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.
 
 ```{r}
-p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
+p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + geom_histogram() 
 plot(p2)
-figure_file = here("results","weight_distribution.png")
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png")
 ggsave(filename = figure_file, plot=p2)
+
+p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + geom_histogram() 
+plot(p3)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png")
+ggsave(filename = figure_file, plot=p3)
+
+p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + geom_histogram() 
+plot(p4)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png")
+ggsave(filename = figure_file, plot=p4)
+
+p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) + geom_histogram() 
+plot(p5)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png")
+ggsave(filename = figure_file, plot=p5) 
 ```
 
-Now height as function of weight.
+We will now explore the new 
+plots 6-10
+
+We will start exploring how many women use each type of birth control compared to the percent of unintended pregnancies related to that birth control.
+Plot 11 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
 
 ```{r}
-p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-plot(p3)
+p11 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
+plot(p11)
 figure_file = here("results","height_weight.png")
@@
 
+Our next group of plots will be violin plots to explore the usage level of birth associated with percentage of unintended pregnancies use for each birth control method. 
+```{r}
+p16 <- mydata %>% ggplot(aes(x=larcster_usage, y=pct_upreg_larcster))
+```
+
+Our last plots will be mosaic plots to compare two categorical variables: region and usage for each type of birth control. 
+```{r}
+library(ggmosaic)
+p17 <- mydata %>% ggplot(aes(x=region, y=larcster_usage, fill = freq)) + geom_mosaic(aes(x=region, fill = larcster_usage))
+```
+
 # Notes

### 2024-02-20 code/processing-code/processingfile-v1.qmd
Part 2 Updates
We fixed an error in one of the variables we created, total_larcster. We began our EDA with histograms and scatter plots with numeric variables.

@@
 rawdata2 <- rawdata %>% 
-                select(country, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534,  sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, sarc_3549, tradmethods_3549, nomethod_3549)
+                select(country, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)
 
@@
 We need to create a variable showing the total usage of each birth control method to remove the effect of age.
+It is worth noting here that the long acting reversible and sterilzation methods are combined in the`pct_upreg_larcster` variable, so we need to consider all the women using both of these methods in the total.
 ```{r}
@@
 rawdata2 <- rawdata2 %>% 
-              mutate(total_larc = larc_1519 + larc_2024 + larc_2534 + larc_3549,
+              mutate(total_larcster = larc_1519 + larc_2024 + larc_2534 + larc_3549 + ster_1519 + ster_2024 + ster_2534 +                          ster_3549,
                      total_sarc = sarc_1519 + sarc_2024 + sarc_2534 + sarc_3549,
@@
 # explore summary of each total to set boundaries for each level of usage
-rawdata2 %>% skim(total_larc, total_sarc, total_trad, total_none)
+rawdata2 %>% skim(total_larcster, total_sarc, total_trad, total_none)
 ```
@@
 rawdata2 <- rawdata2 %>% 
-              mutate(larc_usage = as.factor(case_when(
-                                  total_larc < 11.5 ~ "Low",
-                                  total_larc >= 11.5 & total_larc < 407.4633 ~ "Medium",
-                                  total_larc >= 407.4633 ~ "High")))
+              mutate(larcster_usage = as.factor(case_when(
+                                  total_larcster < 23.94865 ~ "Low",
+                                  total_larcster >= 23.94865 & total_larcster < 1554.7609 ~ "Medium",
+                                  total_larcster >= 1554.7609 ~ "High")))
 
 # check if new variable was created correctly 
-class(rawdata2$larc_usage)
+class(rawdata2$larcster_usage)
 
@@
                                   total_sarc < 114.33113 ~ "Low",
-                                  total_sarc >= 114.33113 & total_larc < 1554.7609 ~ "Medium",
+                                  total_sarc >= 114.33113 & total_sarc < 1554.7609 ~ "Medium",
                                   total_sarc >= 1554.7609	 ~ "High")))
@@
                                   total_trad < 15.05000	 ~ "Low",
-                                  total_trad >= 15.05000 & total_larc < 332.4750 ~ "Medium",
+                                  total_trad >= 15.05000 & total_trad < 332.4750 ~ "Medium",
                                   total_trad >= 332.4750	 ~ "High")))
@@
                                   total_none < 80.07500	 ~ "Low",
-                                  total_none >= 80.07500 & total_larc < 875.2750 ~ "Medium",
+                                  total_none >= 80.07500 & total_none < 875.2750 ~ "Medium",
                                   total_none >= 875.2750	 ~ "High")))

### 2024-02-21 code/eda-code/eda.qmd
Part 2 Updates
I added the skeleton code to produce the remainder of the exploratory plots. I made a note in the processing code about exploring the missing data in the outcome variable.

@@
 
-explore how many people use each method
-explore how many countries are in each category for each method using new factor variable
-explore percentage of unintended pregnancies across methods
-
 # Data exploration through tables
-We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information
+We will begin with creating a summary table of our cleaned variables using the skim()function.
+This provides information about each of the numeric variables including the mean, standard deviation, quartiles, missing observations, and a histogram.
 
@@
 saveRDS(summary_df, file = summarytable_file)
+```
+We also created simple tables for the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
+```{r}
+
 ```
@@
 ```{r}
-p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + geom_histogram() 
+p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + 
+  geom_histogram() + 
+  labs(x="Percent of Unintended Pregnancies", y="Frequency at Country Level")
 plot(p1)
@@
 It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
-We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.
+We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
 
@@
 
-We will now explore the new 
-plots 6-10
+We will now explore the new total usage variables that were created using the number of women using each method across the four age categories.
+This set of plots revealed large outliers in the 'total_larcster`, 'total_sarc', 'total_trad', and `total_none` variables, which prevented the graphs from being very useful.
+The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0.
+
+probably skewed by population size 
+```{r}
+p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram(binwidth = 5000) 
+plot(p6)
+figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
+ggsave(filename = figure_file, plot=p6)
+
+p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() 
+plot(p7)
+figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
+ggsave(filename = figure_file, plot=p7)
+
+p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() 
+plot(p8)
+figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
+ggsave(filename = figure_file, plot=p8)
+
+p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() 
+plot(p9)
+figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
+ggsave(filename = figure_file, plot=p9) 
+```
 
 We will start exploring how many women use each type of birth control compared to the percent of unintended pregnancies related to that birth control.
-Plot 11 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
+Plot 10 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
+Plot 11 uses `pct_upreg_sarc` and `total_sarc` to examine the same relationship as plot 10 but with short-acting reversible methods.
+Plot 12 shows `pct_upreg_trad` versus `total_trad`, and plot 13 shows `pct_upreg_nouse` versus `total_none`. 
+```{r}
+p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
+plot(p10)
+figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
+ggsave(filename = figure_file, plot=p10)
+
+p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=total_sarc)) + geom_point() + geom_smooth(method='lm')
+plot(p11)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
+ggsave(filename = figure_file, plot=p11)
+
+p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=total_trad)) + geom_point() + geom_smooth(method='lm')
+plot(p12)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
+ggsave(filename = figure_file, plot=p12)
+
+p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=total_none)) + geom_point() + geom_smooth(method='lm')
+plot(p13)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
+ggsave(filename = figure_file, plot=p13) 
+```
+
+We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region.
+We adjusted the original scatter plots created above by coloring based on region.
 
 ```{r}
-p11 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
-plot(p11)
-figure_file = here("results","height_weight.png")
-ggsave(filename = figure_file, plot=p3) 
+p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p14)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
+ggsave(filename = figure_file, plot=p14)
+
+p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=total_sarc, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p15)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
+ggsave(filename = figure_file, plot=p15)
+
+p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=total_trad, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p16)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
+ggsave(filename = figure_file, plot=p16)
+
+p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=total_none, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p17)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
+ggsave(filename = figure_file, plot=p17) 
 ```
 
-Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.
+Our next group of plots will be violin plots to explore the usage level of each birth control message associated with percentage of unintended pregnancies use for each birth control method. 
+```{r}
+p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
+  geom_violin() +
+  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
+plot(p18)
+figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p18)
 
-```{r}
-p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
-plot(p4)
-figure_file = here("results","height_weight_stratified.png")
-ggsave(filename = figure_file, plot=p4) 
-```
+p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
+  geom_violin() +
+  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
+plot(p19)
+figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p19)
 
-Our next group of plots will be violin plots to explore the usage level of birth associated with percentage of unintended pregnancies use for each birth control method. 
-```{r}
-p16 <- mydata %>% ggplot(aes(x=larcster_usage, y=pct_upreg_larcster))
-```
+p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
+  geom_violin() +
+  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
+plot(p20)
+figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p20)
 
-Our last plots will be mosaic plots to compare two categorical variables: region and usage for each type of birth control. 
-```{r}
-library(ggmosaic)
-p17 <- mydata %>% ggplot(aes(x=region, y=larcster_usage, fill = freq)) + geom_mosaic(aes(x=region, fill = larcster_usage))
+p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
+  geom_violin() +
+  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
+plot(p21)
+figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p21) 
 ```

### 2024-02-21 code/processing-code/processingfile-v1.qmd
Part 2 Updates
I added the skeleton code to produce the remainder of the exploratory plots. I made a note in the processing code about exploring the missing data in the outcome variable.

@@
 After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.
+`mice` package or `tidymodels`
+or remove because it is main outcome
+or make our own pct_upreg_updated and compare to what they have
 ```{r}

### 2024-02-21 code/processing-code/processingfile-v1.qmd
Part 2 Updates
We explored the pct_upreg variable and decided not to use it as our outcome variable because it is unreliable.

--- 2024-02-20 code/processing-code/processingfile-v1.qmd
@@
 library(here) #to set paths
-library(tidymodels) #to impute missing data
-library(recipes) #to impute missing data
 ```
@@
 
-We should impute the missing data here to avoid introducing bias because 18.18% of the data is missing, and this variable is essential for answering the research question.
-After finding the summary statistics of this variable, the median is 0.4701 and the mean is 0.4679, which suggests that there are minimal outliers in the data.
-```{r}
-#| message: false
-# find mean to use for imputed data
-summary(rawdata2$pct_upreg)
-
-# create a recipe to impute data 
-rec <- recipe(pct_upreg ~ ., data = rawdata2) %>%
-  step_naomit(all_predictors(), -all_outcomes(), skip = TRUE) %>%
-  step_impute_mean(all_predictors()) #which predictors would I use?
-
-# prepare the raw data using the recipe
-data_prep <- prep(rec, training = rawdata2)
-
-# impute the data 
-data_imputed <- bake(data_prep, new_data = NULL) #all pct_upreg are NaN values
-
-# identify rows with missing values in the original dataset
-missing_rows <- which(is.na(rawdata2$pct_upreg))
-
-# replace missing values in the original dataset with imputed values
-rawdata2$pct_upreg[missing_rows] <- data_imputed$pct_upreg[missing_rows]
-
-```
-Once the missing values have been imputed, we should further clean the `pct_upreg` variable to have a typical percentage format. 
-```{r}
-#| message: false
-# convert percentage variable to typical percentage format
-rawdata2 <- rawdata2 %>% 
-              mutate(pct_upreg = pct_upreg * 100)
-
-# ensure mutation worked correctly
-summary(rawdata2$pct_upreg)
-```
-
 We also need to evaluate the 108 missing entries of the percentage of unintended pregnancies by subregion `upreg_subreg`.
@@
 
-To make the covariate variables more useful, it makes sense to create a new factor variable denoting low, medium, and high usage of each birth control method by country.
+We should address the missing data in the general `pct_upreg` here to avoid introducing bias because 18.18% of this variable is missing, and it is essential for answering the research question.
+After reviewing the codebook, the researchers provided an external source for this variable, which we have included in the references.
+The numbers in this source do not match what is included in the data set.
+After exploring the methodology for the Adding It Up study, we found their original data that matches what is recorded in the dataset.
+When we tried to create our own percentage of unintended pregnancy variable, we found that each birth control method adds up to account for 100% of unintended pregnancies.
+Since there are 24 observations missing in what should be the total unintended pregnancy variable, it is likely that the variables with the percentages of unintended pregnancies by birth control method were imputed or estimated.
+Missingness in the collection of data could also explain why the long-acting reversible contraceptives and sterilization method were combined into one category.
+Based on this discovery, we modified the question that we want to answer with the data to use variables that are complete and reliable: `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.
+These variables will allow us to answer the question of is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods among all countries compared to no usage?
+
+```{r}
+#| message: false
+rawdata2 <- rawdata2 %>% 
+            mutate(pct_preg = (pct_upreg_larcster + pct_upreg_sarc + pct_upreg_trad + pct_upreg_nouse))
+summary(rawdata2$pct_preg)
+```
+
+To make the covariate variables more useful for answering our research question, it makes sense to create a new factor variable denoting low, medium, and high usage of each birth control method by country.
 First, we need to determine the boundaries for each level of usage by exploring how many women use each method across age groups.
 We need to create a variable showing the total usage of each birth control method to remove the effect of age.
-It is worth noting here that the long acting reversible and sterilzation methods are combined in the`pct_upreg_larcster` variable, so we need to consider all the women using both of these methods in the total.
+It is worth noting again that the long acting reversible and sterilization methods are combined in the`pct_upreg_larcster` variable, so we need to consider all the women using both of these methods in the total.
 ```{r}

### 2024-02-21 code/eda-code/eda.qmd
Part 2 Updates
Removed 2 template files that we did not need. Explored the outliers in the total_BCmethod variables.

@@
 
-Percent of unintended pregnancies is the main outcome of interest, so we will explore this variable first.
-
-```{r}
-p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + 
-  geom_histogram() + 
-  labs(x="Percent of Unintended Pregnancies", y="Frequency at Country Level")
-plot(p1)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_distribution.png")
-ggsave(filename = figure_file, plot=p1) 
-```
-
 It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
@@
 ```{r}
-p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram(binwidth = 5000) 
+p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram() 
 plot(p6)
@@
 ggsave(filename = figure_file, plot=p9) 
+```
+
+We need to explore where the outliers in these variables are coming from. 
+```{r}
+outliers <- mydata %>% filter(total_larcster > 25000)
 ```

### 2024-02-22 code/eda-code/eda.qmd
Part 2 Updates
Explored standardization methods to remove the effect of outliers before finding the population count variable `wra` which was used to standardize. Updated the processing file to include that variable and updated the manuscript with a more focused research question and the appropriate variables.

--- 2024-02-20 code/eda-code/eda.qmd
@@
 ---
-title: "An example exploratory analysis script"
-author: "Andreas Handel"
-date: "10/28/2021"
+title: "Exploratory Analysis Script"
+author: "Taylor Glass and Arlyn Santiago"
+date: "02/21/24"
 output: html_document
 ---
-
-This Quarto file loads the cleaned data and does some exploring.
-
-I'm only showing it the way where the code is included in the file. 
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
-
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
 
@@
 
-explore how many people use each method
-explore how many countries are in each category for each method using new factor variable
-explore percentage of unintended pregnancies across methods
-
 # Data exploration through tables
-We will begin with creating a summary table of our cleaned variables using the skim()function. This provides information
+We will begin with creating a summary table of our cleaned variables using the skim()function.
+This provides information about each of the numeric variables including the mean, standard deviation, quartiles, missing observations, and a histogram.
 
 ```{r}
+# skim the data to create summary file
 summary_df = skim(mydata)
@@
 ```
+We also created simple tables for the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
+```{r}
+
+```
 
@@
 
-Percent of unintended pregnancies is the main outcome of interest, so we will explore this variable first.
+It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
+We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
 
 ```{r}
-p1 <- mydata %>% ggplot(aes(x=pct_upreg)) + geom_histogram() 
-plot(p1)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_distribution.png")
-ggsave(filename = figure_file, plot=p1) 
-```
-
-It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
-We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.
-
-```{r}
+# explore the percentage of unintended pregnancies related to long acting reversible methods in each country
 p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + geom_histogram() 
@@
 
+# explore the percentage of unintended pregnancies related to short acting reversible methods in each country
 p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + geom_histogram() 
@@
 
+# explore the percentage of unintended pregnancies related to traditional methods in each country
 p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + geom_histogram() 
@@
 
+# explore the percentage of unintended pregnancies related to no birth control usage in each country
 p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) + geom_histogram() 
@@
 
-We will now explore the new 
-plots 6-10
-
-We will start exploring how many women use each type of birth control compared to the percent of unintended pregnancies related to that birth control.
-Plot 11 shows `pct_upreg_larcster` vs `total_larcster` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
+We will now explore the new total usage variables that were created using the number of women using each method across the four age categories.
+This set of plots revealed large outliers in the 'total_larcster`, 'total_sarc', 'total_trad', and `total_none` variables, which prevented the graphs from being very useful.
+The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0.
+This distortion is caused by the two large outliers in the data set.
 
 ```{r}
-p11 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=total_larcster)) + geom_point() + geom_smooth(method='lm')
-plot(p11)
-figure_file = here("results","height_weight.png")
-ggsave(filename = figure_file, plot=p3) 
+# explore the total number of people using long acting reversible methods in each country
+p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram() 
+plot(p6)
+figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
+ggsave(filename = figure_file, plot=p6)
+
+# explore the total number of people using short acting reversible methods in each country
+p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() 
+plot(p7)
+figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
+ggsave(filename = figure_file, plot=p7)
+
+# explore the total number of people using traditional methods in each country
+p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() 
+plot(p8)
+figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
+ggsave(filename = figure_file, plot=p8)
+
+# explore the total number of people not using birth control in each country
+p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() 
+plot(p9)
+figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
+ggsave(filename = figure_file, plot=p9) 
 ```
 
-Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.
+We need to explore where the outliers in these variables are coming from.
+China and India are the two countries that account for the outliers in all 4 variables, which is likely being skewed by population size.
+The first method of standardizing we tried is creating proportions using the total count for each birth control method type.
+After creating these standardized variables, the same issue with outliers remained because China and India still account for the largest proportion of observations.
+Next, we considered using a scaling factor.
+We attempted this approach with the `total_larcster` variable and found the same issues as before.
+The third attempt at standarization utilized a min-max scaling with z-score normalization, but the outlier issue remained.
+```{r}
+# find the countries that represent the outliers
+outliers <- mydata %>% filter(total_larcster > 25000, 
+                              total_sarc > 25000,
+                              total_trad > 5000,
+                              total_none > 10000)
+
+# standardization attempt 1
+mydata <- mydata %>% mutate(larcster_st = total_larcster / sum(total_larcster),
+                            sarc_st = total_sarc / sum(total_sarc),
+                            trad_st = total_trad / sum(total_trad),
+                            none_st = total_none / sum(total_none))
+
+# standardization attempt 2
+scaling_factor_larcster <- max(mydata$total_larcster)
+mydata <- mydata %>% mutate(standardized_larcster = total_larcster / scaling_factor_larcster)
+
+# standardization attempt 3
+min_count <- min(mydata$total_larcster)
+max_count <- max(mydata$total_larcster)
+scaled_counts <- (mydata$total_larcster - min_count) / (max_count - min_count)
+print(scaled_counts)
+mean_count <- mean(mydata$total_larcster)
+sd_count <- sd(mydata$total_larcster)
+z_score_counts <- (mydata$total_larcster - mean_count) / sd_count
+```
+
+After reviewing the raw data set again, we found the `wra` variable that accounts for the number of women of reproductive age (15-49) in each country.
+Once this variable was added to the original processed data set, we could use it to standardize the total counts for each birth control method by population size of women of reprodutive age.
+```{r}
+# population size standarization for total counts per birth control method
+mydata <- mydata %>% mutate(larcster_standard = total_larcster / wra,
+                            sarc_standard = total_sarc / wra,
+                            trad_standard = total_trad / wra, 
+                            none_standard = total_none / wra)
+```
+
+We will start exploring how many women use each type of birth control in each country compared to the percent of unintended pregnancies related to that birth control.
+Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
+Plot 11 uses `pct_upreg_sarc` and `sarc_standard` to examine the same relationship as plot 10 but with short-acting reversible methods.
+Plot 12 shows `pct_upreg_trad` versus `trad_standard`, and plot 13 shows `pct_upreg_nouse` versus `none_standard`.
+```{r}
+# explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
+p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + geom_point() + geom_smooth(method='lm')
+plot(p10)
+figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
+ggsave(filename = figure_file, plot=p10)
+
+# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
+p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + geom_point() + geom_smooth(method='lm')
+plot(p11)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
+ggsave(filename = figure_file, plot=p11)
+
+# explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
+p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + geom_point() + geom_smooth(method='lm')
+plot(p12)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
+ggsave(filename = figure_file, plot=p12)
+
+# explore percentage of unintended pregnancies compared to standardized count of women not using no method
+p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + geom_point() + geom_smooth(method='lm')
+plot(p13)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
+ggsave(filename = figure_file, plot=p13) 
+```
+
+We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region.
+We adjusted the original scatter plots created above by coloring based on region.
 
 ```{r}
-p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
-plot(p4)
-figure_file = here("results","height_weight_stratified.png")
-ggsave(filename = figure_file, plot=p4) 
+# explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
+p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p14)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
+ggsave(filename = figure_file, plot=p14)
+
+# explore percent unintended pregnancies using short acting reversible methods by region
+p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p15)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
+ggsave(filename = figure_file, plot=p15)
+
+# explore percent unintended pregnancies using traditional methods by region
+p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p16)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
+ggsave(filename = figure_file, plot=p16)
+
+# explore percent unintended pregnancies using no method by region
+p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+plot(p17)
+figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
+ggsave(filename = figure_file, plot=p17) 
 ```
 
-Our next group of plots will be violin plots to explore the usage level of birth associated with percentage of unintended pregnancies use for each birth control method. 
+Our next group of plots will be violin plots to explore the usage level of each birth control method, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
 ```{r}
-p16 <- mydata %>% ggplot(aes(x=larcster_usage, y=pct_upreg_larcster))
+# explore long acting reversible and sterilization method usage compared to percent of unintended pregnancies
+p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
+  geom_violin() +
+  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
+plot(p18)
+figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p18)
+
+# explore short acting reversible method usage compared to percent of unintended pregnancies
+p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
+  geom_violin() +
+  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
+plot(p19)
+figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p19)
+
+# explore traditional method usage compared to percent of unintended pregnancies
+p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
+  geom_violin() +
+  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
+plot(p20)
+figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p20)
+
+# explore no method usage compared to percent of unintended pregnancies
+p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
+  geom_violin() +
+  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
+plot(p21)
+figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
+ggsave(filename = figure_file, plot=p21) 
 ```
-
-Our last plots will be mosaic plots to compare two categorical variables: region and usage for each type of birth control. 
-```{r}
-library(ggmosaic)
-p17 <- mydata %>% ggplot(aes(x=region, y=larcster_usage, fill = freq)) + geom_mosaic(aes(x=region, fill = larcster_usage))
-```
-
-# Notes
-
-For your own explorations, tables and figures can be "quick and dirty".
-As long as you can see what's going on, there is no need to polish them.
-That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.).
-Those should look as nice, polished and easy to understand as possible.

### 2024-02-22 products/manuscript/Manuscript.qmd
Part 2 Updates
Explored standardization methods to remove the effect of outliers before finding the population count variable `wra` which was used to standardize. Updated the processing file to include that variable and updated the manuscript with a more focused research question and the appropriate variables.

@@
 # Research Questions
-1) What is the relationship between rates of unintended pregnancies and usage of contraceptive methods across various age groups?
-The main outcome variables will be total percentage of unintended pregnancies (`pct_upreg`). 
-The predictor variables will include percentage of unintended pregnancies across three types of birth control: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`).
-No method variables can serve as control groups for statistical tests.
+1) Is there a significant difference in the percentage of unplanned pregnancies between contraceptive methods among all countries? Does this difference vary by region?
+The main outcome variables will be total percentage of unintended pregnancies per birth control method (`pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`). 
+The control variable will be total percentage of unintended pregnancies among women not using birth control (`pct_upreg_none`)
+The predictor variables will include the total number of people women using each type of birth control standardized by population size, which will be utilized to create a factor variable showing low, medium, and high usage for each method per country.
 Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49. 
@@
 We hope to find relationships between subregions, rates of unintended pregnancies, and contraceptive methods to determine possible interventions that will increase the effectiveness of contraceptive methods.
-
-2) How does marriage status affect percentage of unintended pregnancies across different types of birth control?
-  -outcome variables:
-      unintended pregnancies, total number (pct_upreg)
-      unintended pregnancies, using larc and sterilization (pct_upreg_larcster)
-      unintended pregnancies, using short term reversible (pct_upreg_sarc)
-  -predictor variables:
-      Number wanting to avoid pregnancy, all married women (inneed_married)
-      Number using sterilization, all married women	(ster_married) **combine these two in processing
-      Number using long-acting reversible methods, all married women	(larc_married)
-      Number using short-acting methods, all married women	(sarc_married)
-      Number wanting to avoid pregnancy, formerly married women 	(inneed_formerlymar)
-      Number using sterilization, formerly married women	(ster_formerlymar)
-      Number using long-acting reversible methods, formerly married women	(larc_formerlymar)
-      Number using short-acting methods, formerly married women	(sarc_formerlymar)
-      Number wanting to avoid pregnancy, never married women 	(inneed_nevermar)
-      Number using all modern methods, never married women	(modmethods_nevermar)
-      Number using sterilization, never married women	(ster_nevermar)
-      Number using long-acting reversible methods, never married women	(larc_nevermar)
-      Number using short-acting methods, never married women	(sarc_nevermar)
 

### 2024-02-23 code/eda-code/eda.qmd
I added titles and axis labels to all 21 plots.
I also added brief trend descriptions for all plots. However, I was unsure how to analyze P9, P12, and P15.

@@
 ```{r}
-
+library(gtsummary)
+tabl1 <- mydata %>% select(`region`)
+tabl2 <- mydata %>% select(`larcster_usage`)
+tabl3 <- mydata %>% select(`sarc_usage`)
+tabl4 <- mydata %>% select(`none_usage`)
 ```
@@
 # explore the percentage of unintended pregnancies related to long acting reversible methods in each country
-p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + geom_histogram() 
+p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + geom_histogram() + ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Number of Countries") + ylab("Percent")
 plot(p2)
@@
 # explore the percentage of unintended pregnancies related to short acting reversible methods in each country
-p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + geom_histogram() 
+p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + geom_histogram() + ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Number of Countries") + ylab("Percent")
 plot(p3)
@@
 # explore the percentage of unintended pregnancies related to traditional methods in each country
-p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + geom_histogram() 
+p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + geom_histogram() + ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Number of Countries") + ylab("Percent")
 plot(p4)
@@
 # explore the percentage of unintended pregnancies related to no birth control usage in each country
-p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) + geom_histogram() 
+p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) + geom_histogram() + ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Number of Countries") + ylab("Percent")
 plot(p5)
@@
 ```
+P2: This histogram shows that the percentage of unintended pregnancies is very low, at around 2%.
+
+P3: This histogram shows that the percentage of unintended pregnancies is very low, even for people who do not use SARMs.
+For example, the graph shows that among people who used a SARM 30% of the time, about 2.8% experienced an unintended pregnancy.
+
+P4: This histogram shows that the percentage of unintended pregnancies decreases as the use of traditional methods increases.
+
+P5: This histogram shows that the percentage of unintended pregnancies among women who do not use any contraception varies widely by country.
 
@@
 # explore the total number of people using long acting reversible methods in each country
-p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram() 
+p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram() + ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Population") + ylab("Percent")
 plot(p6)
@@
 # explore the total number of people using short acting reversible methods in each country
-p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() 
+p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() + ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Population") + ylab("Percent")
 plot(p7)
@@
 # explore the total number of people using traditional methods in each country
-p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() 
+p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() + ggtitle("Total Usage of Traditional Methods per Country") + xlab("Population") + ylab("Percent")
 plot(p8)
@@
 # explore the total number of people not using birth control in each country
-p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() 
+p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() + ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Population") + ylab("Percent")
 plot(p9)
@@
 ```
+P6: There appears to be more countries with a lower percentage of LARC usage (around 20-30%) compared to countries with a higher percentage (around 60-70%).
+
+P7: There appears to be more countries with a lower population and higher percentage of SARC usage.
+
+P8: This histogram seems to suggest that there are variations in traditional method usage between different countries.
+
+P9: I'm not sure with this one.
 
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
-p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + geom_point() + geom_smooth(method='lm')
+p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Standardized Count of Women") + ylab("Percent")
 plot(p10)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
-p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + geom_point() + geom_smooth(method='lm')
+p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standardized Count of Women") + ylab("Percent")
 plot(p11)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
-p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + geom_point() + geom_smooth(method='lm')
+p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Standardized Count of Women") + ylab("Percent")
 plot(p12)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women not using no method
-p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + geom_point() + geom_smooth(method='lm')
+p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Standardized Count of Women") + ylab("Percent")
 plot(p13)
@@
 ```
+
+P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to decrease.
+
+P11: As the standardized count of women using contraception increases, the percentage of unintended pregnancies might decrease.
+There appears to be a great amount of variation within the standardized count.
+
+P12: I'm not sure with this one.
+
+P13: There appears to be a variation in unintended pregnancy rates across different countries.
 
@@
 # explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
-p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies Using LARCs and Sterilzation Methods By Region") + xlab("Number of Regions") + ylab("Percent")
 plot(p14)
@@
 # explore percent unintended pregnancies using short acting reversible methods by region
-p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies Using SARCs Methods By Region") + xlab("Number of Regions") + ylab("Percent")
 plot(p15)
@@
 # explore percent unintended pregnancies using traditional methods by region
-p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies Using Traditional Methods By Region") + xlab("Number of Regions") + ylab("Percent")
 plot(p16)
@@
 # explore percent unintended pregnancies using no method by region
-p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + geom_point() + geom_smooth(method='lm') + ggtitle("Percent of Unintended Pregnancies Using No Methods By Region") + xlab("Number of Regions") + ylab("Percent")
 plot(p17)
@@
 ```
+P14: There appears to be a negative correlation.
+As the standardized count of women using contraception increases, the percentage of unintended pregnancies tends to decrease.
+However, the strength of this relationship is difficult to determine.
+
+P15: I'm not sure.
+
+P16: There appears to be a positive correlation. As the percentage of women not using contraception increases, the percentage of unintended pregnancies also tends to increase.
+
+P17: There appears to be a greater overall variation in the No Methods usage per region. The strength of this relationship is difficult to determine.
 
@@
 p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
-  geom_violin() +
+  geom_violin() + ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
   labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
@@
 p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
-  geom_violin() +
+  geom_violin() + ggtitle("Usage of SARCs vs Percent of Unintended Pregnancies") +
   labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
@@
 p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
-  geom_violin() +
+  geom_violin() + ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
   labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
@@
 p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
-  geom_violin() +
+  geom_violin() +  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
   labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
@@
 ```
+P18: It would appear that countries with higher percentages of sterilized women might have lower percentages of unintended pregnancies.
+
+P19: It would appear that the distribution of SARC usage might be skewed towards lower percentages, suggesting that more countries have lower SARC usage compared to those with higher SARC usage.
+
+P20: It would appear that as the percentage of women who have ever used the traditional method increases, the percentage of unintended pregnancies tends to decrease.
+However, the strength of this relationship is difficult to determine.
+
+P21: It would appear that as the percentage of women who do not use a birth control method increases, the percentage of unintended pregnancies tends to inccrease.

### 2024-02-23 products/manuscript/Manuscript.qmd
I added titles and axis labels to all 21 plots.
I also added brief trend descriptions for all plots. However, I was unsure how to analyze P9, P12, and P15.

@@
 
-1. How does marriage status affect percentage of unintended pregnancies across different types of birth control?
+1. Is there a significant difference in the percentage of unplanned pregnancies between three  contraceptive methods among all countries compared to no usage?
 
-2. Is there a significant difference in the percentage of unplanned pregnancies between reversible contraceptives (LARCs?) among the age groups (20-24) and (35-49)?
+2. How does marriage status affect percentage of usage across three different types of birth control per region? (back-up question)
 
@@
 
-https://pubmed.ncbi.nlm.nih.gov/32981858/
-https://pubmed.ncbi.nlm.nih.gov/30687521/ #This study is set in Senegal 
-https://www.sciencedirect.com/science/article/pii/S001078241730478X 
-https://www.sciencedirect.com/science/article/abs/pii/S0020729211005406 #legal status article
+- https://pubmed.ncbi.nlm.nih.gov/32981858/
+- https://www.sciencedirect.com/science/article/pii/S001078241730478X 
+- https://www.sciencedirect.com/science/article/abs/pii/S0020729211005406 #legal status article
+- https://www.guttmacher.org/report/adding-it-up-investing-in-sexual-reproductive-health-2019-methodology (Guttmacher source)
+- https://pubmed.ncbi.nlm.nih.gov/35332057/ (original source of data)

### 2024-02-23 code/eda-code/eda.qmd
Part 2 Update
Updated the plot descriptions in the manuscript. Added additional exploratory data analysis conclusions. Cleaned up the template to include what is relevant to our project and updated the read me files.

--- 2024-02-22 code/eda-code/eda.qmd
@@
 library(here) #for data loading/saving
-library(dplyr)
-library(skimr)
-library(ggplot2)
+library(dplyr) # for basic piping functions
+library(skimr) # for summarizing data
+library(ggplot2) # for visualizing data
+library(gtsummary) # for creating tables
 ```
@@
 # skim the data to create summary file
-summary_df = skim(mydata)
+summary_df <- skim(mydata)
 print(summary_df)
 # save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
+summarytable_file <- here("results", "tables", "summarytable.rds")
 saveRDS(summary_df, file = summarytable_file)
 ```
-We also created simple tables for the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
+
+We also created a simple table with the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
 ```{r}
+# isolate categorical variables into their own data set
+mydata_cat <- mydata %>% select(region, larcster_usage, sarc_usage, trad_usage, none_usage)
 
+# create a summary table with counts of each level of each categorical variable
+summary_table <- tbl_summary(mydata_cat, statistic = all_categorical() ~ "{n}")
+print(summary_table)
 ```
@@
 # explore the percentage of unintended pregnancies related to long acting reversible methods in each country
-p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + geom_histogram() 
+p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + 
+  geom_histogram() +
+  ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
 plot(p2)
@@
 # explore the percentage of unintended pregnancies related to short acting reversible methods in each country
-p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + geom_histogram() 
+p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
+  geom_histogram() + 
+  ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
 plot(p3)
@@
 # explore the percentage of unintended pregnancies related to traditional methods in each country
-p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + geom_histogram() 
+p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + 
+  geom_histogram() + 
+  ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Percentage") + ylab("Number of Countries")
 plot(p4)
@@
 # explore the percentage of unintended pregnancies related to no birth control usage in each country
-p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) + geom_histogram() 
+p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) +
+  geom_histogram() + 
+  ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Percentage") + ylab("Number of countries")
 plot(p5)
@@
 ```
+P2: This histogram is highly skewed right, which shows that most countries experience a very low percentage of unintended pregnancies related to long acting reversible and sterilization methods.
+
+P3: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
+About half of the countries in the dataset report less than 20% of pregnancies are unintended when women use this form of birth control.
+
+P4: This histogram shows that about half of the countries reported less than 10% percent of unintended pregnancies associated with using traditional birth control methods
+
+P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
 
@@
 # explore the total number of people using long acting reversible methods in each country
-p6 <- mydata %>% ggplot(aes(x=total_larcster)) + geom_histogram() 
+p6 <- mydata %>% ggplot(aes(x=total_larcster)) + 
+  geom_histogram() + 
+  ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
 plot(p6)
@@
 # explore the total number of people using short acting reversible methods in each country
-p7 <- mydata %>% ggplot(aes(x=total_sarc)) + geom_histogram() 
+p7 <- mydata %>% ggplot(aes(x=total_sarc)) + 
+  geom_histogram() + 
+  ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
 plot(p7)
@@
 # explore the total number of people using traditional methods in each country
-p8 <- mydata %>% ggplot(aes(x=total_trad)) + geom_histogram() 
+p8 <- mydata %>% ggplot(aes(x=total_trad)) + 
+  geom_histogram() + 
+  ggtitle("Total Usage of Traditional Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
 plot(p8)
@@
 # explore the total number of people not using birth control in each country
-p9 <- mydata %>% ggplot(aes(x=total_none)) + geom_histogram() 
+p9 <- mydata %>% ggplot(aes(x=total_none)) + 
+  geom_histogram() + 
+  ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Number of Women") + ylab("Number of Countries")
 plot(p9)
@@
 ```
-
+The two outliers present in each of these graphs makes them very difficult to interpret.
 We need to explore where the outliers in these variables are coming from.
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
-p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + geom_point() + geom_smooth(method='lm')
+p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + 
+  geom_point() + 
+  geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p10)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
-p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + geom_point() + geom_smooth(method='lm')
+p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
+  geom_point() + 
+  geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p11)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
-p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + geom_point() + geom_smooth(method='lm')
+p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + 
+  geom_point() + 
+  geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p12)
@@
 # explore percentage of unintended pregnancies compared to standardized count of women not using no method
-p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + geom_point() + geom_smooth(method='lm')
+p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + 
+  geom_point() + 
+  geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p13)
@@
 ```
+
+P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to increase.
+There is minimal variation around the linear equation, excluding the two largest percentages accounted for by India and China.
+
+P11: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.
+
+P12: As percentage of unintended pregnancies associated with traditional methods increases, the total number of women using this method increases.
+There is minimal variation in the trend when percentage of unintended pregnancies is less than 8%, but there is more variation when the percentage is greater than 10%.
+
+P13: The same expected positive trend is observed for women who are not using a birth control method, but there is considerable variation in this graph across all percentages of unintended pregnancies.
+Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
 
@@
 # explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
-p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + 
+  geom_point() + geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p14)
@@
 # explore percent unintended pregnancies using short acting reversible methods by region
-p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
+  geom_point() + 
+  geom_smooth(method='lm') + 
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p15)
@@
 # explore percent unintended pregnancies using traditional methods by region
-p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + 
+  geom_point() + 
+  geom_smooth(method='lm') +  
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p16)
@@
 # explore percent unintended pregnancies using no method by region
-p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + geom_point() + geom_smooth(method='lm')
+p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + 
+  geom_point() + 
+  geom_smooth(method='lm') +   
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
 plot(p17)
@@
 ```
+P14: Among women using long acting reversible and sterilization methods, Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
+
+P15: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
+
+P16: Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
+
+P17: This graph is much different than the other three, which is to be expected at this point.
+Oceania has the highest percentage of unintended pregnancies when women are not using birth control.
+The slope of the relationship for each region appears to be similar in this graph with the exception of Oceania having a steeper slope.
+There are considerably more outliers in this graph with African countries
 
@@
   geom_violin() +
+  ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
   labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
@@
   geom_violin() + 
+  ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
   labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
@@
   geom_violin() +  
+  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
   labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
@@
 ```
+P18: The distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
+High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
+
+P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+
+P20: Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
+
+P21: While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
+The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.

### 2024-02-23 products/manuscript/Manuscript.qmd
Part 2 Update
Updated the plot descriptions in the manuscript. Added additional exploratory data analysis conclusions. Cleaned up the template to include what is relevant to our project and updated the read me files.

@@
 
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
+This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
 
@@
 
+# Project Part 2
+# Exploring and processing the data 
+The first priority when cleaning the data was analyzing missing values.
+We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
+We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
+We printed the summary table here becacuse it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
+```{r}
+# show summary table to numeric variables with misssing values
+print(summary_df)
+```
+We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage`.
+We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
+Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
+It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
+```{r}
+print(summary_table)
+```
+
+We created a multitude of exploratory plots to understand our data better.
+The histograms of percent of unintended pregnancies related to each method of birth control showed discrepancies in skewness.
+The `pct_upreg_larcster` and `pct_upreg_trad` variables are highly skewed right, which shows that more countries experience lower rates of unintended pregnancies when using these methods.
+The `pct_upreg_sarc` histogram is not as skewed.
+It shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
+The histogram for `pct_upreg_none` shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
+```{r}
+plot(p2)
+plot(p3)
+plot(p4)
+plot(p5)
+```
+
+The next set of variables we explored were the total counts of women using each birth control method.
+There were two large outliers that prevented the graphs from being useful.
+After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years).
+There is a positive association between the percentage of unintended pregnancies and the total number of women using that method of birth control for all three methods and those not using any method, which is to be expected.
+There is more variation among those using SARMs and traditional methods compared to those using long-acting reversible methods.
+Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
+```{r}
+plot(p10)
+plot(p11)
+plot(p12)
+plot(p13)
+```
+We stratified these four graphs by region to examine how the relationships differ across each of the 6 regions.
+The long acting reversible and sterilization methods graph is a little difficult to decipher, but Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
+The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
+The graph for women not using any birth control methods shows that Oceania has the highest percentage of unintended pregnancies.
+```{r}
+plot(p14)
+plot(p15)
+plot(p16)
+plot(p17)
+```
+The last set of plots we created to explore the data are violin plots using the categorical usage variables for each birth control method.
+The first violin plot shows that the distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
+High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
+It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
+While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
+The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
+```{r}
+plot(p18)
+plot(p19)
+plot(p20)
+plot(p21)
+```
+
 # Summary/Abstract
@@
 Then you can include like this:
-
 Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

### 2024-02-23 products/manuscript/Manuscript.qmd
Attempted Knit
I tried to knit the manuscript document, but I am getting an error that it cannot find an object that it was created in the eda.qmd file.

@@
 ```{r, echo=FALSE, message=FALSE}
-xcode-select --install
 # load a few R packages
-install.packages("dplyr")
-install.packages("naniar")
 library(here)
@@
 We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
-We printed the summary table here becacuse it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
+We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
 ```{r}
-# show summary table to numeric variables with misssing values
+# show summary table to numeric variables with missing values
 print(summary_df)

### 2024-02-23 code/eda-code/eda.qmd
Part 2 Updates
Fixed the manuscript to knit with the summary tables and exploratory plots.

@@
 print(summary_table)
+
+# save to file 
+summarytable_file <- here("results", "tables", "summarytablecat.rds")
+saveRDS(summary_table, file = summarytable_file)
 ```

### 2024-02-23 products/manuscript/Manuscript.qmd
Part 2 Updates
Fixed the manuscript to knit with the summary tables and exploratory plots.

@@
 ## Project Part 1
+
 # Data Source 
@@
 
-# Project Part 2
-# Exploring and processing the data 
-The first priority when cleaning the data was analyzing missing values.
-We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
-We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
-We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
-```{r}
-# show summary table to numeric variables with missing values
-print(summary_df)
-```
-We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage`.
-We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
-Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
-It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
-```{r}
-print(summary_table)
-```
-
-We created a multitude of exploratory plots to understand our data better.
-The histograms of percent of unintended pregnancies related to each method of birth control showed discrepancies in skewness.
-The `pct_upreg_larcster` and `pct_upreg_trad` variables are highly skewed right, which shows that more countries experience lower rates of unintended pregnancies when using these methods.
-The `pct_upreg_sarc` histogram is not as skewed.
-It shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
-The histogram for `pct_upreg_none` shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
-```{r}
-plot(p2)
-plot(p3)
-plot(p4)
-plot(p5)
-```
-
-The next set of variables we explored were the total counts of women using each birth control method.
-There were two large outliers that prevented the graphs from being useful.
-After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years).
-There is a positive association between the percentage of unintended pregnancies and the total number of women using that method of birth control for all three methods and those not using any method, which is to be expected.
-There is more variation among those using SARMs and traditional methods compared to those using long-acting reversible methods.
-Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
-```{r}
-plot(p10)
-plot(p11)
-plot(p12)
-plot(p13)
-```
-We stratified these four graphs by region to examine how the relationships differ across each of the 6 regions.
-The long acting reversible and sterilization methods graph is a little difficult to decipher, but Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
-The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
-The graph for women not using any birth control methods shows that Oceania has the highest percentage of unintended pregnancies.
-```{r}
-plot(p14)
-plot(p15)
-plot(p16)
-plot(p17)
-```
-The last set of plots we created to explore the data are violin plots using the categorical usage variables for each birth control method.
-The first violin plot shows that the distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
-High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
-It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
-Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
-While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
-The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
-```{r}
-plot(p18)
-plot(p19)
-plot(p20)
-plot(p21)
-```
-
 # Summary/Abstract
@@
 Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
+is an example of some - completely random/unrelated - schematic that was generated with Biorender.
 We store those figures in the `assets` folder.
@@
 #| echo: FALSE
-knitr::include_graphics(here("products","assets","antigen-recognition.png"))
+# knitr::include_graphics(here("products","assets","antigen-recognition.png")) #commented out for the sake of knitting the file at this stage in the project 
 ```
@@
 
-## Exploratory/Descriptive analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
-@tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-
+## Exploratory/Descriptive analysis (Project Part 2)
+The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
+We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
+We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
+We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
 ```{r}
 #| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
+#| tbl-cap: "Summary table of numeric variables"
 #| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
+resulttable<- readRDS("../../results/tables/summarytable.rds")
+kable(resulttable)
+```
+
+We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable2.
+We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
+Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
+It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
+```{r}
+#| label: tbl-summarytable2
+#| tbl-cap: "Summary table for categorical variables"
+#| echo: FALSE
+resulttable2 <- readRDS("../../results/tables/summarytablecat.rds")
+kable(resulttable2)
+```
+
+We created a multitude of exploratory plots to understand our data better.
+The histograms of percent of unintended pregnancies related to each method of birth control showed discrepancies in skewness which can be seen @exploratory-result.
+The `pct_upreg_larcster` and `pct_upreg_trad` variables are highly skewed right, which shows that more countries experience lower rates of unintended pregnancies when using these methods.
+The `pct_upreg_sarc` histogram is not as skewed.
+It shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
+The histogram for `pct_upreg_none` shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
+```{r}
+#| label: exploratory-result
+#| fig-cap: "Histograms of Percentage of Unintended Pregnancy Use by Birth Control Method"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png"))
+```
+
+The next set of variables we explored were the total counts of women using each birth control method.
+There were two large outliers that prevented the graphs from being useful.
+After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years).
+There is a positive association between the percentage of unintended pregnancies and the total number of women using that method of birth control for all three methods and those not using any method, which is to be expected.
+There is more variation among those using SARMs and traditional methods compared to those using long-acting reversible methods.
+Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
+These figures can be seen @exploratory-result2.
+```{r}
+#| label: exploratory-result2
+#| fig-cap: "Relationship between Percentage of Unintended Pregnancies and Standardized Population Totals"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png"))
+```
+
+We stratified these four graphs by region to examine how the relationships differ across each of the 6 regions which is shown in @exploratory-result3 .
+The long acting reversible and sterilization methods graph is a little difficult to decipher, but Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
+The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
+The graph for women not using any birth control methods shows that Oceania has the highest percentage of unintended pregnancies.
+```{r}
+#| label: exploratory-result3
+#| fig-cap: "Percentage of Unintended Pregnancies Stratified by Region"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png"))
+```
+
+The last set of plots we created to explore the data are violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4.
+The first violin plot shows that the distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
+High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
+It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
+While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
+The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
+```{r}
+#| label: exploratory-result4
+#| fig-cap: "Violin Plots with Birth Control Usage Levels"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png"))
 ```
@@
 
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+shows a scatterplot figure produced by one of the R scripts.
 
@@
 #| echo: FALSE
-knitr::include_graphics(here("results","height-weight-stratified.png"))
+# knitr::include_graphics(here("results","height-weight-stratified.png")) #commented out for the sake of knitting our project at this stage
 ```
@@
 #| echo: FALSE
-resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
-knitr::kable(resulttable2)
+# resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
+# knitr::kable(resulttable2) #commented out for the sake of knitting our project at this stage
 ```

### 2024-02-23 products/manuscript/Manuscript.qmd
Worked on errors preventing us to knit. Mainly fixed the heading

@@
     highlight-style: github
-bibliography: ../assets/dataanalysis-references.bib
-csl: ../assets/apa.csl
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/apa.csl
 ---

### 2024-03-11 code/analysis-code/statistical-analysis.qmd
Part 3 Updates
I started the statistical analysis and created 6 simple linear models. I added to the methods section in the manuscript and created a second version of the processed code based on the eda-code.

---
title: "Statistical Analysis Script"
author: "Taylor Glass and Arlyn Santiago"
date: "03/12/24"
output: html_document
---
## Load necessary packages
```{r}
#| message: false
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) # for model building 
```
## Load the cleaned data
```{r}
#path to data
data_location <- here("data","processed-data","processeddata2.rds")
#load data
mydata <- readRDS(data_location)
```

## Data fitting and statistical analysis 
The first trend we will investigate is if there is a stronger relationship between total usage of one method of birth control and percentage of unintended pregnancies associated with that method across the three birth control methods and no usage options. These models will allow us to predict unintended pregnancies based on population data about number of women using each type of contraception method per country.

### First Model Fit 
The first model attempts to predict the percentage of unintended pregnancies among women using long acting reversible  and sterilization methods of contraception, 'pct_upreg_larcster', based on the total number of women using this method of contraception. We will use the version of the variable that has been standardized by population size, larcster_standard.
```{r}
# fit linear model using pct_upreg_larcster as outcome, larcster_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit1 <- lm_mod %>% 
            fit(pct_upreg_larcster ~ larcster_standard, mydata) #estimate/train the model

#generate clean output with estimates and p-values
tidy(lmfit1)

# save results from fit into a data frame 
lmtable1 <- tidy(lmfit1)

# save fit results table  
table_file1 = here("results", "resulttable1.rds")
saveRDS(lmtable1, file = table_file1)
```

### Second Model Fit 
The second model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception. We will use the version of the variable that has been standardized by population size, 'sarc_standard'. 
```{r}
# fit linear model using pct_upreg_sarc as outcome, sarc_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit2 <- lm_mod %>% 
            fit(pct_upreg_sarc ~ sarc_standard, mydata) #estimate/train the model

#generate clean output with estimates and p-values
tidy(lmfit2)

# save results from fit into a data frame 
lmtable2 <- tidy(lmfit2)

# save fit results table  
table_file2 = here("results", "resulttable2.rds")
saveRDS(lmtable2, file = table_file2)
```
### Third Model Fit 
The second model attempts to predict the percentage of unintended pregnancies among women using traditional of contraception 'pct_upreg_trad' based on the total number of women using this method of contraception. We will use the version of the variable that has been standardized by population size, 'trad_standard'. 
```{r}
# fit linear model using pct_upreg_trad as outcome, trad_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit3 <- lm_mod %>% 
            fit(pct_upreg_trad ~ trad_standard, mydata) #estimate/train the model

# generate clean output with estimates and p-values
tidy(lmfit3)

# save results from fit into a data frame 
lmtable3 <- tidy(lmfit3)

# save fit results table  
table_file3 = here("results", "resulttable3.rds")
saveRDS(lmtable3, file = table_file3)
```
### Fourth Model Fit 
The fourth model attempts to predict the percentage of unintended pregnancies among women using no method of contraception 'pct_upreg_nouse' based on the total number of women not using a method of contraception. We will use the version of the variable that has been standardized by population size, 'none_standard'. 
```{r}
# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit4 <- lm_mod %>% 
            fit(pct_upreg_nouse ~ none_standard, mydata) #estimate/train the model

# generate clean output with estimates and p-values
tidy(lmfit4)

# save results from fit into a data frame 
lmtable4 <- tidy(lmfit4)

# save fit results table  
table_file4 = here("results", "resulttable4.rds")
saveRDS(lmtable4, file = table_file4)
```
Based on the coefficient estimates for each parameter, the association between total number of women not using a contraception method is most strongly correlated with percentage of unintended pregnancies among those not using contraceptive methods, followed by those who use traditional methods and short-acting reversible methods. The long acting reversible and sterilization methods shows the weakest association based on these four models.

We want to explore if a model with all the variables of total number of women using each type of contraceptive method would be more accurate in estimating percentage of unintended pregnancies.

### Fifth Model Fit 
```{r}
# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit5 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_larcster ~ larcster_standard + sarc_standard + trad_standard + none_standard, mydata)

#generate clean output with estimates and p-values
tidy(lmfit5)

# save results from fit into a data frame 
lmtable5 <- tidy(lmfit5)

# save fit results table  
table_file5 = here("results", "resulttable5.rds")
saveRDS(lmtable5, file = table_file5)
```
These results show that only 'larcster_standard' showed a positive relationship with percentage of unintended pregnancies associated with long acting reversible and standardization techniques. This is to be expected.

We also want to determine if there is a relationship between the region and percentage of unintended pregnancies associated with each method of contraception.

### Sixth Model Fit 
```{r}
# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
lm_mod <- linear_reg() #specify the type of model
lmfit6 <- lm_mod %>% #estimate/train the model 
            fit(pct_upreg_larcster ~ region, mydata)

#generate clean output with estimates and p-values
tidy(lmfit6)

# save results from fit into a data frame 
lmtable6 <- tidy(lmfit6)

# save fit results table  
table_file6 = here("results", "resulttable6.rds")
saveRDS(lmtable6, file = table_file6)
```

### 2024-03-11 code/eda-code/eda.qmd
Part 3 Updates
I started the statistical analysis and created 6 simple linear models. I added to the methods section in the manuscript and created a second version of the processed code based on the eda-code.

@@
 ```{r}
-#load needed packages. make sure they are installed.
+#load necessary packages
 library(here) #for data loading/saving
@@
 
-Load the data.
+# Load the data
 
 ```{r}
-#Path to data
+#path to data
 data_location <- here("data","processed-data","processeddata.rds")
@@
 ```{r}
-# population size standarization for total counts per birth control method
+# population size standardization for total counts per birth control method
 mydata <- mydata %>% mutate(larcster_standard = total_larcster / wra,
@@
 The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
+
+# Save data 
+Lastly, we will save this data with the new standardized total usage variables for each method of birth control as the second version of our processed data with only the variables necessary for the statistical analysis and modeling portion of the project.
+This new dataset is saved in the same location as the first processed dataset.
+```{r}
+# select variables necessary for statistical analysis/modeling 
+processeddata2 <- mydata %>% select(country, wra, region, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad,  pct_upreg_nouse, larcster_usage, sarc_usage, trad_usage, none_usage, larcster_standard, sarc_standard, trad_standard, none_standard)
+
+# save it to the same folder with the first version of the processed data
+save_data_location <- here("data","processed-data","processeddata2.rds")
+saveRDS(processeddata2, file = save_data_location)
+```

### 2024-03-11 code/processing-code/processing-file.qmd
Part 3 Updates
I started the statistical analysis and created 6 simple linear models. I added to the methods section in the manuscript and created a second version of the processed code based on the eda-code.

--- 2024-02-21 code/processing-code/processingfile-v1.qmd
@@
 
-code for low, medium, and high use per country using case_when()
-
 ```{r}
@@
 rawdata2 <- rawdata %>% 
-                select(country, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)
+                select(country, wra, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)
 
@@
 
-Fortunately, the rest of the variables required for this analysis are complete.
+Fortunately, the rest of the variables required for this analysis are complete including the `wra` variable with population size that we will need to standardize population counts later in the exploratory data analysis process.
 Each of the four variables describing the percentage of unintended pregnancies based on birth control type also need to be converted to traditional percentage format.
@@
 Based on this discovery, we modified the question that we want to answer with the data to use variables that are complete and reliable: `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse`.
-These variables will allow us to answer the question of is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods among all countries compared to no usage?
+These variables will allow us to answer the question of is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods among all countries compared to no usage.
 

### 2024-03-11 products/manuscript/Manuscript.qmd
Part 2 Error
Added sessionInfo and looked for reasons why others would get an error when rendering the manuscript.
Checking for errors
Found the version of Quarto we are using and checked that the renv package is updated.
Part 3 Updates
I started the statistical analysis and created 6 simple linear models. I added to the methods section in the manuscript and created a second version of the processed code based on the eda-code.

@@
 Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49.
+These variables will be combined across age groups to create a total usage variable for each method.
 -Number of women 15-19 using long-acting reversible methods `larc_1519`
@@
 
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies across age groups (20-24) and (35-49) to investigate the effectiveness of each contraceptive method.Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies.
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of each contraceptive method.
+Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies.
 According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group.
@@
 
-1. Is there a significant difference in the percentage of unplanned pregnancies between three  contraceptive methods among all countries compared to no usage?
-
-2. How does marriage status affect percentage of usage across three different types of birth control per region? (back-up question)
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
+1. Is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods and no usage between regions of the world?
 
@@
 _Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+Reproducing this project requires R, RStudio, and Microsoft word.
+The README files in each folder contain descriptions of the documents located in that folder to help guide users through the reproducible processes.
+Begin with the processing-file.qmd in the code > processing-code folder.
+This file loads the original dataset from the Guttmacher Institute's Adding It Up project, loads the codebook for the dataset, and glimpses at the data.
+We chose 30 variables to clean to find the percentage of unintended pregnancies based on each type of contraceptive methods: long acting reversible and sterilization methods, short acting reversible methods, traditional methods, and no contraceptive usage.
+After the missing data was explored, we chose to use the percentage of unintended pregnancies variables related to each method of birth control because the total variable had unexplained missingness.
+The processed data is saved in the data > processed-data folder as processeddata.rds.
+
+The next step is exploratory data analysis.
+The eda.qmd file in the code > eda-code folder contains all the code for this process.
+We produce a couple summary tables before we begin a simple exploration of trends in percentage of unintended pregnancies by country in our first set of histograms.
+The second set of histograms explore contraceptive usage methods by country.
+These graphs revealed outliers, so we standardized by population size for each country.
+The next sets of plots are line graphs showing total usage of contraceptive methods to percentage of unintended pregnancies associated with that method.
+All of these figures are saved in the Supplementary-Material.qmd file in the products > manuscript > supplement folder.
+The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing, which are included in this manuscript.
+The exploratory data analysis process led to another version of the processed data with standardized total usage variables saved as processeddata2.rds in the data > procsssed-data folder.
+
+Step three is the statistical analayis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder.
+
+The Manuscript.qmd file for this document can be found in the products > manuscript folder.
 

### 2024-03-13 code/eda-code/eda.qmd
Cleaning
I deleted the extra exploratory-analysis file and moved some of the processing code from eda.qmd to processing-qmd, so there is only one processed datset.

--- 2024-02-23 code/eda-code/eda.qmd
@@
 ```
-The two outliers present in each of these graphs makes them very difficult to interpret.
-We need to explore where the outliers in these variables are coming from.
-China and India are the two countries that account for the outliers in all 4 variables, which is likely being skewed by population size.
-The first method of standardizing we tried is creating proportions using the total count for each birth control method type.
-After creating these standardized variables, the same issue with outliers remained because China and India still account for the largest proportion of observations.
-Next, we considered using a scaling factor.
-We attempted this approach with the `total_larcster` variable and found the same issues as before.
-The third attempt at standarization utilized a min-max scaling with z-score normalization, but the outlier issue remained.
-```{r}
-# find the countries that represent the outliers
-outliers <- mydata %>% filter(total_larcster > 25000, 
-                              total_sarc > 25000,
-                              total_trad > 5000,
-                              total_none > 10000)
 
-# standardization attempt 1
-mydata <- mydata %>% mutate(larcster_st = total_larcster / sum(total_larcster),
-                            sarc_st = total_sarc / sum(total_sarc),
-                            trad_st = total_trad / sum(total_trad),
-                            none_st = total_none / sum(total_none))
-
-# standardization attempt 2
-scaling_factor_larcster <- max(mydata$total_larcster)
-mydata <- mydata %>% mutate(standardized_larcster = total_larcster / scaling_factor_larcster)
-
-# standardization attempt 3
-min_count <- min(mydata$total_larcster)
-max_count <- max(mydata$total_larcster)
-scaled_counts <- (mydata$total_larcster - min_count) / (max_count - min_count)
-print(scaled_counts)
-mean_count <- mean(mydata$total_larcster)
-sd_count <- sd(mydata$total_larcster)
-z_score_counts <- (mydata$total_larcster - mean_count) / sd_count
-```
-
-After reviewing the raw data set again, we found the `wra` variable that accounts for the number of women of reproductive age (15-49) in each country.
-Once this variable was added to the original processed data set, we could use it to standardize the total counts for each birth control method by population size of women of reprodutive age.
-```{r}
-# population size standarization for total counts per birth control method
-mydata <- mydata %>% mutate(larcster_standard = total_larcster / wra,
-                            sarc_standard = total_sarc / wra,
-                            trad_standard = total_trad / wra, 
-                            none_standard = total_none / wra)
-```
-
-We will start exploring how many women use each type of birth control in each country compared to the percent of unintended pregnancies related to that birth control.
+The outliers were explored and addressed in the processing-file.qmd in the code > processing-code folder.
+Now that we have standardized population variables, We will start exploring how many women use each type of birth control in each country compared to the percent of unintended pregnancies related to that birth control.
 Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.

### 2024-03-13 code/processing-code/processing-file.qmd
Cleaning
I deleted the extra exploratory-analysis file and moved some of the processing code from eda.qmd to processing-qmd, so there is only one processed datset.

@@
 ```
+The exploratory data analysis in the eda.qmd file in the code > eda-code folder revealed outliers in the data that need to be addressed.
+
+The two outliers make the histograms showing the distribution of total usage of birth control method by country very difficult to interpret.
+We need to explore where the outliers in these variables are coming from.
+```{r}
+# find the countries that represent the outliers
+outliers <- rawdata2 %>% filter(total_larcster > 25000, 
+                              total_sarc > 25000,
+                              total_trad > 5000,
+                              total_none > 10000)
+
+# print the outliers
+outliers
+```
+
+China and India are the two countries that account for the outliers in all 4 variables, which is likely being skewed by population size.
+
+The first method of standardizing we tried is creating proportions using the total count for each birth control method type.
+After creating these standardized variables, the same issue with outliers remained because China and India still account for the largest proportion of observations.
+Next, we considered using a scaling factor.
+We attempted this approach with the `total_larcster` variable and found the same issues as before.
+The third attempt at standarization utilized a min-max scaling with z-score normalization, but the outlier issue remained.
+```{r}
+# standardization attempt 1
+rawdata2 <- rawdata2 %>% mutate(larcster_st = total_larcster / sum(total_larcster),
+                            sarc_st = total_sarc / sum(total_sarc),
+                            trad_st = total_trad / sum(total_trad),
+                            none_st = total_none / sum(total_none))
+
+# standardization attempt 2
+scaling_factor_larcster <- max(rawdata2$total_larcster)
+rawdata2 <- rawdata2 %>% mutate(standardized_larcster = total_larcster / scaling_factor_larcster)
+
+# standardization attempt 3
+min_count <- min(rawdata2$total_larcster)
+max_count <- max(rawdata2$total_larcster)
+scaled_counts <- (rawdata2$total_larcster - min_count) / (max_count - min_count)
+print(scaled_counts)
+mean_count <- mean(rawdata2$total_larcster)
+sd_count <- sd(rawdata2$total_larcster)
+z_score_counts <- (rawdata2$total_larcster - mean_count) / sd_count
+```
+
+After reviewing the raw data set again, we found the `wra` variable that accounts for the number of women of reproductive age (15-49) in each country.
+Once this variable was added to the original processed data set, we could use it to standardize the total counts for each birth control method by population size of women of reproductive age.
+```{r}
+# population size standardization for total counts per birth control method
+rawdata2 <- rawdata2 %>% mutate(larcster_standard = total_larcster / wra,
+                            sarc_standard = total_sarc / wra,
+                            trad_standard = total_trad / wra, 
+                            none_standard = total_none / wra)
+```
 

### 2024-03-13 code/analysis-code/statistical-analysis.qmd
Additional Predictors
I used the codebook to find 5 more predictors for our outcome. I processed them and made sure there was no missingness. I did a little bit of EDA and created simple linear models with each of them.

@@
 ```
+
 ## Load the cleaned data
@@
 #path to data
-data_location <- here("data","processed-data","processeddata2.rds")
+data_location <- here("data","processed-data","processeddata.rds")
 #load data
@@
 ```
+### Seventh Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, pct_currentlymarried as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit7 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ pct_currentlymarried, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit7)
+
+# save results from fit into a data frame 
+lmtable7 <- tidy(lmfit7)
+
+# save fit results table  
+table_file7 = here("results", "resulttable7.rds")
+saveRDS(lmtable7, file = table_file7)
+```
+
+# Eighth Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, pct_nevermarried as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit8 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ pct_nevermarried, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit8)
+
+# save results from fit into a data frame 
+lmtable8 <- tidy(lmfit8)
+
+# save fit results table  
+table_file8 = here("results", "resulttable7.rds")
+saveRDS(lmtable8, file = table_file8)
+```
+
+# Ninth Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit9 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ rate_matdeaths, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit9)
+
+# save results from fit into a data frame 
+lmtable9 <- tidy(lmfit9)
+
+# save fit results table  
+table_file9 = here("results", "resulttable9.rds")
+saveRDS(lmtable9, file = table_file9)
+```
+
+# Tenth Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_upreg as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit10 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ rate_matdeaths_upreg, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit10)
+
+# save results from fit into a data frame 
+lmtable10 <- tidy(lmfit10)
+
+# save fit results table  
+table_file10 = here("results", "resulttable10.rds")
+saveRDS(lmtable10, file = table_file10)
+```
+
+# Eleventh Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_abs as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit11 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ rate_matdeaths_abs, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit11)
+
+# save results from fit into a data frame 
+lmtable11 <- tidy(lmfit11)
+
+# save fit results table  
+table_file11 = here("results", "resulttable9.rds")
+saveRDS(lmtable11, file = table_file11)
+```

### 2024-03-13 code/eda-code/eda.qmd
Additional Predictors
I used the codebook to find 5 more predictors for our outcome. I processed them and made sure there was no missingness. I did a little bit of EDA and created simple linear models with each of them.

@@
 The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
+
+We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`.
+Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers.
+```{r}
+# explore the region predictor
+p22 <- mydata %>% ggplot(aes(x=region)) + 
+  geom_bar() +
+  ggtitle("Countries in Each Region") + 
+  xlab("Region") + ylab("Number of Countries")
+plot(p22)
+figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
+ggsave(filename = figure_file, plot=p22)
+
+# explore the current marriage predictor
+p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Currently Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p23)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p23)
+
+# explore the never married predictor
+p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Never Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p24)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p24)
+
+# explore the total maternal mortality rate predictor
+p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
+  geom_histogram() +
+  ggtitle("Total Rate of Maternal Deaths") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p25)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p25)
+
+# explore the maternal mortality rate from unintended pregnancies predictor
+p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p26)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p26)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths_upreg > 6000)
+
+# explore the maternal mortality rate from abortions
+p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Abortions") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p27)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p27)
+
+# explore outliers in maternal mortality from abortions
+mydata %>% filter(rate_matdeaths_abs > 500)
+
+```
+
+P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers.
+We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad and South Sudan that account for the highest rates of maternal mortality in this category.
+
+P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.

### 2024-03-13 code/processing-code/processing-file.qmd
Additional Predictors
I used the codebook to find 5 more predictors for our outcome. I processed them and made sure there was no missingness. I did a little bit of EDA and created simple linear models with each of them.

@@
 # Identifying variables to be cleaned 
-We want to explore the relationship between unintended pregnancies and birth control methods across different age groups.
-We will use three demographic variable: `country`, `region`, and `subregion`.
-The variable for the total percentage of unintended pregnancies, `pct_upreg` will be the main outcome variable.
-We could also use `upreg_subreg` to explore unintended pregnancies by subregion.
-The predictor variables will include percentage of unintended pregnancies across three types of birth control: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`).
-No method variables can serve as control groups for statistical tests.
+We want to explore the relationship between unintended pregnancies and contraception methods.
+We will use three predictor variables: total number of women using the contraceptive method, region, number of safe abortions,  .
+The variable for the total percentage of unintended pregnancies, `pct_upreg` will be the main outcome variable with additional options for outcome variables of total percentage of unintended pregnancies by each contraceptive method: long acting reversible methods (`pct_upreg_larcster`), short term reversible methods (`pct_upreg_sarc`), traditional methods (`pct_upreg_trad`), and no method (`pct_upreg_nouse`).
+We will use multiple predictor variables including: total number of women using the contraceptive method, region, subregion, percentage of married women, percentage of women never married, total maternal deaths per 100,000 live births, deaths from unintended pregnancy per 100,000 live births, and deaths related to abortion per 100000 abortions.
 Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49.
+These variables will be summed to create total number of women using each type of contraception:
 -Number of women 15-19 using long-acting reversible methods `larc_1519`
@@
 rawdata2 <- rawdata %>% 
-                select(country, wra, region, subregion, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)
+                select(country, wra, region, subregion, pct_currentlymarried, pct_nevermarried, rate_matdeaths, rate_matdeaths_upreg, rate_matdeaths_abs, pct_upreg, upreg_subreg, pct_upreg_larcster, pct_upreg_sarc, pct_upreg_trad, pct_upreg_nouse, larc_1519, ster_1519, sarc_1519, tradmethods_1519, nomethod_1519, larc_2024, ster_2024, sarc_2024, tradmethods_2024, nomethod_2024, larc_2534, ster_2534, sarc_2534, tradmethods_2534,  nomethod_2534, larc_3549, ster_3549, sarc_3549, tradmethods_3549, nomethod_3549)
 
@@
 Fortunately, the rest of the variables required for this analysis are complete including the `wra` variable with population size that we will need to standardize population counts later in the exploratory data analysis process.
-Each of the four variables describing the percentage of unintended pregnancies based on birth control type also need to be converted to traditional percentage format.
+Each of the four variables describing the percentage of unintended pregnancies based on birth control type and the two variables described percentage of married and non-married women also need to be converted to traditional percentage format.
 ```{r}
@@
                      pct_upreg_trad = pct_upreg_trad * 100, 
-                     pct_upreg_nouse = pct_upreg_nouse * 100)
+                     pct_upreg_nouse = pct_upreg_nouse * 100,
+                     pct_currentlymarried = pct_currentlymarried * 100,
+                     pct_nevermarried = pct_nevermarried * 100)
 
@@
 summary(rawdata2$pct_upreg_nouse)
+summary(rawdata2$pct_currentlymarried)
+summary(rawdata2$pct_nevermarried)
 ```
@@
 ```
+
 After finding the distribution of number of women using each type of birth control, we can use that information to create a new factor variable denoting low, medium, and high usage for each method.

### 2024-03-14 code/analysis-code/statistical-analysis.qmd
Statistical Analysis Update
I cleaned this section to only include the one outcome we chose with models for each of the 7 predictors. I also added RMSE values for 2 models to begin comparing performance. I updated some of the manuscript to reflect these changes.

--- 2024-03-11 code/analysis-code/statistical-analysis.qmd
@@
 ## Data fitting and statistical analysis 
-The first trend we will investigate is if there is a stronger relationship between total usage of one method of birth control and percentage of unintended pregnancies associated with that method across the three birth control methods and no usage options.
-These models will allow us to predict unintended pregnancies based on population data about number of women using each type of contraception method per country.
+The outcome of interest in all of the following models is the percentage of unintended pregnancies associated with usage of short acting reversible contraception methods.
+The predictors used in our models include total number of women using short acting reversible methods standardized by population size, region, percent of women currently married, percent of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy, and rate of maternal deaths due to abortion.
+We are considering adding another predictor using a political freedom index for each country.
 
 ### First Model Fit 
-The first model attempts to predict the percentage of unintended pregnancies among women using long acting reversible  and sterilization methods of contraception, 'pct_upreg_larcster', based on the total number of women using this method of contraception.
-We will use the version of the variable that has been standardized by population size, larcster_standard.
+The second model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception.
+We will use the version of the variable that has been standardized by population size, 'sarc_standard'.
 ```{r}
-# fit linear model using pct_upreg_larcster as outcome, larcster_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, sarc_standard as predictor
 lm_mod <- linear_reg() #specify the type of model
 lmfit1 <- lm_mod %>% 
-            fit(pct_upreg_larcster ~ larcster_standard, mydata) #estimate/train the model
+            fit(pct_upreg_sarc ~ sarc_standard, mydata) #estimate/train the model
 
@@
 ```
+There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.
 
 ### Second Model Fit 
-The second model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception.
-We will use the version of the variable that has been standardized by population size, 'sarc_standard'.
 ```{r}
-# fit linear model using pct_upreg_sarc as outcome, sarc_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, region as predictor
 lm_mod <- linear_reg() #specify the type of model
-lmfit2 <- lm_mod %>% 
-            fit(pct_upreg_sarc ~ sarc_standard, mydata) #estimate/train the model
+lmfit2 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_larcster ~ region, mydata)
 
@@
 ```
+This model shows that the Latin America/the Carribean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region.
+The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.
+
 ### Third Model Fit 
-The second model attempts to predict the percentage of unintended pregnancies among women using traditional of contraception 'pct_upreg_trad' based on the total number of women using this method of contraception.
-We will use the version of the variable that has been standardized by population size, 'trad_standard'.
 ```{r}
-# fit linear model using pct_upreg_trad as outcome, trad_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, pct_currentlymarried as predictor
 lm_mod <- linear_reg() #specify the type of model
-lmfit3 <- lm_mod %>% 
-            fit(pct_upreg_trad ~ trad_standard, mydata) #estimate/train the model
+lmfit3 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ pct_currentlymarried, mydata)
 
-# generate clean output with estimates and p-values
+#generate clean output with estimates and p-values
 tidy(lmfit3)
@@
 ```
-### Fourth Model Fit 
-The fourth model attempts to predict the percentage of unintended pregnancies among women using no method of contraception 'pct_upreg_nouse' based on the total number of women not using a method of contraception.
-We will use the version of the variable that has been standardized by population size, 'none_standard'.
+This model shows there is a negative relationship between the number of women currently married and the outcome.
+As the percentage of currently married women increases by 1, the percentage of unintended pregnancies decreases by a factor of 0.4181571.
+
+# Fourth Model Fit 
 ```{r}
-# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, pct_nevermarried as predictor
 lm_mod <- linear_reg() #specify the type of model
-lmfit4 <- lm_mod %>% 
-            fit(pct_upreg_nouse ~ none_standard, mydata) #estimate/train the model
+lmfit4 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ pct_nevermarried, mydata)
 
-# generate clean output with estimates and p-values
+#generate clean output with estimates and p-values
 tidy(lmfit4)
@@
 ```
-Based on the coefficient estimates for each parameter, the association between total number of women not using a contraception method is most strongly correlated with percentage of unintended pregnancies among those not using contraceptive methods, followed by those who use traditional methods and short-acting reversible methods.
-The long acting reversible and sterilization methods shows the weakest association based on these four models.
+This model shows a positive relationship between the percentage of women never married and the outcome.
+As the percentage of women never married increases by 1, the percentage of unintended pregnancies increases by a factor of 0.2474218.
 
-We want to explore if a model with all the variables of total number of women using each type of contraceptive method would be more accurate in estimating percentage of unintended pregnancies.
-
-### Fifth Model Fit 
+# Fifth Model Fit 
 ```{r}
-# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths as predictor
 lm_mod <- linear_reg() #specify the type of model
 lmfit5 <- lm_mod %>% #estimate/train the model 
-            fit(pct_upreg_larcster ~ larcster_standard + sarc_standard + trad_standard + none_standard, mydata)
+            fit(pct_upreg_sarc ~ rate_matdeaths, mydata)
 
@@
 ```
-These results show that only 'larcster_standard' showed a positive relationship with percentage of unintended pregnancies associated with long acting reversible and standardization techniques.
-This is to be expected.
+This model shows that total rate of maternal deaths is slightly negatively correlated with the outcome based on a small coefficient estimate of -0.01482568.
 
-We also want to determine if there is a relationship between the region and percentage of unintended pregnancies associated with each method of contraception.
-
-### Sixth Model Fit 
+# Sixth Model Fit 
 ```{r}
-# fit linear model using pct_upreg_nouse as outcome, none_standard as predictor
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_upreg as predictor
 lm_mod <- linear_reg() #specify the type of model
 lmfit6 <- lm_mod %>% #estimate/train the model 
-            fit(pct_upreg_larcster ~ region, mydata)
+            fit(pct_upreg_sarc ~ rate_matdeaths_upreg, mydata)
 
@@
 ```
+This model shows that total rate of maternal deaths due to unintended pregnancies has an even weaker association with the outcome based on a small coefficient estimate of -0.003158542.
+
+# Seventh Model Fit 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome, rate_matdeaths_abs as predictor
+lm_mod <- linear_reg() #specify the type of model
+lmfit7 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ rate_matdeaths_abs, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit7)
+
+# save results from fit into a data frame 
+lmtable7 <- tidy(lmfit7)
+
+# save fit results table  
+table_file7 = here("results", "resulttable7.rds")
+saveRDS(lmtable7, file = table_file7)
+```
+This model shows that total rate of maternal deaths due to abortion has the strongest association with the outcome out of the three maternal mortality rates included based on a coefficient estimate of -0.033353527.
+
+### Eighth Model Fit 
+We will explore all 7 predictors using in a multiple linear regression model to determine how the associations  interact. 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome with all 7 predictors
+lm_mod <- linear_reg() #specify the type of model
+lmfit8 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + pct_nevermarried + rate_matdeaths + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit8)
+
+# save results from fit into a data frame 
+lmtable8 <- tidy(lmfit8)
+
+# save fit results table  
+table_file8 = here("results", "resulttable8.rds")
+saveRDS(lmtable8, file = table_file8)
+```
+This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433.
+The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor.
+The Oceania region now has a negative association with the outcome.
+Both of the marriage status predictors are associated with a decrease in the outcome.
+The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.
+
+## Assessing Model Performance
+It is important to consider appropriate metrics to measure model performance.
+We will use RMSE at this point to compare the model using the number of women using short acting reversible contraception predictor and the model using the percent currently married predictor.
+```{r}
+# create predictions using the sarc_standard predictor model
+poppred <- predict(lmfit1, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values of Y
+popdata <- bind_cols(poppred, mydata$sarc_standard) %>%
+            rename(sarc_standard = "...2")
+
+# find RMSE and R squared to determine model fit
+rmse1 <- rmse(popdata, truth = sarc_standard, estimate = .pred)
+
+# create predictions using the pct_currentlymarriedd predictor model
+marriedpred <- predict(lmfit3, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values of Y
+marrieddata <- bind_cols(marriedpred, mydata$pct_currentlymarried) %>%
+            rename(pct_currentlymarried = "...2")
+
+# find RMSE and R squared to determine model fit
+rmse2 <- rmse(marrieddata, truth = pct_currentlymarried, estimate = .pred)
+
+# compare the metrics 
+print(rmse1)
+print(rmse2)
+```
+
+The RMSE metric favors the model with the total number of women using the short acting reversible contraception method over the model with the percentage of women currently married, which is to be expected.

### 2024-03-14 products/manuscript/Manuscript.qmd
Statistical Analysis Update
I cleaned this section to only include the one outcome we chose with models for each of the 7 predictors. I also added RMSE values for 2 models to begin comparing performance. I updated some of the manuscript to reflect these changes.

@@
 This variable has 24 missing values, which will need to be cleaned.
-A simple visualization shows that African and Latin American/Carribean countries have the highest percentages of unintended pregnancies.
+A simple visualization shows that African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
 
 # Research Questions
-1) Is there a significant difference in the percentage of unplanned pregnancies between contraceptive methods among all countries? Does this difference vary by region?
-The main outcome variables will be total percentage of unintended pregnancies per birth control method (`pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`). 
-The control variable will be total percentage of unintended pregnancies among women not using birth control (`pct_upreg_none`)
-The predictor variables will include the total number of people women using each type of birth control standardized by population size, which will be utilized to create a factor variable showing low, medium, and high usage for each method per country.
-Covariates will include the number of women using each of these methods across four different age groups: 15-19, 20-24, 25-34, and 35-49.
-These variables will be combined across age groups to create a total usage variable for each method.
--Number of women 15-19 using long-acting reversible methods `larc_1519`
--Number of women 15-19 using short-acting methods `sarc_1519`
--Number of women 15-19 using traditional methods `tradmethods_1519`
--Number of women 15-19 using no method `nomethod_1519`
--Number of women 20-24 using long-acting reversible methods `larc_2024`
--Number of women 20-24 using short-acting methods	`sarc_2024`
--Number of women 20-24 using traditional methods	`tradmethods_2024`
--Number of women 20-24 using no method	`nomethod_2024`
--Number of women 25-34 using long-acting reversible methods	`larc_2534`
--Number of women 25-34 using short-acting methods	`sarc_2534`
--Number of women 25-34 using traditional methods	`tradmethods_2534`
--Number of women 25-34 using no method	`nomethod_2534`
--Number of women 35-49 using long-acting reversible method	`larc_3549`
--Number of women 35-49 using short-acting method	`sarc_3549`
--Number of women 35-49 using all traditional methods	`tradmethods_3549`
--Number of women 35-49 using no method	`nomethod_3549`
+1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries? 
+    The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).
 
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies across age groups to investigate the effectiveness of each method.
-We hope to find relationships between subregions, rates of unintended pregnancies, and contraceptive methods to determine possible interventions that will increase the effectiveness of contraceptive methods.
+    The predictor variables will include the total number of people women using each type of birth control                standardized by population size, region, percentage of women currently married, percentage of women never married,     total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due     to abortion.
+
+    A full list of variables can be found in the processing-file in the code > processing-code folder.
 
 # How to analyze 
-All of the variables used in this analysis are numeric, so we can build linear regression models to analyze how each method of contraception affects the percentage of unintended pregnancies.
+Most of the predictors used in this analysis are numeric, so we can build linear regression models to analyze how each predictor affects the percentage of unintended pregnancies among women using short acting reversible methods.
 We can group countries together based on the subregion variable to compare how these relationships vary across the world as shown in the crude visualization above.

### 2024-03-14 code/analysis-code/statistical-analysis.qmd
Manuscript
Finished describing how to reproduce the project up to this point and added the statistical analysis in the manuscript.

@@
 # save fit results table  
-table_file1 = here("results", "resulttable1.rds")
+table_file1 = here("results", "tables", "resulttable1.rds")
 saveRDS(lmtable1, file = table_file1)
@@
 # save fit results table  
-table_file2 = here("results", "resulttable2.rds")
+table_file2 = here("results", "tables", "resulttable2.rds")
 saveRDS(lmtable2, file = table_file2)
@@
 # save fit results table  
-table_file3 = here("results", "resulttable3.rds")
+table_file3 = here("results", "tables", "resulttable3.rds")
 saveRDS(lmtable3, file = table_file3)
@@
 # save fit results table  
-table_file4 = here("results", "resulttable4.rds")
+table_file4 = here("results", "tables", "resulttable4.rds")
 saveRDS(lmtable4, file = table_file4)
@@
 # save fit results table  
-table_file5 = here("results", "resulttable5.rds")
+table_file5 = here("results", "tables", "resulttable5.rds")
 saveRDS(lmtable5, file = table_file5)
@@
 # save fit results table  
-table_file6 = here("results", "resulttable6.rds")
+table_file6 = here("results", "tables", "resulttable6.rds")
 saveRDS(lmtable6, file = table_file6)
@@
 # save fit results table  
-table_file7 = here("results", "resulttable7.rds")
+table_file7 = here("results", "tables", "resulttable7.rds")
 saveRDS(lmtable7, file = table_file7)
@@
 # save fit results table  
-table_file8 = here("results", "resulttable8.rds")
+table_file8 = here("results", "tables", "resulttable8.rds")
 saveRDS(lmtable8, file = table_file8)

### 2024-03-14 products/manuscript/Manuscript.qmd
Manuscript
Finished describing how to reproduce the project up to this point and added the statistical analysis in the manuscript.

@@
 1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?
+
 The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).
@@
 
-1. Is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods and no usage between regions of the world?
+1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?
+
+The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).
+
+The predictor variables will include the total number of people women using each type of birth control                standardized by population size, region, percentage of women currently married, percentage of women never married,    total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due    to abortion.
+
+A full list of variables can be found in the processing-file in the code > processing-code folder.
 
@@
 The eda.qmd file in the code > eda-code folder contains all the code for this process.
-We produce a couple summary tables before we begin a simple exploration of trends in percentage of unintended pregnancies by country in our first set of histograms.
+We produced a couple summary tables before we began a simple exploration of trends in percentage of unintended pregnancies by country in our first set of histograms.
 The second set of histograms explore contraceptive usage methods by country.
@@
 The next sets of plots are line graphs showing total usage of contraceptive methods to percentage of unintended pregnancies associated with that method.
+The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing.
 All of these figures are saved in the Supplementary-Material.qmd file in the products > manuscript > supplement folder.
-The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing, which are included in this manuscript.
-The exploratory data analysis process led to another version of the processed data with standardized total usage variables saved as processeddata2.rds in the data > procsssed-data folder.
+This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods as our single outcome because this method is becoming increasingly popular, so we feel that results from this contraception method will be the most relevant.
 
-Step three is the statistical analayis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder.
+Step three is the statistical analysis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder.
+We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion.
+We also assess the performance of the models in this file.
+Result tables for each model are stored in the results > tables folder.
 
@@
 _Explain anything related to your statistical analyses._
+We began with a simple linear regression model with each of the 7 predictor variables.
+We also included a multiple linear regression model with all 7 predictors.
+We included the RMSE for two of the models to begin exploring model fit.
 
@@
 
-shows a scatterplot figure produced by one of the R scripts.
+We completed a simple linear regression analysis with each of the 7 predictors.
+The predictor with the strongest relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
 
 ```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| label: basic-statistical-analysis
+#| fig-cap: "Strongest Predictor: sarc_standard"
 #| echo: FALSE
-# knitr::include_graphics(here("results","height-weight-stratified.png")) #commented out for the sake of knitting our project at this stage
+readRDS(here("results","tables", "resulttable1.rds")) 
 ```
+The model using the region predictor revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region.
+The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
+```{r}
+#| label: basic-statistical-analysis
+#| fig-cap: "Regional Predictor"
+#| echo: FALSE
+readRDS(here("results","tables", "resulttable2.rds")) 
+```
+We created one model with all 7 predictors using multiple linear regression.
+This model shows the strongest predictor is the number of women using short acting reversible contraception.
+The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor.
+The Oceania region now has a negative association with the outcome.
+Both of the marriage status predictors are associated with a decrease in the outcome.
+The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.
+```{r}
+#| label: basic-statistical-analysis
+#| fig-cap: "Multiple Linear Regression with 7 Predictors"
+#| echo: FALSE
+readRDS(here("results","tables", "resulttable8.rds")) 
+```
+The full statistical analysis can be found in the statistical-analysis.qmd file in the code > analysis-code folder.
 

### 2024-03-15 code/eda-code/eda.qmd
Moved all plots about other birth control methods to Supp.
I also added information under Data Acquisition and created a table for the predictor variables. However, I was having wifi errors so let me know if you can see the changes.

--- 2024-03-13 code/eda-code/eda.qmd
@@
 
-It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well.
-We will create a histogram for the `pct_upreg_larcster`, `pct_upreg_sarc`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
+It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well. We will create a histogram for the `pct_upreg_sarc` variable.
 
 ```{r}
-# explore the percentage of unintended pregnancies related to long acting reversible methods in each country
-p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + 
-  geom_histogram() +
-  ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
-plot(p2)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png")
-ggsave(filename = figure_file, plot=p2)
 
-# explore the percentage of unintended pregnancies related to short acting reversible methods in each country
+# Explore the percentage of unintended pregnancies related to short acting reversible methods in each country
 p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
@@
 
-# explore the percentage of unintended pregnancies related to traditional methods in each country
-p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + 
-  geom_histogram() + 
-  ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Percentage") + ylab("Number of Countries")
-plot(p4)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png")
-ggsave(filename = figure_file, plot=p4)
-
-# explore the percentage of unintended pregnancies related to no birth control usage in each country
-p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) +
-  geom_histogram() + 
-  ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Percentage") + ylab("Number of countries")
-plot(p5)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png")
-ggsave(filename = figure_file, plot=p5) 
 ```
-P2: This histogram is highly skewed right, which shows that most countries experience a very low percentage of unintended pregnancies related to long acting reversible and sterilization methods.
 
@@
 
-P4: This histogram shows that about half of the countries reported less than 10% percent of unintended pregnancies associated with using traditional birth control methods
+We will now explore the new total usage variables that were created using the number of women using each method across the 'total_sarc' variable. 
+```{r}
 
-P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
-
-We will now explore the new total usage variables that were created using the number of women using each method across the four age categories.
-This set of plots revealed large outliers in the 'total_larcster`, 'total_sarc', 'total_trad', and `total_none` variables, which prevented the graphs from being very useful.
-The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0.
-This distortion is caused by the two large outliers in the data set.
-
-```{r}
-# explore the total number of people using long acting reversible methods in each country
-p6 <- mydata %>% ggplot(aes(x=total_larcster)) + 
-  geom_histogram() + 
-  ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
-plot(p6)
-figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
-ggsave(filename = figure_file, plot=p6)
-
-# explore the total number of people using short acting reversible methods in each country
+# Explore the total number of people using Short Acting Reversible Methods in each country
 p7 <- mydata %>% ggplot(aes(x=total_sarc)) + 
@@
 
-# explore the total number of people using traditional methods in each country
-p8 <- mydata %>% ggplot(aes(x=total_trad)) + 
-  geom_histogram() + 
-  ggtitle("Total Usage of Traditional Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
-plot(p8)
-figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
-ggsave(filename = figure_file, plot=p8)
-
-# explore the total number of people not using birth control in each country
-p9 <- mydata %>% ggplot(aes(x=total_none)) + 
-  geom_histogram() + 
-  ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Number of Women") + ylab("Number of Countries")
-plot(p9)
-figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
-ggsave(filename = figure_file, plot=p9) 
 ```
@@
 Now that we have standardized population variables, We will start exploring how many women use each type of birth control in each country compared to the percent of unintended pregnancies related to that birth control.
-Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
+
 Plot 11 uses `pct_upreg_sarc` and `sarc_standard` to examine the same relationship as plot 10 but with short-acting reversible methods. 
-Plot 12 shows `pct_upreg_trad` versus `trad_standard`, and plot 13 shows `pct_upreg_nouse` versus `none_standard`.
 ```{r}
-# explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
-p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + 
-  geom_point() + 
-  geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p10)
-figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
-ggsave(filename = figure_file, plot=p10)
 
@@
 
-# explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
-p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + 
-  geom_point() + 
-  geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p12)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
-ggsave(filename = figure_file, plot=p12)
-
-# explore percentage of unintended pregnancies compared to standardized count of women not using no method
-p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + 
-  geom_point() + 
-  geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p13)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
-ggsave(filename = figure_file, plot=p13) 
 ```
-
-P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to increase.
-There is minimal variation around the linear equation, excluding the two largest percentages accounted for by India and China.
 
@@
 
-P12: As percentage of unintended pregnancies associated with traditional methods increases, the total number of women using this method increases.
-There is minimal variation in the trend when percentage of unintended pregnancies is less than 8%, but there is more variation when the percentage is greater than 10%.
-
-P13: The same expected positive trend is observed for women who are not using a birth control method, but there is considerable variation in this graph across all percentages of unintended pregnancies.
-Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
-
-We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region.
+We are also interested in how these relationships between percent of unintended pregnancies using Short Acting Reversible Methods and the number of people using each method varies across region.
 We adjusted the original scatter plots created above by coloring based on region.
@@
 ```{r}
-# explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
-p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + 
-  geom_point() + geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p14)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
-ggsave(filename = figure_file, plot=p14)
 
@@
 
-# explore percent unintended pregnancies using traditional methods by region
-p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + 
-  geom_point() + 
-  geom_smooth(method='lm') +  
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p16)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
-ggsave(filename = figure_file, plot=p16)
-
-# explore percent unintended pregnancies using no method by region
-p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + 
-  geom_point() + 
-  geom_smooth(method='lm') +   
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p17)
-figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
-ggsave(filename = figure_file, plot=p17) 
 ```
-P14: Among women using long acting reversible and sterilization methods, Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
 
@@
 
-P16: Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
-
-P17: This graph is much different than the other three, which is to be expected at this point.
-Oceania has the highest percentage of unintended pregnancies when women are not using birth control.
-The slope of the relationship for each region appears to be similar in this graph with the exception of Oceania having a steeper slope.
-There are considerably more outliers in this graph with African countries
-
-Our next group of plots will be violin plots to explore the usage level of each birth control method, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
+Our next group of plots will be violin plots to explore the usage level of Short Acting Reversible Methods, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
 ```{r}
-# explore long acting reversible and sterilization method usage compared to percent of unintended pregnancies
-p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
-  geom_violin() +
-  ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
-  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
-plot(p18)
-figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
-ggsave(filename = figure_file, plot=p18)
 
@@
 
-# explore traditional method usage compared to percent of unintended pregnancies
-p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
-  geom_violin() + 
-  ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
-  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
-plot(p20)
-figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
-ggsave(filename = figure_file, plot=p20)
-
-# explore no method usage compared to percent of unintended pregnancies
-p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
-  geom_violin() +  
-  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
-  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
-plot(p21)
-figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
-ggsave(filename = figure_file, plot=p21) 
 ```
-P18: The distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
-High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
 
 P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
-
-P20: Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
-
-P21: While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
-The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.

### 2024-03-15 products/manuscript/Manuscript.qmd
Moved all plots about other birth control methods to Supp.
I also added information under Data Acquisition and created a table for the predictor variables. However, I was having wifi errors so let me know if you can see the changes.

@@
 
+```{r}
+# Load libraries
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+# Descriptive statistics for numeric variables
+pred_summary <- mydata %>%
+  summarise(across(c(wra, pct_currentlymarried,pct_nevermarried,rate_matdeaths, rate_matdeaths_upreg, rate_matdeaths_abs), mean, sd))
+
+# Descriptive statistics for categorical variable (assuming region is a factor)
+summary_factor <- mydata %>%
+  group_by(region) %>%
+  summarise(n = n())
+
+# Combine into a single dataframe
+summary_pred <- rbind(
+  pred_summary %>% summarise(Variable = "pct_upreg_sarc", ., .names = quote(names(.))),
+  pred_summary %>% select(-pct_upreg_sarc),
+  summary_factor %>% summarise(Variable = "region", n)
+)
+
+# Create table 
+kable(summary_pred, caption = "Table Depicting Predictor Variables", booktabs = TRUE) %>%
+  kable_styling(full_width = F, font_size = 10) %>%
+  column_def(widths = c(200, 150, 150))
+
+```
+
+We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`.
+Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers.
+```{r}
+# explore the region predictor
+p22 <- mydata %>% ggplot(aes(x=region)) + 
+  geom_bar() +
+  ggtitle("Countries in Each Region") + 
+  xlab("Region") + ylab("Number of Countries")
+plot(p22)
+figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
+ggsave(filename = figure_file, plot=p22)
+
+# explore the current marriage predictor
+p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Currently Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p23)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p23)
+
+# explore the never married predictor
+p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Never Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p24)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p24)
+
+# explore the total maternal mortality rate predictor
+p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
+  geom_histogram() +
+  ggtitle("Total Rate of Maternal Deaths") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p25)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p25)
+
+# explore the maternal mortality rate from unintended pregnancies predictor
+p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p26)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p26)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths_upreg > 6000)
+
+# explore the maternal mortality rate from abortions
+p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Abortions") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p27)
+figure_file = here("results", "figures", "exploratory-figures", "married_distribution.png")
+ggsave(filename = figure_file, plot=p27)
+
+# explore outliers in maternal mortality from abortions
+mydata %>% filter(rate_matdeaths_abs > 500)
+
+```
+
+P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers.
+We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad and South Sudan that account for the highest rates of maternal mortality in this category.
+
+P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.
+
 The Manuscript.qmd file for this document can be found in the products > manuscript folder.
@@
 _As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+
+This research aims to analyze the factors influencing unintended pregnancies and contraception use in low-and-middle-income countries.
+The Guttmacher Institute's Adding It Up project provides the foundation, offering data on sexual and reproductive health services in these regions.Focusing on women of reproductive age (15-49), the study usese information from nationally representative surveys and population projections to estimate unintended pregnancy rates.
+
+While data on the overall percentage of unintended pregnancies was unavailable, we strategically utilized data on rates related to specific birth control methods.
+This approach allows us to analyze the relationship between these rates and various social and health factors.
+Factors like the number of women utilizing different birth control methods, maternal death rates, and marital status will be meticulously examined.
+All 7 of these predictor variables were found within the study's codebook, which can can be found in the `Data` folder of this project.
+By discovering these relationships, the research aims to shed light on the complex landscape of contraception use and unintended pregnancy in low-and -middle-income countries.
+This knowledge can empower policymakers and healthcare providers to make informed decisions that ultimately improve reproductive health outcomes for women.
 

### 2024-03-15 products/manuscript/supplement/Supplementary-Material.qmd
Moved all plots about other birth control methods to Supp.
I also added information under Data Acquisition and created a table for the predictor variables. However, I was having wifi errors so let me know if you can see the changes.

---
title: "Supplement to Global Distribution of Unintended Pregnancies and Contraceptive Usage Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis_template_references.bib
csl: ../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
#load necessary packages
library(here) #for data loading/saving
library(dplyr) # for basic piping functions
library(skimr) # for summarizing data
library(ggplot2) # for visualizing data
library(gtsummary) # for creating tables
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

# Overview

Readers will be able to find our analysis of the different contraception methods through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.

```{r}
# explore the percentage of unintended pregnancies related to long acting reversible methods in each country
p2 <- mydata %>% ggplot(aes(x=pct_upreg_larcster)) + 
  geom_histogram() +
  ggtitle("Percent Unintended Pregnancy Using Long Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p2)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png")
ggsave(filename = figure_file, plot=p2)

# explore the percentage of unintended pregnancies related to traditional methods in each country
p4 <- mydata %>% ggplot(aes(x=pct_upreg_trad)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Traditional Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png")
ggsave(filename = figure_file, plot=p4)

# explore the percentage of unintended pregnancies related to no birth control usage in each country
p5 <- mydata %>% ggplot(aes(x=pct_upreg_nouse)) +
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using No Birth Control") + xlab("Percentage") + ylab("Number of countries")
plot(p5)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png")
ggsave(filename = figure_file, plot=p5) 
```

P2: This histogram is highly skewed right, which shows that most countries experience a very low percentage of unintended pregnancies related to long acting reversible and sterilization methods.

P4: This histogram shows that about half of the countries reported less than 10% percent of unintended pregnancies associated with using traditional birth control methods

P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.

The following set of plots revealed large outliers in the 'total_larcster`,'total_trad', and `total_none` variables, which prevented the graphs from being very useful. The histograms look like most observations for each of these four variables are 0, but that is not possible because the summary table showed that the minimum value for all four of these variables is greater than 0. This distortion is caused by the two large outliers in the data set.

```{r}
# explore the total number of people using long acting reversible methods in each country
p6 <- mydata %>% ggplot(aes(x=total_larcster)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Long Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p6)
figure_file = here("results", "figures", "exploratory-figures", "total_larcster_distribution.png")
ggsave(filename = figure_file, plot=p6)

# explore the total number of people using traditional methods in each country
p8 <- mydata %>% ggplot(aes(x=total_trad)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Traditional Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p8)
figure_file = here("results", "figures", "exploratory-figures", "total_trad_distribution.png")
ggsave(filename = figure_file, plot=p8)

# explore the total number of people not using birth control in each country
p9 <- mydata %>% ggplot(aes(x=total_none)) + 
  geom_histogram() + 
  ggtitle("Total Usage of No Birth Control Method per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p9)
figure_file = here("results", "figures", "exploratory-figures", "total_none_distribution.png")
ggsave(filename = figure_file, plot=p9)

```
Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.

```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using long-acting reversible and sterilization methods
p10 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p10)
figure_file = here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png")
ggsave(filename = figure_file, plot=p10) 
```

P10: As the standardized count of women using LARCs and sterilization methods increases, the percentage of unintended pregnancies tends to increase. There is minimal variation around the linear equation, excluding the two largest percentages accounted for by India and China.

Plot 12 shows `pct_upreg_trad` versus `trad_standard` 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using traditional methods
p12 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p12)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png")
ggsave(filename = figure_file, plot=p12) 
```

P12: As percentage of unintended pregnancies associated with traditional methods increases, the total number of women using this method increases. There is minimal variation in the trend when percentage of unintended pregnancies is less than 8%, but there is more variation when the percentage is greater than 10%.

Plot 13 shows `pct_upreg_nouse` versus `none_standard`
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women not using no method
p13 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p13)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png")
ggsave(filename = figure_file, plot=p13) 
```

P13: The same expected positive trend is observed for women who are not using a birth control method, but there is considerable variation in this graph across all percentages of unintended pregnancies. Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.

We are also interested in how these relationships between percent of unintended pregnancies using each birth control method and the number of people using each method varies across region. We adjusted the original scatter plots created above by coloring based on region.
```{r}
# explore percent unintended pregnancies using long acting reversible and sterilziation methods by region
p14 <- mydata %>% ggplot(aes(x=pct_upreg_larcster, y=larcster_standard, color = region)) + 
  geom_point() + geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using LARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p14)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png")
ggsave(filename = figure_file, plot=p14)

# explore percent unintended pregnancies using traditional methods by region
p16 <- mydata %>% ggplot(aes(x=pct_upreg_trad, y=trad_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +  
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using Traditional Methods") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p16)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png")
ggsave(filename = figure_file, plot=p16)

# explore percent unintended pregnancies using no method by region
p17 <- mydata %>% ggplot(aes(x=pct_upreg_nouse, y=none_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') +   
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using No Method") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p17)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png")
ggsave(filename = figure_file, plot=p17)

```

P14: Among women using long acting reversible and sterilization methods, Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.

P16: Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.

P17: This graph is much different than the other three, which is to be expected at this point. Oceania has the highest percentage of unintended pregnancies when women are not using birth control. The slope of the relationship for each region appears to be similar in this graph with the exception of Oceania having a steeper slope. There are considerably more outliers in this graph with African countries

Our next group of plots will be violin plots to explore the usage level of each birth control method, a categorical variable, associated with percentage of unintended pregnancies use for each birth control method, a numerical variable.
```{r}
# explore long acting reversible and sterilization method usage compared to percent of unintended pregnancies
p18 <- mydata %>% ggplot(aes(x = larcster_usage, y = pct_upreg_larcster)) +
  geom_violin() +
  ggtitle("Usage of LARCs and Sterilization vs Percent of Unintended Pregnancies") +
  labs(x = "Long Acting Reversible and Sterilization Usage", y = "Percentage Unintended Pregnancies") 
plot(p18)
figure_file = here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p18)

# explore traditional method usage compared to percent of unintended pregnancies
p20 <- mydata %>% ggplot(aes(x = trad_usage, y = pct_upreg_trad)) +
  geom_violin() + 
  ggtitle("Usage of Traditional Method vs Percent of Unintended Pregnancies") +
  labs(x = "Traditional Method Usage", y = "Percentage Unintended Pregnancies") 
plot(p20)
figure_file = here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p20)

# explore no method usage compared to percent of unintended pregnancies
p21 <- mydata %>% ggplot(aes(x = none_usage, y = pct_upreg_nouse)) +
  geom_violin() +  
  ggtitle("No Birth Control Usage vs Percent of Unintended Pregnancies") +
  labs(x = "No Birth Control Usage", y = "Percentage Unintended Pregnancies") 
plot(p21)
figure_file = here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p21)

```

P18: The distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods. High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.

P20: Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.

P21: While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2024-03-15 products/manuscript/Manuscript.qmd
Updated Predictor Variables table.

@@
 
-The predictor variables will include the total number of people women using each type of birth control standardized by population size, region, percentage of women currently married, percentage of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due to abortion.
+The predictor variables will include:
+1. The total number of people women using each type of birth control standardized by population size 2. Region
+3. Percentage of women currently married
+4. Percentage of women never married
+5. Total rate of maternal deaths
+6. Rate of maternal deaths due to unintended pregnancy 
+7. Rate of maternal deaths due to abortion.
 
@@
 # Load libraries
-library(knitr)
+
 library(kableExtra)
-library(dplyr)
 
-# Descriptive statistics for numeric variables
-pred_summary <- mydata %>%
-  summarise(across(c(wra, pct_currentlymarried,pct_nevermarried,rate_matdeaths, rate_matdeaths_upreg, rate_matdeaths_abs), mean, sd))
+# Selecting variables
+selected_variables <- mydata[c("pct_upreg_sarc", "region", "wra", "pct_currentlymarried", "pct_nevermarried","rate_matdeaths", "rate_matdeaths_upreg", "rate_matdeaths_abs")]
 
-# Descriptive statistics for categorical variable (assuming region is a factor)
-summary_factor <- mydata %>%
-  group_by(region) %>%
-  summarise(n = n())
-
-# Combine into a single dataframe
-summary_pred <- rbind(
-  pred_summary %>% summarise(Variable = "pct_upreg_sarc", ., .names = quote(names(.))),
-  pred_summary %>% select(-pct_upreg_sarc),
-  summary_factor %>% summarise(Variable = "region", n)
-)
+# Converting region from factor to numeric
+selected_variables$region <- as.numeric(as.factor(selected_variables$region))
 
 # Create table 
-kable(summary_pred, caption = "Table Depicting Predictor Variables", booktabs = TRUE) %>%
-  kable_styling(full_width = F, font_size = 10) %>%
-  column_def(widths = c(200, 150, 150))
+kable(selected_variables, caption = "Table Depicting Predictor Variables", booktabs = TRUE) %>%
+  kable_styling(full_width = F, font_size = 10)
 

### 2024-03-15 code/eda-code/eda.qmd
Final Part 3 Updates
I organized the eda file and added some final interpretations. I cleaned the manuscript file to only include relevant information in the correct order and made sure the document renders.

@@
 ```{r}
+#| message: false
 #load necessary packages
@@
 library(gtsummary) # for creating tables
+library(kableExtra) # for creating tables
 ```
@@
 ```
+
+# Summarize the data 
+```{r}
+table(mydata$region)
+table(mydata$subregion)
+summary(mydata$pct_upreg)
+ggplot(mydata, aes(x = country, y = pct_upreg, fill = region)) +
+  geom_col(position = "dodge") +
+  labs(title = "Percentage Unintended Pregnancies by Country",
+       x = "Country",
+       y = "Percentage")
+```
+
+The dataset includes 53 African countries, 36 Asian countries, 11 European Countries, 24 Latin American/Caribbean countries, and 8 Oceanic countries.There are 16 subregions that provide better detail about each country's location.
+The average percentage of unintended pregnancies at the country level is 46.79%.
+with a maximum of 75.65%.
+This variable has 24 missing values, which will need to be cleaned.
+A simple visualization shows that African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
 
@@
 
-We also created a simple table with the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
+We created a simple table with the categorical variables: `region`, `larcster_usage`, `sarc_usage`, and `none_usage`.
 ```{r}
@@
 
+We also wanted a summary table with the predictor information to include in our manuscript.
+```{r}
+# Selecting variables
+selected_variables <- mydata[c("pct_upreg_sarc", "region", "wra", "pct_currentlymarried", "pct_nevermarried","rate_matdeaths", "rate_matdeaths_upreg", "rate_matdeaths_abs")]
+
+# Converting region from factor to numeric
+selected_variables$region <- as.numeric(as.factor(selected_variables$region))
+
+# Asia = 1
+# Africa = 2
+# Europe = 3
+# Latin America & the Caribbean = 4
+# Oceania = 5
+
+# Create table 
+kable(selected_variables, caption = "Table Depicting Predictor Variables", booktabs = TRUE) %>%
+  kable_styling(full_width = F, font_size = 10)
+
+# Save to file 
+summarytable_file <- here("results", "tables", "predictortable.rds")
+saveRDS(predictor_table, file = summarytable_file)
+```
+
 # Data exploration through figures
@@
 We will make histogram plots for the continuous outcomes in our dataset.
+After exploring thepercent of unintended pregnancies related to each type of contraception method, we chose to focus on short acting reversible methods as our single outcome.
+All of the exploratory analysis on the other 3 potential outcomes can be found in the Supplementary-Material.qmd in the products > manuscript > supplement folder.
+The following exploratory analysis and subsequent statistical analysis will include the `pct_upreg_sarc` outcome only.
 
-It is important to consider the type of birth control methods associated with percent of unintended pregnancies as well. We will create a histogram for the `pct_upreg_sarc` variable.
+To begin exploring our outcome, we will create a histogram for the `pct_upreg_sarc` variable.
 
@@
 P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+
+We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`.
+Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers.
+```{r}
+# explore the region predictor
+p22 <- mydata %>% ggplot(aes(x=region)) + 
+  geom_bar() +
+  ggtitle("Countries in Each Region") + 
+  xlab("Region") + ylab("Number of Countries")
+plot(p22)
+figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
+ggsave(filename = figure_file, plot=p22)
+
+# explore the current marriage predictor
+p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Currently Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p23)
+figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
+ggsave(filename = figure_file, plot=p23)
+
+# explore the never married predictor
+p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Never Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(p24)
+figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
+ggsave(filename = figure_file, plot=p24)
+
+# explore the total maternal mortality rate predictor
+p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
+  geom_histogram() +
+  ggtitle("Total Rate of Maternal Deaths") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p25)
+figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
+ggsave(filename = figure_file, plot=p25)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths > 1000)
+
+# explore the maternal mortality rate from unintended pregnancies predictor
+p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p26)
+figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
+ggsave(filename = figure_file, plot=p26)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths_upreg > 6000)
+
+# explore the maternal mortality rate from abortions
+p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Abortions") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(p27)
+figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
+ggsave(filename = figure_file, plot=p27)
+
+# explore outliers in maternal mortality from abortions
+mydata %>% filter(rate_matdeaths_abs > 500)
+
+```
+P22: The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
+about 35 countries.
+Latin America/the Caribbean accounts for about 25 countries.
+Europe and Oceania account for the least number of observations.
+
+P23: The distribution of women currently married is approximately normal with an average around 57% and no major outliers.
+
+P24: The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.
+
+P25: The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers.
+We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.
+
+P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
+Chad and South Sudan also account for the highest rates of maternal mortality in this category.
+
+P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.

### 2024-03-15 code/processing-code/processing-file.qmd
Final Part 3 Updates
I organized the eda file and added some final interpretations. I cleaned the manuscript file to only include relevant information in the correct order and made sure the document renders.

@@
 ```
+There are 132 observations, one row for each low- to middle- income country, of the 399 variable included in the Adding It Up All Women Dataset.
+Some examples of numeric variables included are percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+There are three important character variables that can be used to group the numeric variables: country name, region, and sub region.
 
@@
 
-We will load and examine the codebook first.
+We will load and examine the codebook to learn more about the dataset.
 
@@
 ```
+
 # Identifying variables to be cleaned 

### 2024-03-15 products/manuscript/Manuscript.qmd
Updated numeric scale on Predictor Variable table
Final Part 3 Updates
I organized the eda file and added some final interpretations. I cleaned the manuscript file to only include relevant information in the correct order and made sure the document renders.

--- 2024-03-14 products/manuscript/Manuscript.qmd
@@
 ```
-## Project Part 1
-
-# Data Source 
-The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world.
-The data used here is from their Adding It Up project.
-The goals of this project include estimating the need for, impact of, and costs associated with providing sexual and reproductive health services.
-Low and middle income countries are the target audience for this research, specifically women of reproductive age (15-49).
-Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted.
-Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others.
-The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.
-
-# Load and describe the data 
-```{r}
-library(here)
-AIUdata <- read_csv(here("data", "raw-data", "AIU All Women Dataset.csv"))
-dim(AIUdata)
-str(AIUdata)
-```
-There are 132 observations, one row for each low- to middle- income country, of the 399 variable included in the Adding It Up All Women Dataset.
-Some examples of numeric variables included are percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
-There are three important character variables that can be used to group the numeric variables: country name, region, and sub region.
-
-# Summarize the data 
-```{r}
-table(AIUdata$region)
-table(AIUdata$subregion)
-summary(AIUdata$pct_upreg)
-ggplot(AIUdata, aes(x = country, y = pct_upreg, fill = region)) +
-  geom_col(position = "dodge") +
-  labs(title = "Percentage Unintended Pregnancies by Country",
-       x = "Country",
-       y = "Percentage")
-```
-The dataset includes 53 African countries, 36 Asian countries, 11 European Countries, 24 Latin American/Caribbean countries, and 8 Oceanic countries.There are 16 subregions that provide better detail about each country's location.
-The average percentage of unintended pregnancies at the country level is 46.79%.
-with a maximum of 75.65%.
-This variable has 24 missing values, which will need to be cleaned.
-A simple visualization shows that African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
-
-# Research Questions
-1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?
-
-The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).
-
-The predictor variables will include the total number of people women using each type of birth control                standardized by population size, region, percentage of women currently married, percentage of women never married,    total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due    to abortion.
-
-A full list of variables can be found in the processing-file in the code > processing-code folder.
-
-# How to analyze 
-Most of the predictors used in this analysis are numeric, so we can build linear regression models to analyze how each predictor affects the percentage of unintended pregnancies among women using short acting reversible methods.
-We can group countries together based on the subregion variable to compare how these relationships vary across the world as shown in the crude visualization above.
-There are other interesting variables that can be used for grouping for analysis such as wealth, urban/rural location, and marriage status.
 
@@
 
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of each contraceptive method.
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,  short acting reversible, and traditional.
 Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies.
@@
 Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study.
-The table provided above under `Summarize the data` shows that African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
+African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
 This could be due to the current legal status of contraception within Latin American countries.
@@
 _Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
+The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world.
+The data used here is from their Adding It Up project.
+The goals of this project include estimating the need for, impact of, and costs associated with providing sexual and reproductive health services.
+Low and middle income countries are the target audience for this research, specifically women of reproductive age (15-49).
+Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted.
+Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others.
+The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.
 
-This data originates from the Guttmacher Institute's Adding It Up project.
-The data contains numeric variables: percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+For the purpose of this project, we chose several numeric variables from the data source: percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
 There are only three character variables that can be used to group the numeric variables: country name, region, and sub region.
-So far we have dvided the observations into "low", "medium", and "high" category of birth control usage.
 
@@
 This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods as our single outcome because this method is becoming increasingly popular, so we feel that results from this contraception method will be the most relevant.
+Additionally, the exploratory data analysis process showed that the distribution of this outcome variable is less skewed than the other options, and the stratification by region is clearer, which allows for better understanding.
 
@@
 
+This research aims to analyze the factors influencing unintended pregnancies and contraception use in low-and-middle-income countries.
+The Guttmacher Institute's Adding It Up project provides the foundation, offering data on sexual and reproductive health services in these regions.Focusing on women of reproductive age (15-49), the study usese information from nationally representative surveys and population projections to estimate unintended pregnancy rates.
+
+While data on the overall percentage of unintended pregnancies was unavailable, we strategically utilized data on rates related to specific birth control methods.
+This approach allows us to analyze the relationship between these rates and various social and health factors.
+Factors like the number of women utilizing different birth control methods, maternal death rates, and marital status will be meticulously examined.
+All 7 of these predictor variables were found within the study's codebook, which can can be found in the `Data` folder of this project.
+By discovering these relationships, the research aims to shed light on the complex landscape of contraception use and unintended pregnancy in low-and -middle-income countries.
+This knowledge can empower policymakers and healthcare providers to make informed decisions that ultimately improve reproductive health outcomes for women.
+
 ## Data import and cleaning
@@
 
-## Exploratory/Descriptive analysis (Project Part 2)
-The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
-We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
-We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
-We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
+## Exploratory/Descriptive analysis
+
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @tbl-summarytable.
 ```{r}
@@
 #| echo: FALSE
-resulttable<- readRDS("../../results/tables/summarytable.rds")
-kable(resulttable)
-```
-
-We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable2.
-We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
-Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
-It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
-```{r}
-#| label: tbl-summarytable2
-#| tbl-cap: "Summary table for categorical variables"
-#| echo: FALSE
-resulttable2 <- readRDS("../../results/tables/summarytablecat.rds")
-kable(resulttable2)
+predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
+predictortable
 ```
@@
 We created a multitude of exploratory plots to understand our data better.
-The histograms of percent of unintended pregnancies related to each method of birth control showed discrepancies in skewness which can be seen @exploratory-result.
-The `pct_upreg_larcster` and `pct_upreg_trad` variables are highly skewed right, which shows that more countries experience lower rates of unintended pregnancies when using these methods.
-The `pct_upreg_sarc` histogram is not as skewed.
-It shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
-The histogram for `pct_upreg_none` shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
+The first @exploratory-result shows the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
+The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
 ```{r}
@@
 #| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_larcster_distribution.png"))
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_trad_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_nouse_distribution.png"))
 ```
@@
 After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years).
-There is a positive association between the percentage of unintended pregnancies and the total number of women using that method of birth control for all three methods and those not using any method, which is to be expected.
-There is more variation among those using SARMs and traditional methods compared to those using long-acting reversible methods.
-Additionally, the lowest percentage in this category is around 35%, which is much higher than the percentages associated with the three birth control methods.
-These figures can be seen @exploratory-result2.
+There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected.
+This association is displayed in @exploratory-result2.
 ```{r}
 #| label: exploratory-result2
-#| fig-cap: "Relationship between Percentage of Unintended Pregnancies and Standardized Population Totals"
+#| fig-cap: "Relationship between Percentage of Unintended Pregnancies and Standardized Population Total"
 #| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctuppreg_totallarcster.png"))
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totaltrad.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalnone.png"))
 ```
 
-We stratified these four graphs by region to examine how the relationships differ across each of the 6 regions which is shown in @exploratory-result3 .
-The long acting reversible and sterilization methods graph is a little difficult to decipher, but Asia has the highest percentages of unintended pregnancies, followed by Latin America/the Caribbean and Africa.
-The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.Among women using traditional birth control methods, Europe has the sharpest increase in unintended pregnancies as population increases, but Asia has the highest percentage of unintended pregnancies in this category.
-The graph for women not using any birth control methods shows that Oceania has the highest percentage of unintended pregnancies.
+We stratified these four graphs by region to examine how the relationship differs across each of the t regions which is shown in @exploratory-result3 .
+The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
 ```{r}
@@
 #| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_larcster_stratified.png"))
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_trad_stratified.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_none_stratified.png"))
 ```
 
-The last set of plots we created to explore the data are violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4.
-The first violin plot shows that the distribution of percentage of unintended pregnancies varies a lot among usage levels of long acting reversible and sterilization methods.
-High usage countries have a consistent percentage of unintended pregnancies, but countries with a medium usage level tend to have lower percentages of unintended pregnancies.
+We also created violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4.
 It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
-Among women using traditional birth control methods, the distribution of percentage of unintended pregnancies is similar among medium and high usage countries, but low usage countries are skewed towards lower percentage of unintended pregnancies.
-While the distribution of percentage of unintended pregnancies is similar across the three usage levels for women not using any birth control method, the high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
+The high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
 The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
@@
 #| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "larcster_pctupreg_violin.png"))
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "trad_pctupreg_violin.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "none_pctupreg_violin.png"))
+```
+
+Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @exploratory-result5.
+The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
+about 35 countries.
+Latin America/the Caribbean accounts for about 25 countries.
+Europe and Oceania account for the least number of observations.
+The distribution of women currently married is approximately normal with an average around 57% and no major outliers.
+The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.
+The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers.
+We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.
+The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
+Chad and South Sudan also account for the highest rates of maternal mortality in this category.
+The distribution of maternal mortality rate from abortions is also highly skewed to the right.
+The outliers are accounted for by Gambia and Mauritania in this predictor.
+```{r}
+#| label: exploratory-result5
+#| fig-cap: "Predictor Variable Distributions"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
+knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))
 ```
@@
 
-We completed a simple linear regression analysis with each of the 7 predictors.
+We completed a simple linear regression analysis with each of the 7 predictors shown in @statistical-analysis1.
 The predictor with the strongest relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
@@
 ```{r}
-#| label: basic-statistical-analysis
+#| label: statistical-analysis1
 #| fig-cap: "Strongest Predictor: sarc_standard"
@@
 ```
-The model using the region predictor revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region.
+
+The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region.
 The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
+These associations can be seen in @statistical-analysis2.
 ```{r}
-#| label: basic-statistical-analysis
+#| label: statistical-analysis2
 #| fig-cap: "Regional Predictor"

### 2024-03-15 products/manuscript/supplement/Supplementary-Material.qmd
Final Part 3 Updates
I organized the eda file and added some final interpretations. I cleaned the manuscript file to only include relevant information in the correct order and made sure the document renders.

@@
 
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
 For illustrative purposes, I'm doing the supplement as pdf.
@@
 
-Of course you would choose the format based on needs.
+# Overview
+This file provides extra information that is not included in our manuscript or the main processing and exploratory data analysis files.
+This information provides guidance on our thought processes behind the scenes and explains why we chose a single outcome to study.
 
-I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
+# Code and file information
 
-# Overview
+Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
+Essentially, give a full set of instructions to re-generate everything.
 
-Readers will be able to find our analysis of the different contraception methods through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
+{{< pagebreak >}}
+
+# Additional Method Details
+
+Often, the main manuscript only allows for an overview description of the methods.
+Use the supplement to describe all your methods, models and approaches in a lot of detail.
+Reference specific parts of your code as needed.
+
+{{< pagebreak >}}
+
+# Additional results
+
+Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
+
+Readers will be able to find our analysis of the different contraception methods that we did not select for our main outcome through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
 
@@
 ```
+
 Plot 10 shows `pct_upreg_larcster` vs `larcster_standard` to analyze women using long acting reversible methods and sterilization versus the percentage of unintended pregnancies using those methods.
@@
 
-# Code and file information
+## Additional result
+The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
+We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
+We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
+We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
+```{r}
+#| label: tbl-summarytable1
+#| tbl-cap: "Summary table to discover missingness"
+#| echo: FALSE
+resulttable1 <- readRDS("../../results/tables/summarytable.rds")
+kable(resulttable1)
+```
 
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
-
-{{< pagebreak >}}
-
-# Additional Method Details
-
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
-
-{{< pagebreak >}}
-
-# Additional results
-
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
-
-## Example additional result
+We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable2.
+We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
+Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
+It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
+```{r}
+#| label: tbl-summarytable2
+#| tbl-cap: "Summary table for categorical variables"
+#| echo: FALSE
+resulttable2 <- readRDS("../../results/tables/summarytablecat.rds")
+kable(resulttable2)
+```
 

### 2024-03-25 code/analysis-code/statistical-analysis.qmd
Module 11 Discussion
I made some updates based on the summary we have to write about our project.

@@
 #| message: false
+library(dplyr) # for basic syntax
 library(broom) #for cleaning up output from lm()
@@
 
-# create a data frame with the predictions and true values of Y
+# create a data frame with the predictions and true values 
 popdata <- bind_cols(poppred, mydata$sarc_standard) %>%
@@
 rmse1 <- rmse(popdata, truth = sarc_standard, estimate = .pred)
+rsq1 <- rsq(popdata, truth = sarc_standard, estimate = .pred)
 
-# create predictions using the pct_currentlymarriedd predictor model
+# create predictions using the pct_currentlymarried predictor model
 marriedpred <- predict(lmfit3, new_data = mydata) %>% 
@@
 
-# create a data frame with the predictions and true values of Y
+# create a data frame with the predictions and true values 
 marrieddata <- bind_cols(marriedpred, mydata$pct_currentlymarried) %>%
@@
 rmse2 <- rmse(marrieddata, truth = pct_currentlymarried, estimate = .pred)
+rsq2 <- rsq(marrieddata, truth = pct_currentlymarried, estimate = .pred)
+
+# create predictions using the rate of maternal deaths predictor model
+matdeathspred <- predict(lmfit5, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+matdeathdata <- bind_cols(matdeathspred, mydata$rate_matdeaths) %>%
+            rename(rate_matdeaths = "...2")
+
+# find RMSE and R squared to determine model fit
+rmse3 <- rmse(matdeathdata, truth = rate_matdeaths, estimate = .pred)
+rsq3 <- rsq(matdeathdata, truth = rate_matdeaths, estimate = .pred)
 
@@
 print(rmse1)
+print(rsq1)
 print(rmse2)
+print(rsq2)
+print(rmse3)
+print(rsq3)
 ```

### 2024-03-27 code/analysis-code/statistical-analysis.qmd
Part 4 Updates
I added ANOVA and Tukey HSD tests to explore relationships between regions. I worked on the model performance section by finding RMSE values for the most important predictors and the MLR model. I also added a null model to use for comparison in CV analysis next.

--- 2024-03-14 code/analysis-code/statistical-analysis.qmd
@@
 ---
-title: "Statistical Analysis Script"
+title: "Statistical Analysis"
 author: "Taylor Glass and Arlyn Santiago"
@@
 library(tidymodels) # for model building 
+library(yardstick) # for finding performance metrics 
+library(rsample) #for splitting the data into testing/training
 ```
@@
 ### First Model Fit 
-The second model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception.
+The first model attempts to predict the percentage of unintended pregnancies among women using short acting reversible methods of contraception 'pct_upreg_sarc' based on the total number of women using this method of contraception.
 We will use the version of the variable that has been standardized by population size, 'sarc_standard'.
@@
 saveRDS(lmtable2, file = table_file2)
+
 ```
-This model shows that the Latin America/the Carribean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region.
+This model shows that the Latin America/the Caribbean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region.
 The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.
+
+### ANOVA for region variable
+We want to compare the mean percent of unintended pregnancies among those using short acting reversible methods across the five regions: Asia, Africa, Europe, Latin America and the Caribbean, and Oceania.
+We will start with an ANOVA test to determine the relationship between region and outcome.
+We are more interested in pairwise comparisons between the regions, so we also conducted Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA.
+```{r}
+# conduct an ANOVA test with the region variable
+regionANOVA <- aov(pct_upreg_sarc ~ region, data = mydata)
+
+# explore ANOVA results 
+summary(regionANOVA)
+
+# add Tukey's HSD test for pairwise comparisons
+tukey_result <- TukeyHSD(regionANOVA)
+
+# explore the pairwise comparisons
+tidy(tukey_result)
+```
+ANOVA results show that the region variable is statistically significantly associated with the outcome variable given the p-value less than 0.001.
+The Tukey HSD test revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia.
+The relationships between Oceania and Europe and Latin America/the Caribbean are statistically significant with p-values less than 0.0001.
+Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.
+The lowest difference in outcome between regions is between Asia and Africa, but this is not a statistically significant assocaition.
+Two statistically significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.
 
@@
 
+### Nineth model fit 
+It is important to create the null model for future assessment of model performance, which just calculates the mean outcome without any predictors present. 
+```{r}
+# create a null model
+null_model <- null_model(mode = "regression") %>% 
+    set_engine("parsnip") %>%
+    fit(pct_upreg_sarc ~ 1, data = mydata)
+
+#generate clean output with estimates and p-values
+tidy(null_model)
+
+# save results from fit into a data frame 
+nulltable <- tidy(null_model)
+
+# save fit results table  
+table_file9 = here("results", "tables", "resulttable9.rds")
+saveRDS(nulltable, file = table_file9)
+```
+
 ## Assessing Model Performance
 It is important to consider appropriate metrics to measure model performance.
-We will use RMSE at this point to compare the model using the number of women using short acting reversible contraception predictor and the model using the percent currently married predictor.
+We will use RMSE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor.
+We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above.
 ```{r}
+# set seed for reproducibility 
+set.seed(2468)
+
 # create predictions using the sarc_standard predictor model
@@
 
-# find RMSE and R squared to determine model fit
+# find RMSE to determine model fit
 rmse1 <- rmse(popdata, truth = sarc_standard, estimate = .pred)
 
-# create predictions using the pct_currentlymarriedd predictor model
-marriedpred <- predict(lmfit3, new_data = mydata) %>% 
+# create predictions using the pct_nevermarried predictor model
+marriedpred <- predict(lmfit4, new_data = mydata) %>% 
                   select(.pred)
 
-# create a data frame with the predictions and true values of Y
-marrieddata <- bind_cols(marriedpred, mydata$pct_currentlymarried) %>%
-            rename(pct_currentlymarried = "...2")
+# create a data frame with the predictions and true values 
+marrieddata <- bind_cols(marriedpred, mydata$pct_nevermarried) %>%
+            rename(pct_nevermarried = "...2")
 
-# find RMSE and R squared to determine model fit
-rmse2 <- rmse(marrieddata, truth = pct_currentlymarried, estimate = .pred)
+# find RMSE to determine model fit
+rmse2 <- rmse(marrieddata, truth = pct_nevermarried, estimate = .pred)
+
+# create predictions using the rate of maternal deaths due to abortion predictor model
+matdeathspred <- predict(lmfit7, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+matdeathdata <- bind_cols(matdeathspred, mydata$rate_matdeaths_abs) %>%
+            rename(rate_matdeaths_abs = "...2")
+
+# find RMSE to determine model fit
+rmse3 <- rmse(matdeathdata, truth = rate_matdeaths_abs, estimate = .pred)
 
@@
 print(rmse2)
+print(rmse3)
 ```
 
-The RMSE metric favors the model with the total number of women using the short acting reversible contraception method over the model with the percentage of women currently married, which is to be expected.
+The RMSE metric favors the model with the percentage of women never married because it has the lowest RMSE value of 17.01493.
+The model with the total number of women using short acting reversible methods is slightly worse with an RMSE 18.22695.
+The model with the rate of maternal deaths due to abortion performs much worse, with a high RMSE value of 129.482.
+
+It is important to determine if our model with the region predictor performs well because the main question to be answered by this analysis depends on this predictor.
+Since region is a categorical variable, we cannot evaluate the RMSE performance metric.
+```{r}
+# set seed for reproducibility 
+set.seed(2468)
+
+# create predictions using the sarc_standard predictor model
+regionpred <- predict(lmfit2, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+regiondata <- bind_cols(regionpred, mydata$region) %>%
+            rename(region = "...2")
+
+# find RMSE to determine model fit
+rmse4 <- rmse(regiondata, truth = region, estimate = .pred)
+```
+
+We want to see how the model with all 7 predictors performs in comparison to the single predictor models. 
+```{r}
+# set seed for reproducibility 
+set.seed(2468)
+
+# create predictions using the multiple linear regression model
+allpred <- predict(lmfit8, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+alldata <- bind_cols(allpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find RMSE to determine model fit
+rmse5 <- rmse(alldata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse5)
+```
+As expected, this model performs better than the single predictor models with a low RMSE score of 4.209.
+
+## Cross-Validation to Assess Model Performance
+While the RMSE performance is a good place to start for evaluating model performance, we need to determine how well the model performs on data it has not seen before to determine its usefulness.
+We will use the cross-validation process to do this.
+The first step is to split the data into testing and training portions.
+```{r}
+# set seed for reproducibility
+set.seed(2468)
+
+# put 75% into the training data
+datasplit <- initial_split(mydata, prop = 3/4)
+
+# create data frames for the two sets
+train_data <- training(datasplit)
+test_data <- testing(datasplit)
+```

### 2024-03-28 code/eda-code/eda.qmd
Updated the predictor variable rds file error
Updated the summary table for our predictor variables

@@
 
-# Converting region from factor to numeric
-selected_variables$region <- as.numeric(as.factor(selected_variables$region))
-
-# Asia = 1
-# Africa = 2
-# Europe = 3
-# Latin America & the Caribbean = 4
-# Oceania = 5
+selected_variables <- selected_variables %>% 
+  arrange(region)
 
@@
 summarytable_file <- here("results", "tables", "predictortable.rds")
-saveRDS(predictor_table, file = summarytable_file)
+saveRDS(predictortable, file = summarytable_file)
 ```

### 2024-03-28 code/analysis-code/statistical-analysis.qmd
Statistical Analysis Updates
I added another model with a combination of predictors and found more RMSE values to compare models. I found correlations between predictors and adjusted the multivariate model accordingly.

@@
 Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.
-The lowest difference in outcome between regions is between Asia and Africa, but this is not a statistically significant assocaition.
+The lowest difference in outcome between regions is between Asia and Africa, but this is not a statistically significant association.
 Two statistically significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.
@@
 
+### Assessing correlation between predictors
+There are some variables that are intuitively correlated, which means they should not be included in the same model.
+We want to explore those correlations before building our multivariate linear regression model.
+```{r}
+# find correlation coefficients 
+cor(mydata$pct_currentlymarried, mydata$pct_nevermarried)
+cor(mydata$rate_matdeaths, mydata$rate_matdeaths_upreg)
+cor(mydata$rate_matdeaths, mydata$rate_matdeaths_abs)
+cor(mydata$rate_matdeaths_upreg, mydata$rate_matdeaths_abs)
+```
+The percent of currently married women and the percent of never married women have a correlation coefficient of -0.8588, so these two variables are highly correlated.
+The total rate of maternal deaths and rate of maternal deaths due to unintended pregnancies is the most highly correlated pair with a coefficient of 0.8904.
+The correlation between total rate of maternal deaths and rate of maternal deaths due to abortion is slightly lower with a coefficient of 0.8023.
+The two variables with the least correlation out of these four pairs are rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion with a coefficient of 0.6486.
+Based on these associations, we will only include the percent of currently married women in the multivariate model because it has a slightly larger effect on the outcome with a more significant p-value in its single predictor model compared to the percent of never married women predictor.
+We will include both the rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion beccause the total rate of maternal deaths is so highly correlated with each of these variables.
+
 ### Eighth Model Fit 
-We will explore all 7 predictors using in a multiple linear regression model to determine how the associations  interact. 
+Based on the exploration of correlation between predictors, we include 5 predictors in the multiple linear regression model to determine how the associations interact. 
 ```{r}
@@
 lmfit8 <- lm_mod %>% #estimate/train the model 
-            fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + pct_nevermarried + rate_matdeaths + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)
+            fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)
 
@@
 ```
+We also want to observe models with different combinations of predictors to determine if there is any evidence of interaction or effect measure modification.
+Multivariate analysis allows for the assessment of the association between each predictor and the outcome while adjusting for the effects of other variables.
+
+### Tenth Model fit 
+We will create a model with all 3 rate of maternal mortality predictors to compare to the three models with each predictor on its own to further explore the correlation between these variables. 
+```{r}
+# fit linear model using pct_upreg_sarc as outcome with rate of maternal mortality predictors
+lm_mod <- linear_reg() #specify the type of model
+lmfit10 <- lm_mod %>% #estimate/train the model 
+            fit(pct_upreg_sarc ~ rate_matdeaths + rate_matdeaths_upreg + rate_matdeaths_abs, mydata)
+
+#generate clean output with estimates and p-values
+tidy(lmfit10)
+
+# save results from fit into a data frame 
+lmtable10 <- tidy(lmfit10)
+
+# save fit results table  
+table_file10 = here("results", "tables", "resulttable10.rds")
+saveRDS(lmtable10, file = table_file10)
+```
+It appears that there is an association between these three variables because the direction of the association between rate of maternal deaths and outcome has flipped.
+This model shows that rate of maternal deaths is associated with an increase in the outcome by a factor of 0.011267, while the single predictor model showed a slightly negatively correlation with the outcome based on a small coefficient estimate of -0.01482568.
+The other two variables retain their negative association with the outcome variable with similar coefficient measurements.
 
@@
 # create a data frame with the predictions and true values 
-popdata <- bind_cols(poppred, mydata$sarc_standard) %>%
-            rename(sarc_standard = "...2")
+popdata <- bind_cols(poppred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
 
 # find RMSE to determine model fit
-rmse1 <- rmse(popdata, truth = sarc_standard, estimate = .pred)
+rmse1 <- rmse(popdata, truth = pct_upreg_sarc, estimate = .pred)
 
-# create predictions using the pct_nevermarried predictor model
-marriedpred <- predict(lmfit4, new_data = mydata) %>% 
+# create predictions using the pct_curmarried predictor model
+marriedpred <- predict(lmfit3, new_data = mydata) %>% 
                   select(.pred)
@@
 # create a data frame with the predictions and true values 
-marrieddata <- bind_cols(marriedpred, mydata$pct_nevermarried) %>%
-            rename(pct_nevermarried = "...2")
+marrieddata <- bind_cols(marriedpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
 
 # find RMSE to determine model fit
-rmse2 <- rmse(marrieddata, truth = pct_nevermarried, estimate = .pred)
+rmse2 <- rmse(marrieddata, truth = pct_upreg_sarc, estimate = .pred)
 
 # create predictions using the rate of maternal deaths due to abortion predictor model
-matdeathspred <- predict(lmfit7, new_data = mydata) %>% 
+matdeaths_abspred <- predict(lmfit7, new_data = mydata) %>% 
                   select(.pred)
@@
 # create a data frame with the predictions and true values 
-matdeathdata <- bind_cols(matdeathspred, mydata$rate_matdeaths_abs) %>%
-            rename(rate_matdeaths_abs = "...2")
+matdeath_absdata <- bind_cols(matdeaths_abspred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
 
 # find RMSE to determine model fit
-rmse3 <- rmse(matdeathdata, truth = rate_matdeaths_abs, estimate = .pred)
+rmse3 <- rmse(matdeath_absdata, truth = pct_upreg_sarc, estimate = .pred)
+
+# create predictions using the null model
+nullpred <- predict(lmfit9, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+nulldata <- bind_cols(nullpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find RMSE to determine model fit
+rmse4 <- rmse(nulldata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 print(rmse3)
+print(rmse4)
 ```
 
-The RMSE metric favors the model with the percentage of women never married because it has the lowest RMSE value of 17.01493.
-The model with the total number of women using short acting reversible methods is slightly worse with an RMSE 18.22695.
-The model with the rate of maternal deaths due to abortion performs much worse, with a high RMSE value of 129.482.
+The RMSE metric favors the model with the total number of women using short acting reversible contraception methods based on a low RMSE value of 5.143319.
+The model with the percentage of women currently married performs worse with an RMSE value of 10.17591.
+The same conclusion can be drawn for the model with the rate of maternal mortality due to abortions predictor only, given the RMSE value of 10.14128.
+The null model has an RMSE value of 9.842267, so all models should have an RMSE lower than this to be considered useful.
 
@@
 # create a data frame with the predictions and true values 
-regiondata <- bind_cols(regionpred, mydata$region) %>%
-            rename(region = "...2")
+regiondata <- bind_cols(regionpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
 
 # find RMSE to determine model fit
-rmse4 <- rmse(regiondata, truth = region, estimate = .pred)
+rmse5 <- rmse(regiondata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse5)
 ```
 
-We want to see how the model with all 7 predictors performs in comparison to the single predictor models. 
+The results of this RMSE value are disappointing because 18.11857 is much larger than the null model's RMSE of 9.842267.
+However, this issue could possibly be explained by the fact that Africa has the most significant differences in our outcome compared to the other regions.
+We will still include this predictor in our multivariate regression model based on the significant Tukey conclusions and the large differences in outcome displayed in our exploratory data analysis results.
+
+After examining the relationship between the rate of maternal mortality variables, we want to determine which model performs the best.
+We have already found the RMSE value for the rate of maternal deaths due to abortion, which is stored as `rmse3`.
+We need to find the RMSE values for the other two single predictor models and the model with all 3 predictors.
+```{r}
+# create predictions using the rate of maternal deaths predictor model
+matdeathsallpred <- predict(lmfit5, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+matdeathalldata <- bind_cols(matdeathsallpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find RMSE to determine model fit
+rmse6 <- rmse(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse6)
+
+# create predictions using the rate of maternal deaths from unintended pregnancies predictor model
+matdeaths_upregpred <- predict(lmfit6, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+matdeath_upregdata <- bind_cols(matdeaths_upregpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find RMSE to determine model fit
+rmse7 <- rmse(matdeath_upregdata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse7)
+
+# create predictions using model with all three rate predictors
+totalratepred <- predict(lmfit10, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+totalratedata <- bind_cols(totalratepred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find RMSE to determine model fit
+rmse8 <- rmse(totalratedata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse8)
+
+# display RMSE values for each model
+print(rmse6) #rate of maternal mortality predictor only
+print(rmse7) #rate of maternal mortality due to unintended pregnancy predictor only
+print(rmse3) #rate of maternal mortality due to abortion predictor only
+print(rmse8) #all 3 rates of maternal mortality predictors together 
+```
+The RMSE values shows that the three single predictor models perform worse than the null model, which is not ideal.
+The model with all 3 maternal morality rate predictors performs exactly the same as the null model, which makes sense because the coefficients estimating the effect of each predictor on the outcome are so small.
+
+We want to see how the model with all 5 predictors performs in comparison to the single predictor models. 
 ```{r}
@@
 # find RMSE to determine model fit
-rmse5 <- rmse(alldata, truth = pct_upreg_sarc, estimate = .pred)
-print(rmse5)
+rmse9 <- rmse(alldata, truth = pct_upreg_sarc, estimate = .pred)
+print(rmse9)
 ```
-As expected, this model performs better than the single predictor models with a low RMSE score of 4.209.
+As expected, this model performs better than the single predictor models with the lowest RMSE score of 4.209.
 
@@
 ```
+
+We will only cross-validate the multivariate regression model `lmfit8` because it performs the best in terms of RMSE. 
+```{r}
+
+```

### 2024-04-01 code/analysis-code/statistical-analysis.qmd
Statistical Analysis Updates
I added MAE as another performance metric and completed cross-validation for the two best models. I graphed the testing and training predictions using the MLR model. I updated the manuscript methods and results for part 4.

@@
 library(rsample) #for splitting the data into testing/training
+library(knitr) # to create performance metric tables 
+library(parsnip) # to create a workflow for cross-validation
 ```
@@
 tidy(tukey_result)
+
+# save fit results table  
+table_file13 = here("results", "tables", "resulttable13.rds")
+saveRDS(tukey_result, file = table_file13)
 ```
@@
 Based on these associations, we will only include the percent of currently married women in the multivariate model because it has a slightly larger effect on the outcome with a more significant p-value in its single predictor model compared to the percent of never married women predictor.
-We will include both the rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion beccause the total rate of maternal deaths is so highly correlated with each of these variables.
+We will include both the rate of maternal deaths due to unintended pregnancies and rate of maternal deaths due to abortion because the total rate of maternal deaths is so highly correlated with each of these variables.
 
@@
 It is important to consider appropriate metrics to measure model performance.
-We will use RMSE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor.
+We will use RMSE and MAE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor.
 We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above.
+We chose these performance metrics because they are commonly used for linear regression models like the ones we have created.
+RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values.
+We include both metrics to increase the validity of our results.
 ```{r}
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse1 <- rmse(popdata, truth = pct_upreg_sarc, estimate = .pred)
+mae1 <- mae(popdata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse2 <- rmse(marrieddata, truth = pct_upreg_sarc, estimate = .pred)
+mae2 <- mae(marrieddata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse3 <- rmse(matdeath_absdata, truth = pct_upreg_sarc, estimate = .pred)
+mae3 <- mae(matdeath_absdata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse4 <- rmse(nulldata, truth = pct_upreg_sarc, estimate = .pred)
+mae4 <- mae(matdeath_absdata, truth = pct_upreg_sarc, estimate = .pred)
 
-# compare the metrics 
-print(rmse1)
-print(rmse2)
-print(rmse3)
-print(rmse4)
+# create a matrix with RMSE and MAE values
+results <- as.table(matrix(c(rmse1$.estimate, mae1$.estimate, rmse2$.estimate, mae2$.estimate, rmse3$.estimate, mae3$.estimate,
+                             rmse4$.estimate, mae4$.estimate), nrow = 4, ncol = 2, byrow = TRUE))
+
+# add column names
+colnames(results) <- c("RMSE", "MAE")
+
+# add row names
+rownames(results) <- c("Model 1-Population Size Using SARCs", "Model 2-Currently Married", 
+                             "Model 3-Rate of Maternal Deaths(Abortion)", "Model 4-Null")
+
+# print the table
+kable(results)
+
+# save performance metrics results table  
+table_file11 = here("results", "tables", "resulttable11.rds")
+saveRDS(results, file = table_file11)
 ```
 
-The RMSE metric favors the model with the total number of women using short acting reversible contraception methods based on a low RMSE value of 5.143319.
-The model with the percentage of women currently married performs worse with an RMSE value of 10.17591.
-The same conclusion can be drawn for the model with the rate of maternal mortality due to abortions predictor only, given the RMSE value of 10.14128.
-The null model has an RMSE value of 9.842267, so all models should have an RMSE lower than this to be considered useful.
+Both metrics favor the model with the total number of women using short acting reversible contraception methods based on the lowest RMSE value of 5.143319 and lowest MAE value of 3.953767.
+The model with the percentage of women currently married performs worse with an RMSE value of 10.17591 and MAE value of 8.107884.
+The same conclusion can be drawn for the model with the rate of maternal mortality due to abortions predictor only, given the RMSE value of 10.14128 and MAE value of 8.21687.
+The null model has an RMSE value of 9.842267 and MAE of 8.21687, so all models should have performance metrics lower than these values to be considered useful.
 
 It is important to determine if our model with the region predictor performs well because the main question to be answered by this analysis depends on this predictor. 
-Since region is a categorical variable, we cannot evaluate the RMSE performance metric.
 ```{r}
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse5 <- rmse(regiondata, truth = pct_upreg_sarc, estimate = .pred)
+mae5 <- mae(regiondata, truth = pct_upreg_sarc, estimate = .pred)
 print(rmse5)
+print(mae5)
 ```
 
-The results of this RMSE value are disappointing because 18.11857 is much larger than the null model's RMSE of 9.842267.
+The results of these performance metrics are disappointing because 18.11857 is much larger than the null model's RMSE of 9.842267 and 14.66513 is also larger than the null model's MAE.
 However, this issue could possibly be explained by the fact that Africa has the most significant differences in our outcome compared to the other regions.
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse6 <- rmse(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
-print(rmse6)
+mae6 <- mae(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse7 <- rmse(matdeath_upregdata, truth = pct_upreg_sarc, estimate = .pred)
-print(rmse7)
+mae7 <- mae(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse8 <- rmse(totalratedata, truth = pct_upreg_sarc, estimate = .pred)
-print(rmse8)
+mae8 <- mae(matdeathalldata, truth = pct_upreg_sarc, estimate = .pred)
 
-# display RMSE values for each model
-print(rmse6) #rate of maternal mortality predictor only
-print(rmse7) #rate of maternal mortality due to unintended pregnancy predictor only
-print(rmse3) #rate of maternal mortality due to abortion predictor only
-print(rmse8) #all 3 rates of maternal mortality predictors together 
+# create a matrix with RMSE and MAE values
+results2 <- as.table(matrix(c(rmse6$.estimate, mae6$.estimate, rmse7$.estimate, mae7$.estimate, rmse3$.estimate, mae3$.estimate,
+                              rmse8$.estimate, mae8$.estimate, rmse4$.estimate, mae4$.estimate), nrow = 5, ncol = 2, byrow = TRUE))
+
+# add column names
+colnames(results2) <- c("RMSE", "MAE")
+
+# add row names
+rownames(results2) <- c("Maternal Morality Model", "Maternal Mortality due to Unintended Pregnancies Model", 
+                        "Maternal Mortality due to Abortion Model", "All 3 Maternal Mortality Predictors Model", "Null Model")
+
+# print the table
+kable(results2)
+
+# save performance metrics results table  
+table_file12 = here("results", "tables", "resulttable12.rds")
+saveRDS(results2, file = table_file12)
+
 ```
-The RMSE values shows that the three single predictor models perform worse than the null model, which is not ideal.
+The performance metrics show that the three single predictor models perform worse than the null model, which is not ideal.
 The model with all 3 maternal morality rate predictors performs exactly the same as the null model, which makes sense because the coefficients estimating the effect of each predictor on the outcome are so small.
+There is more variation in RMSE values compared to MAE values, but the conclusions are the same.
 
@@
 
-# find RMSE to determine model fit
+# find performance metrics to determine model fit
 rmse9 <- rmse(alldata, truth = pct_upreg_sarc, estimate = .pred)
+mae9 <- mae(alldata, truth = pct_upreg_sarc, estimate = .pred)
 print(rmse9)
+print(mae9)
 ```
-As expected, this model performs better than the single predictor models with the lowest RMSE score of 4.209.
+As expected, this model performs better than the single predictor models with the lowest RMSE score of 4.361363 and lowest MAE score of 3.386193.
 
 ## Cross-Validation to Assess Model Performance
-While the RMSE performance is a good place to start for evaluating model performance, we need to determine how well the model performs on data it has not seen before to determine its usefulness.
+While the RMSE and MAE performance metrics are a good place to start for evaluating model performance, we need to determine how well the model performs on data it has not seen before to determine its usefulness.
 We will use the cross-validation process to do this.
@@
 
-We will only cross-validate the multivariate regression model `lmfit8` because it performs the best in terms of RMSE. 
+We will only cross-validate the single predictor model with the highest performance, `lmfit1`, with the total population size using short acting reversible contraception and the  multivariate regression model,`lmfit8`, because it performs the best in terms of RMSE and MAE.
+Our dataset is relatively small with only 132 observations, so we will use 5 folds for the cross-validation process.
 ```{r}
+# set seed for reproducibility
+set.seed(2468)
 
+# create 10 folds
+folds <- vfold_cv(train_data, v=5)
+
+# define the model specification
+lm_spec <- linear_reg() %>% 
+  set_engine("lm")
+
+# initialize a workflow and fit it to 'lmfit1'
+my_workflow <- workflow() %>%
+  add_model(lm_spec) %>% 
+  add_formula(pct_upreg_sarc ~ sarc_standard) %>% #specify the formula since we are not using a recipe
+  fit(data = train_data)
+
+# resample the model 
+my_resample <- my_workflow %>% 
+                fit_resamples(folds)
+
+# initialize a workflow and fit it to 'lmfit8'
+my_workflow2 <- workflow() %>%
+  add_model(lm_spec) %>% 
+  add_formula(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs) %>% #specify the formula since we are not using a recipe
+  fit(data = train_data)
+
+# resample the model 
+my_resample2 <- my_workflow2 %>% 
+                fit_resamples(folds)
+
+# collect the metrics from both models
+collect_metrics(my_resample)
+collect_metrics(my_resample2)
 ```
+The cross-validated results lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
+The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319.
+While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still slightly lower at 4.405716.
+Additionally, the standard error the multiple linear regression model is 0.2768, which is lower than the standard error for the single predictor model of 0.3528.
+This means there is higher variance in the single predictor model.
+The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data.
+Overall, the multiple linear regression model performs better.
+
+Since we have concluded that the model with all 5 predictors performs the best, we will use this model to make predictions for the test data.
+We will create a graph to show how well the model performed on the training and testing data.
+```{r}
+# set seed for reproducibility
+set.seed(2468)
+
+# create predictions using MLR model with the test data
+testpred <- predict(my_workflow2, new_data = test_data) %>% 
+                  select(.pred)
+
+# create predictions using MLR model with training data
+trainpred <- predict(my_workflow2, new_data = train_data) %>% 
+                  select(.pred)
+
+# create data frame with the observed values predicted values from model 2
+plotdata <- data.frame(
+  observed = c(train_data$pct_upreg_sarc), 
+  training_model2 = c(trainpred$.pred), 
+  test_model2 = c(testpred$.pred),
+  model = rep(c("Observed", "Training Model", "Testing Model"), each = nrow(train_data))) # add label indicating the model
+
+# create a visual representation with 2 sets of predicted and observed values 
+ggplot(plotdata, aes(x = observed)) +
+  geom_point(aes(y = training_model2, color = "training model"), shape = 3) +
+  geom_point(aes(y = test_model2, color = "testing model"), shape = 3) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") + #add the 45 degree line
+  xlim(0, 50) + #limit the x-axis values
+  ylim(0, 50) + #limit the y-axis values
+  labs(x = "Observed Values", y = "Predicted Values", title = "Comparing Training and Testing Data on MLR Model")
+```
+The training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
+The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
+While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data.

### 2024-04-01 products/manuscript/Manuscript.qmd
Statistical Analysis Updates
I added another model with a combination of predictors and found more RMSE values to compare models. I found correlations between predictors and adjusted the multivariate model accordingly.
Statistical Analysis Updates
I added MAE as another performance metric and completed cross-validation for the two best models. I graphed the testing and training predictions using the MLR model. I updated the manuscript methods and results for part 4.

@@
 
-1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?
+1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region?
 
@@
 We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion.
-We also assess the performance of the models in this file.
-Result tables for each model are stored in the results > tables folder.
+We also explore the differences in the outcome between regions using an ANOVA test and Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA.
+Result tables for each model and test are stored in the results > tables folder.
+
+Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder.
+We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE).
+We chose these performance metrics because they are commonly used for linear regression models like the ones we have created.
+RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values.
+We include both metrics to increase the validity of our results.
+Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before.
 
@@
 _Explain anything related to your statistical analyses._
-We began with a simple linear regression model with each of the 7 predictor variables.
-We also included a multiple linear regression model with all 7 predictors.
-We included the RMSE for two of the models to begin exploring model fit.
+We began with a simple linear regression model with each of the 7 predictor variables to determine the association betwen each variable and the outcome.
+We explored the region variable further because our main goal is to understand the difference in outcome by region.
+We used ANOVA to determine the difference between regions, and we added Tukey's Honeslty Significant Difference test to examine pairwise associations.
+This led to discovering correlations between some of the predictors, which we examined through correlation coefficients.
+We tried out a combination of these correlated predictors in 2 different models to further examine the interactions.
+After considering which predictors are most impactful on our outcome, we narrowed the final predictors down to a list of 5.
+We included a multiple linear regression model with all 5 of these predictors.
+We used the RMSE and MAE performance metrics to evaluate all 9 models that we created to determine which one fit best.
+Lastly, we used cross-validation to determine how well our multiple linear regression model performs on data it has not seen before.
 
@@
 ```
-We created one model with all 7 predictors using multiple linear regression.
+The Tukey HSD test for the region predictor, shown in @statistical-analysis3, revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia.
+The differences between Oceania and Europe and Latin America/the Caribbean are the largest.
+Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.
+Two more significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.
+```{r}
+#| label: statistical-analysis3
+#| fig-cap: "Tukey HSD Test Results"
+#| echo: FALSE
+readRDS(here("results","tables", "resulttable13.rds")) 
+```
+
+We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4.
 This model shows the strongest predictor is the number of women using short acting reversible contraception.
@@
 The Oceania region now has a negative association with the outcome.
-Both of the marriage status predictors are associated with a decrease in the outcome.
-The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.
+The currently married predictor is associated with a weak decrease in the outcome.
+The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome.
 ```{r}
-#| label: basic-statistical-analysis
-#| fig-cap: "Multiple Linear Regression with 7 Predictors"
+#| label: statistical-analysis4
+#| fig-cap: "Multiple Linear Regression with 5 Predictors"
 #| echo: FALSE

### 2024-04-02 code/analysis-code/statistical-analysis.qmd
Part 4 Updates
Completed bootstrapping and graphed the results. Added LASSO regression and tuned the parameters. Organized all results in the manuscript.

@@
 library(parsnip) # to create a workflow for cross-validation
+library(glmnet) # for LASSO regression
 ```
@@
 # create a matrix with RMSE and MAE values
-results2 <- as.table(matrix(c(rmse6$.estimate, mae6$.estimate, rmse7$.estimate, mae7$.estimate, rmse3$.estimate, mae3$.estimate,
-                              rmse8$.estimate, mae8$.estimate, rmse4$.estimate, mae4$.estimate), nrow = 5, ncol = 2, byrow = TRUE))
+results2 <- as.table(matrix(c(rmse6$.estimate, mae6$.estimate, rmse7$.estimate, mae7$.estimate, rmse3$.estimate, mae3$.estimate, rmse8$.estimate, mae8$.estimate, rmse4$.estimate, mae4$.estimate), nrow = 5, ncol = 2, byrow = TRUE))
 
@@
 As expected, this model performs better than the single predictor models with the lowest RMSE score of 4.361363 and lowest MAE score of 3.386193.
+
+We want to have a table comparing the RMSE and MAE values for the three single predictor models of interest along with the full model with all 5 predictors, which will all be compared to the null model.
+```{r}
+# create a matrix with RMSE and MAE values
+results3 <- as.table(matrix(c(rmse1$.estimate, mae1$.estimate, rmse2$.estimate, mae2$.estimate,
+                             rmse3$.estimate, mae3$.estimate, rmse9$.estimate, mae9$.estimate,
+                             rmse4$.estimate, mae4$.estimate), nrow = 5, ncol = 2, byrow = TRUE))
+
+# add column names
+colnames(results3) <- c("RMSE", "MAE")
+
+# add row names
+rownames(results3) <- c("Population Size Using SARCs", "Percent Currently Married", 
+                       "Rate of Maternal Deaths due to Abortion", "Full Model with 5 Predictors", "Null Model")
+
+# print the table
+kable(results3)
+
+# save performance metrics results table  
+table_file14 = here("results", "tables", "resulttable14.rds")
+saveRDS(results3, file = table_file14)
+```
 
@@
 While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still slightly lower at 4.405716.
-Additionally, the standard error the multiple linear regression model is 0.2768, which is lower than the standard error for the single predictor model of 0.3528.
+Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528.
 This means there is higher variance in the single predictor model.
@@
 Overall, the multiple linear regression model performs better.
+
+We want a table displaying these results as well. 
+```{r}
+# create a matrix with RMSE and MAE values
+results4 <- as.table(matrix(c(4.822226, 0.25279125, 0.786067, 0.04292202, 4.405716, 0.27679502, 0.820600, 0.02815933), nrow = 2, ncol = 4, byrow = TRUE))
+
+# add column names
+colnames(results4) <- c("RMSE", "Standard Error", "R-Squared", "Standard Error")
+
+# add row names
+rownames(results4) <- c("Single Predictor Model", "Full Predictor Model")
+
+# print the table
+kable(results4)
+
+# save performance metrics results table  
+table_file15 = here("results", "tables", "resulttable15.rds")
+saveRDS(results4, file = table_file15)
+```
 
@@
 # create a visual representation with 2 sets of predicted and observed values 
-ggplot(plotdata, aes(x = observed)) +
+figure1 <- ggplot(plotdata, aes(x = observed)) +
               geom_point(aes(y = training_model2, color = "training model"), shape = 3) +
@@
               labs(x = "Observed Values", y = "Predicted Values", title = "Comparing Training and Testing Data on MLR Model")
+figure1
+
+# save final figure
+figure_file = here("results", "figures", "final-figures", "fullmodelpred.png")
+ggsave(filename = figure_file, plot=figure1) 
 ```
@@
 While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data.
+
+To examine the uncertainty in the model further, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. 
+```{r}
+# set the seed for reproducibility
+set.seed(2468)
+
+# create 100 bootstraps with the training data
+bootstraps <- bootstraps(train_data, times = 100)
+
+# create empty vector to store predictions list 
+preds_bs <- vector("list", length = length(bootstraps))
+
+# write a loop to fit the model to each bootstrap and make predictions
+for (i in 1:length(bootstraps)) {
+
+  bootstrap_sample <- analysis(bootstraps$splits[[i]])  # isolate the bootstrap sample
+
+  model <- lm(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs, data = bootstrap_sample) # fit the model using the bootstrap sample
+
+  predictions <- predict(model, newdata = train_data) # make predictions with the training data
+
+  preds_bs[[i]] <- predictions # store predictions in the empty vector
+}
+# create an empty array to store the predictions
+num_samples <- length(preds_bs)
+num_datapoints <- length(preds_bs[[1]])  
+preds_array <- array(NA, dim = c(num_samples, num_datapoints))
+
+# fill the array with predictions from bootstrappping
+for (i in 1:num_samples) {
+  preds_array[i,] <- unlist(preds_bs[[i]])
+}
+
+# find the median and 95% confidence intervals of each prediction
+preds <- preds_array %>%
+          apply(2, quantile,  c(0.05, 0.5, 0.95)) %>%
+          t()
+
+# create a data frame with all the necessary variables
+finalbootstrap <- data.frame(
+  observed = c(train_data$pct_upreg_sarc), 
+  point_estimates = c(trainpred$.pred),
+  medians = preds[, 2],
+  lower_bound = preds[, 1],
+  upper_bound = preds[, 3]
+)
+
+# create visualization with all 5 variables 
+figure2 <- ggplot(finalbootstrap, aes(x = observed, y = point_estimates)) +
+              geom_point(color = "black") +  
+              geom_point(aes(y = medians), color = "coral") + 
+              geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.1, color = "blue") + 
+              geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +  # add 45                 degree line
+              xlim(0, 50) + # limit the x-axis values
+              ylim(0, 50) + # limit the y-axis values
+              labs(x = "Observed Values", y = "Predicted Values", title = "Uncertainty in Predictions with Final Model")
+figure2
+
+# save final figure
+figure_file2 = here("results", "figures", "final-figures", "bootstrappred.png")
+ggsave(filename = figure_file2, plot=figure2) 
+```
+This figure shows that 95% confidence intervals, shown as dark blue lines, for the predictions that do not fall on the expected 45 degree line do not intersect with the line either.
+The dark blue lines show that the confidence intervals are narrow, which increases the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions.
+
+Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression.
+This regression analysis shrinks the coefficients towards the mean, which allows it to perform subset selection.
+Coefficients can reach zero, and that results in that variable being removed from the model.
+This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting.
+```{r}
+# set seed for reproducibility
+set.seed(2468)
+
+# LASSO regression workflow
+lasso_mod <- linear_reg(penalty = 0.1, mixture = 1, mode = "regression") %>% #set the penalty
+                set_engine("glmnet")
+
+lasso_results <- lasso_mod %>% 
+            fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg +
+                  rate_matdeaths_abs, data = mydata)
+
+# print results 
+tidy(lasso_results)
+
+# save results from fit into a data frame 
+lmtable11 <- tidy(lasso_results)
+
+# save fit results table  
+table_file16 = here("results", "tables", "resulttable16.rds")
+saveRDS(lmtable11, file = table_file16)
+```
+As expected, the coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
+The `sarc_standard` predictor still has the strongest relationship with the outcome.
+The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome.
+The direction of the relationship between the rate of maternal deaths due to abortions and the outcome flipped from the first full regression model to this model, but the coefficient is so small that this difference is not significant.
+
+It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best.
+We will use a grid search to ensure that we try all combinations of the parameters in our grid.
+We will complete the tuning process with a 5-fold cross-validation repeated 5 times.
+```{r}
+# set seed for reproducibility
+set.seed(2468)
+
+# repeat lasso model workflow
+lasso_mod2 <- linear_reg(penalty = tune(), mixture = 1, mode = "regression") %>% # allow for tuning
+                set_engine("glmnet")
+
+# change variable type of region for tune_grid()
+mydata$region <- as.numeric(mydata$region)
+
+# define the recipe
+recipe <- recipe(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg +
+                  rate_matdeaths_abs, data = mydata)
+
+# define the workflow
+workflow <- workflow() %>%
+  add_recipe(recipe) %>%
+  add_model(lasso_mod2)
+
+# define a parameter grid with 50 values linearly spaced on a log scale
+model_grid <- grid_regular(penalty(range = c(log10(1e-5), log10(1e2))), levels = 50)
+
+# create the real samples cross-validation 
+real_samples <- vfold_cv(mydata, v = 5, repeats = 5)
+
+# tune the model 
+tuned_model <- workflow %>% 
+                tune_grid(resamples = real_samples,
+                          grid = model_grid,
+                          metrics = metric_set(rmse))
+
+# observe the tuning process
+figure3 <- tuned_model %>% 
+             autoplot()
+figure3
+
+# save the final figure 
+figurefile3 <- here("results", "figures", "final-figures", "lassoparameters.png")
+ggsave(filename = figurefile3, plot = figure3)
+```
+As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
+The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1.
+When the regularization penalty increases to 10, the RMSE increases to almost 11.
+These observations are expected because the RMSE can only increase in LASSO regression because the penalty parameter is added to the SSR term to determine the value of the cost function.
+It is expected that the model will perform worse according to the RMSE scale as the penalty parameter increases.
+This plot shows that penalty parameters less than 1 are preferable for model performance.
+Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression.

### 2024-04-02 products/manuscript/Manuscript.qmd
Part 4 Updates
Completed bootstrapping and graphed the results. Added LASSO regression and tuned the parameters. Organized all results in the manuscript.

@@
 The Manuscript.qmd file for this document can be found in the products > manuscript folder.
-
-## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-# knitr::include_graphics(here("products","assets","antigen-recognition.png")) #commented out for the sake of knitting the file at this stage in the project 
-```
 
@@
 ```
+Details about the 9 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.
+
+## Full analysis
+
+_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
+This might again be code that is best placed in one or several separate R scripts that need to be well documented.
+You want the code to produce figures and data ready for display as tables, and save those.
+Then you load them here._
+
+We wanted to analyze the differences in the outcome between regions further, so we completed an ANOVA test and subsequet Tukey HSD test.
 The Tukey HSD test for the region predictor, shown in @statistical-analysis3, revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia.
@@
 ```
-
 We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4.
@@
 ```
-The full statistical analysis can be found in the statistical-analysis.qmd file in the code > analysis-code folder.
+To determine which model performed the best, we started by examining the RMSE and MAE performance metrics.
+We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion.
+@statistical-analysis5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
+The model with the population size using SARCs is the only single predictor model that performs better than the null.
+The full model with all 5 predictors peforms the best with the lowest RMSE and MAE values.
+```{r}
+#| label: statistical-analysis5
+#| fig-cap: "Performance Metric Comparisons"
+#| echo: FALSE
+readRDS(here("results","tables", "resulttable14.rds")) 
+```
+We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces.
+The cross-validated results, shown in @statistical-analysis6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
+The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319.
+While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model.
+Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528.
+This means there is higher variance in the single predictor model.
+The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data.
+Overall, the multiple linear regression model performs better.
+```{r}
+#| label: statistical-analysis6
+#| fig-cap: "Cross Validation Performance Metric Comparisons"
+#| echo: FALSE
+readRDS(here("results","tables", "resulttable15.rds")) 
+```
 
-## Full analysis
+Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance.
+As seen in @statistical-analysis7, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
+The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
+While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data.
+```{r}
+#| label: statistical-analysis7
+#| fig-cap: "Evaluating Final Model Performance on Test Data"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
+```
+We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data.
+The figure shown in @statistical-analysis8 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
+The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction.
+For the predictions that do not fall on the expected 45 degree line, the confidence intervals do not intersect with the line either.
+These findings increase the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions.
+```{r}
+#| label: statistical-analysis8
+#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
+```
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
+Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression.
+This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting.
+As seen in @statistical-analysis9, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
+The `sarc_standard` predictor still has the strongest relationship with the outcome.
+The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome.
+This process confirms that our final model should contain all 5 predictors.
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: statistical-analysis9
+#| fig-cap: "LASSO Regression Results"
 #| echo: FALSE
-# resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
-# knitr::kable(resulttable2) #commented out for the sake of knitting our project at this stage
+readRDS(here("results","tables", "resulttable16.rds")) 
+```
+It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @statistical-analysis10.
+As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
+The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1.
+When the regularization penalty increases to 10, the RMSE increases to almost 11.
+Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression based on an RMSE value slightly greater than 5, which i comparable to the RMSE found in the first regression model with all 5 predictors.
+```{r}
+#| label: statistical-analysis10
+#| fig-cap: "Parameter Tuning for LASSO Regression"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
 ```

### 2024-04-04 products/manuscript/Manuscript.qmd
Worked on abstract.

@@
 _Write a summary of your project._
+Our study investigates the association between unintended pregnancy rates and short-acting reversible contraception (SARC) use across low- and middle-income countries.
+Data from the Guttmacher Institute's Adding It Up project was used, which included information on reproductive-aged women (15-49) and various demographic factors.
+Single and multivariate linear regression models were employed to examine relationships between predictor variables and the outcome variable (unintended pregnancy rate).
+
+The exploratory data analysis revealed a less skewed distribution of unintended pregnancies among SARC users compared to other contraceptive methods.
+This led researchers to focus on SARC use for the remainder of the study.
+The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia.
+The Oceania and Europe region had weaker associations.
+Marriage status (both currently and never married) was inversely associated with unintended pregnancy rates.
+This means that as the percent of women who are currently married or never married increases, the unintended preganacy rate decreses.
+Maternal mortality rates, however, showed a positive correlation.
+
+A full model incorporating five predictors (total SARC users, region, marital status, and maternal mortality rates due to unintended pregnancy and abortion) was built.
+Despite performing the best among the evaluated models and being validated through cross-validation, the full model exhibited tendencies towards overfitting the data.
+
+Lasso regression, a technique that performs variable selection, was then employed to address these overfitting concerns.
+The analysis identified the total number of SARC users as the strongest predictor, followed by region.
+Marital status and maternal mortality rates displayed weaker negative associations with unintended pregnancy rates.
+While the direction of the association between abortion-related maternal mortality and unintended pregnancy rates flipped in the Lasso model, the coefficient (effect size) remained minimal.
+
+Overall, this study highlights the influence of regional variations, SARC use, and marital status on unintended pregnancy rates in low- and middle-income countries.
+The findings suggest a need for further investigation into the factors influencing maternal mortality rates and unintended pregnancies.
 

### 2024-04-04 code/eda-code/eda.qmd
Updated Predictor Variable table

@@
 # Selecting variables
-selected_variables <- mydata[c("pct_upreg_sarc", "region", "wra", "pct_currentlymarried", "pct_nevermarried","rate_matdeaths", "rate_matdeaths_upreg", "rate_matdeaths_abs")]
-
-selected_variables <- selected_variables %>% 
-  arrange(region)
+selected_variables <- mydata[c("pct_upreg_sarc","pct_currentlymarried","pct_nevermarried","rate_matdeaths", "rate_matdeaths_upreg", "rate_matdeaths_abs","region", "wra")]
 
 # Create table 
-kable(selected_variables, caption = "Table Depicting Predictor Variables", booktabs = TRUE) %>%
-  kable_styling(full_width = F, font_size = 10)
+selected_variables %>%
+  tbl_summary(
+    by = region,
+    label = list(wra ~ "Number of women of reproductive age, 15-49"),
+    type = all_continuous() ~ "continuous2",
+    statistic = list(all_continuous() ~ c("{mean}","{min}, {max}"))) %>%
+modify_header(label ~ "**Predictor Variable**") %>%
+modify_caption("**Table 1. Predictor Variables**") %>%
+modify_spanning_header(all_stat_cols() ~ "Region")
 

### 2024-04-05 products/manuscript/Manuscript.qmd
Updated our references section.

@@
 A full model incorporating five predictors (total SARC users, region, marital status, and maternal mortality rates due to unintended pregnancy and abortion) was built.
-Despite performing the best among the evaluated models and being validated through cross-validation, the full model exhibited tendencies towards overfitting the data.
+Despite performing the best among the evaluated models and being validated through cross-validation and bootstrap method, the full model exhibited tendencies towards overfitting the data.
 
@@
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
 {{< pagebreak >}}
@@
 # References
+_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
 
-- https://pubmed.ncbi.nlm.nih.gov/32981858/
-- https://www.sciencedirect.com/science/article/pii/S001078241730478X 
-- https://www.sciencedirect.com/science/article/abs/pii/S0020729211005406 #legal status article
-- https://www.guttmacher.org/report/adding-it-up-investing-in-sexual-reproductive-health-2019-methodology (Guttmacher source)
-- https://pubmed.ncbi.nlm.nih.gov/35332057/ (original source of data)
+[@heviaLegalStatusEmergency2012], [@jonathanmarcbearakCountryspecificEstimatesUnintended2022], [@kavanaughContraceptiveMethodUse2018], [@rivetContraceptionUseAttitudes2021]

### 2024-04-05 products/manuscript/supplement/Supplementary-Material.qmd
Part 4 Updates
Added method descriptions in the supplemental file and wrote instructions for reproducibility. Updated manuscript with code processing summary and CV/bootstrapping explanations.

@@
 
-For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
-
 # Overview
@@
 
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
+Reproducing this project requires R, RStudio, Microsoft word, and LaTex.
+The README files in each folder contain descriptions of the documents located in that folder to help guide users through the reproducible processes.
+
+Begin with the processing-file.qmd in the code > processing-code folder.
+This file loads the original dataset from the Guttmacher Institute's Adding It Up project, loads the codebook for the dataset, and glimpses at the data.
+The processing-code.qmd file shows how we modified existing variable, created new variables, and saved the processed data.
+The processed data is saved in the data > processed-data folder as processeddata.rds.
+
+Visit the eda.qmd file in the code > eda-code folder to find code for the exploratory data analysis process.
+This code generates all the exploratory figures presented in the Exploratory/descriptive analysis results in the manuscript.qmd file as well as additional figures for the outcome not selected for the final analysis as shown later in this supplementary material.
+All exploratory figures are saved in the results > figures > exploratory-figures folder.
+
+Continue to the statistical-analysis.qmd file in the code > analysis-code folder to find code for the statistical analysis and model building portion of the study.
+All model results tables are stored in the results > tables folder.
+All model figures examining model performance are stored in the results > figures > final-figures folder.
+
+View the manuscript in the products > manuscript folder to find a summary of the study along with the most important findings and final conclusions.
+This supplemental material is stored in the products > manuscript > supplement folder.
 
@@
 
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
+We only briefly mentioned that we used cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript.
+Cross-validation was chosen as a method for determining how well the model performs on data it has not seen before because that is the main point of building models.
+Our original models, found in resulttables1-10 in the results > tables folder, are created using all of the data.
+The cross-validation process built models with 75% of the data that was designated as the training data, and the remaining 25% was used as testing data.
+Our dataset is small with only 132 observations, so the insights provided here are not robust.
+However, we still want to know how well these variables predict our outcome to determine how meaningful each predictor is, so we completed cross-validation anyways to answer this question.
+We also used this approach due to the risk of overfitting in our model based on interaction between variables.
+Plotting the observed and predicted values for our model allowed us to look for systematic deviations from the expected relationship visualized with a 45 degree line as shown in @supplement-1.
+Since we did conclude that the model overfits the data based on unhealthy deviations of the testing data from the 45 degree line, our use of cross-validation is appropriate and useful.
+Lastly, this approach is an easily interpretable way to increase confidence in the performance of our models based on metrics because both proceses lead to the same conclusions.
+```{r}
+#| label: supplement-1
+#| fig-cap: "Evaluating Final Model Performance on Test Data"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
+```
+
+Bootstrapping was chosen to determine variability and uncertainty in the data.
+This process is similar to cross-validation, but the resampling process simulates a new dataset by performing sampling with replacement to obtain the same sample size as the original dataset.
+This is crucial for our small dataset of 132 observations, which is why we chose to bootstrap with 100 resamples.
+We are still limited by our small sample size for this approach, but it is worth completing for the purpose of investigating uncertainty.
+The model estimates parameters for each of the 100 samples, which creates a distribution of parameter estimates that can be used to create confidence intervals.
+We chose to create standard 95% confidence intervals based on convention, which can be seen in @supplement-2.
+```{r}
+#| label: supplement-2
+#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
+```
+
+LASSO regression is important in this study because we need to manage the bias-variance tradeoff since we have evidence that our model overfits the data.
+By applying a penalty parameter to the cost function, LASSO regression shrinks the coefficients of the parameter estimates, so these estimates will always be smaller than the ones provided by the original linear model.
+As the penalty parameter increases, a larger penalty term is added to the SSR, which increases the value of the cost function.
+As the penalty increases, the RMSE value increases, and the performance of the model is considered worse and worse comparatively.
+We chose a parameter of 0.1 first because that is standard for LASSO regression, but we also completed a parameter tuning process to validate that choice.
+As shown in @supplement-3, the RMSE performance metric increases exponentially when the penalty parameter increases past 1, which validates our previous selection of 0.1 for the penalty parameter because that will produce the lowest RMSE possible.
+Since coefficients can be reduced all the way to zero, this analysis completes feature selection as well.
+Shrinking the coefficients also helps combat colinearity in variables, which we established between the percent of currently married women and percent of never married women, and the three rates of maternal mortality, respectively.
+LASSO models are also easier to interpet, which increases the validity of our findings.
+```{r}
+#| label: supplement-3
+#| fig-cap: "Parameter Tuning for LASSO Regression"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
+```
+
+# Additional data cleaning explanation
+The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
+We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
+This variable is only 82% complete, but the individual percent of unintended pregnancy variables correponding with different types of birth control methods are each 100% complete.
+Since the missingness of this data is not well explained except by the hypothesis that this variable was found using methods outside of the AIU protocol, we decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
+We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
+```{r}
+#| label: tbl-summarytable2
+#| tbl-cap: "Summary table to discover missingness"
+#| echo: FALSE
+resulttable2 <- readRDS("../../results/tables/summarytable.rds")
+kable(resulttable2)
+```
 
@@
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
-
-Readers will be able to find our analysis of the different contraception methods that we did not select for our main outcome through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables.
+Readers will be able to find our analysis of the different contraception methods that we did not select for our main outcome through figures we've made for the `pct_upreg_larcster`, `pct_upreg_trad`, and `pct_upreg_nouse` variables in this section.
 
@@
 
-## Additional result
+We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable1.
+These variables were not used in the main final analysi of the paper.
+We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
+Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
+It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
+```{r}
+#| label: tbl-summarytable1
+#| tbl-cap: "Summary table for categorical variables"
+#| echo: FALSE
+resulttable1 <- readRDS("../../results/tables/summarytablecat.rds")
+kable(resulttable1)
+```
+
+# Additional data cleaning explanation
 The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
@@
 ```{r}
-#| label: tbl-summarytable1
+#| label: tbl-summarytable2
 #| tbl-cap: "Summary table to discover missingness"
 #| echo: FALSE
-resulttable1 <- readRDS("../../results/tables/summarytable.rds")
-kable(resulttable1)
-```
-
-We explored the distribution of the countries by five categorical variables: `region`, `larcster_usage`, `sarc_usage`, `trad_usage`, and `none_usage` which is displayed in @tbl-summarytable2.
-We found that most countries are located in Africa followed by Asia and Latin America/the Caribbean.
-Only 20 countries have high usage of long acting reversible and sterilization methods compared to 33 countries with high usage of short acting reversible and traditional methods.
-It is interesting that the distributions for short acting reversible method, traditional method, and no method usages are the same.
-```{r}
-#| label: tbl-summarytable2
-#| tbl-cap: "Summary table for categorical variables"
-#| echo: FALSE
-resulttable2 <- readRDS("../../results/tables/summarytablecat.rds")
+resulttable2 <- readRDS("../../results/tables/summarytable.rds")
 kable(resulttable2)
-```
-
-@tbl-resulttable1 shows an additional table summarizing a model fit.
-
-```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
-#| echo: FALSE
-resulttable1 = readRDS(here("results","resulttable1.rds"))
-knitr::kable(resulttable1)
-```
-
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight.png"))
 ```
@@
 
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
 {{< pagebreak >}}

### 2024-04-05 code/eda-code/eda.qmd
Updated manuscript to reflect the updates for the pred.table

@@
 # Create table 
-selected_variables %>%
+predictortable <-selected_variables %>%
   tbl_summary(

### 2024-04-05 products/manuscript/Manuscript.qmd

@@
 # Summary/Abstract
-_Write a summary of your project._
 Our study investigates the association between unintended pregnancy rates and short-acting reversible contraception (SARC) use across low- and middle-income countries.
@@
 Marriage status (both currently and never married) was inversely associated with unintended pregnancy rates.
-This means that as the percent of women who are currently married or never married increases, the unintended preganacy rate decreses.
-Maternal mortality rates, however, showed a positive correlation.
+This means that as the percent of women who are currently married or never married increases, the unintended preganacy rate decreases.
+Maternal mortality rates showed a positive correlation with the outcoe.
 
 A full model incorporating five predictors (total SARC users, region, marital status, and maternal mortality rates due to unintended pregnancy and abortion) was built.
-Despite performing the best among the evaluated models and being validated through cross-validation and bootstrap method, the full model exhibited tendencies towards overfitting the data.
+Despite performing the best among the evaluated models and being validated through cross-validation and the bootstrap method, the full model exhibited tendencies towards overfitting the data.
 
-Lasso regression, a technique that performs variable selection, was then employed to address these overfitting concerns.
+Lasso regression, a technique that performs variable selection, was employed to address these overfitting concerns.
 The analysis identified the total number of SARC users as the strongest predictor, followed by region.
@@
 ## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
 
@@
 ## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
 The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world.
@@
 ## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
 
-1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region?
+Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region?
 
@@
 # Methods 
-
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
 Reproducing this project requires R, RStudio, and Microsoft word.
@@
 The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing.
-All of these figures are saved in the Supplementary-Material.qmd file in the products > manuscript > supplement folder.
+All of these figures are displayed and summarized in the Supplementary-Material.qmd file in the products > manuscript > supplement folder.
 This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods as our single outcome because this method is becoming increasingly popular, so we feel that results from this contraception method will be the most relevant.
 Additionally, the exploratory data analysis process showed that the distribution of this outcome variable is less skewed than the other options, and the stratification by region is clearer, which allows for better understanding.
+All exploratory figures are saved in the results > figures > exploratory-figures folder.
 
@@
 Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before.
+This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets.
+To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions.
+This process is similar to cross-validation because it relies on resamling the training data.
+Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance.
+It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best.
+We use a grid search to ensure that we try all combinations of the parameters in our grid.
+We complete the tuning process with a 5-fold cross-validation repeated 5 times.
+We are still considering adding a random forest procedure to examine the variance in the model.
+Each of these processes are crucial in determining which model performs best for our dataset.
+Result tables for each model and test are stored in the results > tables folder.
+Figures created from cross-validation, boostrapping, and LASSO regression are stored in the results > figure > final-figures folder.
 
@@
 ## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-
 This research aims to analyze the factors influencing unintended pregnancies and contraception use in low-and-middle-income countries.
@@
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+As shown in the processing-file.qmd in the code > processing-code folder, we import the data from a .csv file that was downloaded directly from the Guttmacher AIU project website.
+We also included the codebook in this file.
+After taking a glimpse at the data, we identified the variables we need to clean and created a new dataset with those variables only.
+We explored missingness in the data, which is further explained in the supplementary-material.qmd file found in the products > manuscript > supplement folder.
+Simple processing tasks include converting variable types, converting decimal value into true percentages, creating variables to find total usage of each birth control method across age groups, exploring outliers, and creating new standardized total usage values based on population size.
 
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
 We began with a simple linear regression model with each of the 7 predictor variables to determine the association betwen each variable and the outcome.
@@
 ## Basic statistical analysis
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
 We completed a simple linear regression analysis with each of the 7 predictors shown in @statistical-analysis1.
@@
 ## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
 We wanted to analyze the differences in the outcome between regions further, so we completed an ANOVA test and subsequet Tukey HSD test.

### 2024-04-05 products/manuscript/Manuscript.qmd
Attempt at making highlighted references

@@
 
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,  short acting reversible, and traditional.
-Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies.
-According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group.
-Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study.
+We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional.
+Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022].
+According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group .
+Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021].
 African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies.
@@
 Argentina and Brazil lack legal recognition, but offer protocols and guides.
-Mexico requires provision to victims of sexual and domestic violence (Hevia, 2012).
+Mexico requires provision to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012].
+Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].
 

### 2024-04-15 code/analysis-code/statistical-analysis.qmd
Part 4 Updates
Added method descriptions in the supplemental file and wrote instructions for reproducibility. Updated manuscript with code processing summary and CV/bootstrapping explanations.
Removed ANOVA and Tukey.
Changed lmfit9 to null model

@@
 The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.
-
-### ANOVA for region variable
-We want to compare the mean percent of unintended pregnancies among those using short acting reversible methods across the five regions: Asia, Africa, Europe, Latin America and the Caribbean, and Oceania.
-We will start with an ANOVA test to determine the relationship between region and outcome.
-We are more interested in pairwise comparisons between the regions, so we also conducted Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA.
-```{r}
-# conduct an ANOVA test with the region variable
-regionANOVA <- aov(pct_upreg_sarc ~ region, data = mydata)
-
-# explore ANOVA results 
-summary(regionANOVA)
-
-# add Tukey's HSD test for pairwise comparisons
-tukey_result <- TukeyHSD(regionANOVA)
-
-# explore the pairwise comparisons
-tidy(tukey_result)
-
-# save fit results table  
-table_file13 = here("results", "tables", "resulttable13.rds")
-saveRDS(tukey_result, file = table_file13)
-```
-ANOVA results show that the region variable is statistically significantly associated with the outcome variable given the p-value less than 0.001.
-The Tukey HSD test revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia.
-The relationships between Oceania and Europe and Latin America/the Caribbean are statistically significant with p-values less than 0.0001.
-Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.
-The lowest difference in outcome between regions is between Asia and Africa, but this is not a statistically significant association.
-Two statistically significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.
 
@@
 # create predictions using the null model
-nullpred <- predict(lmfit9, new_data = mydata) %>% 
+nullpred <- predict(null_model, new_data = mydata) %>% 
                   select(.pred)
@@
 Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression.
+
+We are considering adding a random forest model to further examine the variance in our model.

### 2024-04-15 code/eda-code/eda.qmd
Added labels to figures in manuscript, switched axis labels.

@@
 # Explore the percentage of unintended pregnancies related to short acting reversible methods in each country
-p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
+plot_1 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
   geom_histogram() + 
   ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
-plot(p3)
+plot(plot_1)
 figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png")
-ggsave(filename = figure_file, plot=p3)
+ggsave(filename = figure_file, plot=plot_1)
 
@@
 
-P3: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
+Plot 1: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods.
 About half of the countries in the dataset report less than 20% of pregnancies are unintended when women use this form of birth control.
@@
 # Explore the total number of people using Short Acting Reversible Methods in each country
-p7 <- mydata %>% ggplot(aes(x=total_sarc)) + 
+plot_2 <- mydata %>% ggplot(aes(x=total_sarc)) + 
   geom_histogram() + 
   ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
-plot(p7)
+plot(plot_2)
 figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
-ggsave(filename = figure_file, plot=p7)
+ggsave(filename = figure_file, plot=plot_2)
 
@@
 
-Plot 11 uses `pct_upreg_sarc` and `sarc_standard` to examine the same relationship as plot 10 but with short-acting reversible methods. 
+Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
 ```{r}
@@
 # explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
-p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
+plot_3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
   geom_point() + 
@@
   ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p11)
+plot(plot_3)
 figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
-ggsave(filename = figure_file, plot=p11)
+ggsave(filename = figure_file, plot=plot_3)
 
@@
 
-P11: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.
+Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.
 
@@
 # explore percent unintended pregnancies using short acting reversible methods by region
-p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
+plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
   geom_point() + 
@@
   ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
-plot(p15)
+plot(plot_4)
 figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
-ggsave(filename = figure_file, plot=p15)
+ggsave(filename = figure_file, plot=plot_4)
 
@@
 
-P15: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
+Plot 4: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
 
@@
 # explore short acting reversible method usage compared to percent of unintended pregnancies
-p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
+plot_5 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
   geom_violin() + ggtitle("Usage of SARCs vs Percent of Unintended Pregnancies") +
   labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
-plot(p19)
+plot(plot_5)
 figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
-ggsave(filename = figure_file, plot=p19)
+ggsave(filename = figure_file, plot=plot_5)
 
@@
 
-P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+Plot 5: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
 
@@
 # explore the region predictor
-p22 <- mydata %>% ggplot(aes(x=region)) + 
+plot_6 <- mydata %>% ggplot(aes(x=region)) + 
   geom_bar() +
@@
   xlab("Region") + ylab("Number of Countries")
-plot(p22)
+plot(plot_6)
 figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
-ggsave(filename = figure_file, plot=p22)
+ggsave(filename = figure_file, plot=plot_6)
 
 # explore the current marriage predictor
-p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
+plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
   geom_histogram() +
@@
   xlab("Percentage") + ylab("Number of Countries")
-plot(p23)
+plot(plot_7)
 figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
@@
 # explore the never married predictor
-p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
+plot_8 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
   geom_histogram() +
@@
   xlab("Percentage") + ylab("Number of Countries")
-plot(p24)
+plot(plot_8)
 figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
-ggsave(filename = figure_file, plot=p24)
+ggsave(filename = figure_file, plot=plot_8)
 
 # explore the total maternal mortality rate predictor
-p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
+plot_9 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
   geom_histogram() +
@@
   xlab("Maternal Deaths") + ylab("Number of Countries")
-plot(p25)
+plot(plot_9)
 figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
-ggsave(filename = figure_file, plot=p25)
+ggsave(filename = figure_file, plot=plot_9)
 
@@
 # explore the maternal mortality rate from unintended pregnancies predictor
-p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
+plot_10 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
   geom_histogram() +
@@
   xlab("Maternal Deaths") + ylab("Number of Countries")
-plot(p26)
+plot(plot_10)
 figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
-ggsave(filename = figure_file, plot=p26)
+ggsave(filename = figure_file, plot=plot_10)
 
@@
 # explore the maternal mortality rate from abortions
-p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
+plot_11 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
   geom_histogram() +
@@
   xlab("Maternal Deaths") + ylab("Number of Countries")
-plot(p27)
+plot(plot_11)
 figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
-ggsave(filename = figure_file, plot=p27)
+ggsave(filename = figure_file, plot=plot_11)
 
@@
 ```
-P22: The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
+Plot 6: The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
 about 35 countries.
@@
 
-P23: The distribution of women currently married is approximately normal with an average around 57% and no major outliers.
+Plot 7: The distribution of women currently married is approximately normal with an average around 57% and no major outliers. 
+Plot 8: The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.
 
-P24: The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.
-
-P25: The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers.
+Plot 9: The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers.
 We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.
 
-P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
+Plot 10: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
 Chad and South Sudan also account for the highest rates of maternal mortality in this category.
 
-P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.
+Plot 11: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.

### 2024-04-15 products/manuscript/Manuscript.qmd
Added labels to figures in manuscript, switched axis labels.

@@
 #| label: tbl-summarytable
-#| tbl-cap: "Summary table of numeric variables"
+#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
 #| echo: FALSE
@@
 #| label: exploratory-result
-#| fig-cap: "Histograms of Percentage of Unintended Pregnancy Use by Birth Control Method"
+#| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
 #| echo: FALSE
@@
 #| label: exploratory-result2
-#| fig-cap: "Relationship between Percentage of Unintended Pregnancies and Standardized Population Total"
+#| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
 #| echo: FALSE
@@
 
-We stratified these four graphs by region to examine how the relationship differs across each of the t regions which is shown in @exploratory-result3 .
+We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @exploratory-result3 .
 The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
@@
 #| label: exploratory-result3
-#| fig-cap: "Percentage of Unintended Pregnancies Stratified by Region"
+#| fig-cap: "Figure 2: Graph Depicting Percent Unintended Pregnancies Using SARC by region"
 #| echo: FALSE
@@
 #| label: exploratory-result4
-#| fig-cap: "Violin Plots with Birth Control Usage Levels"
+#| fig-cap: "Plot 5: Violin Plot Depicitng Usage of SARc Per Region"
 #| echo: FALSE
@@
 #| label: statistical-analysis1
-#| fig-cap: "Strongest Predictor: sarc_standard"
+#| fig-cap: "Table 2: Strongest Predictor: sarc_standard"
 #| echo: FALSE
@@
 #| label: statistical-analysis2
-#| fig-cap: "Regional Predictor"
+#| fig-cap: "Table 3: Regional Predictor"
 #| echo: FALSE
@@
 ## Full analysis
-We wanted to analyze the differences in the outcome between regions further, so we completed an ANOVA test and subsequet Tukey HSD test.
-The Tukey HSD test for the region predictor, shown in @statistical-analysis3, revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia.
-The differences between Oceania and Europe and Latin America/the Caribbean are the largest.
-Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users.
-Two more significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome.
-```{r}
-#| label: statistical-analysis3
-#| fig-cap: "Tukey HSD Test Results"
-#| echo: FALSE
-readRDS(here("results","tables", "resulttable13.rds")) 
-```
+
 We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4.
@@
 #| label: statistical-analysis4
-#| fig-cap: "Multiple Linear Regression with 5 Predictors"
+#| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
 #| echo: FALSE
@@
 #| label: statistical-analysis5
-#| fig-cap: "Performance Metric Comparisons"
+#| fig-cap: "Table 5: Performance Metric Comparisons"
 #| echo: FALSE
@@
 #| label: statistical-analysis6
-#| fig-cap: "Cross Validation Performance Metric Comparisons"
+#| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
 #| echo: FALSE
@@
 #| label: statistical-analysis7
-#| fig-cap: "Evaluating Final Model Performance on Test Data"
+#| fig-cap: "Figure 3: Evaluating Final Model Performance on Test Data"
 #| echo: FALSE
@@
 #| label: statistical-analysis8
-#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
+#| fig-cap: "Figure 4: Evaluation of Uncertainty in Full Model Predictions"
 #| echo: FALSE
@@
 #| label: statistical-analysis9
-#| fig-cap: "LASSO Regression Results"
+#| fig-cap: "Table 7: LASSO Regression Results"
 #| echo: FALSE
@@
 #| label: statistical-analysis10
-#| fig-cap: "Parameter Tuning for LASSO Regression"
+#| fig-cap: "Figure 5: Parameter Tuning for LASSO Regression"
 #| echo: FALSE

### 2024-04-15 code/analysis-code/statistical-analysis.qmd
Part 5 Updates
Added random forest model and tuned the parameters. Found AIC criteria and analyzed. Added explanation of outcome predictor to clear up confusion.

@@
 library(glmnet) # for LASSO regression
+library(ranger) # for random forest model 
 ```
@@
 The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
-While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data.
+While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data because the testing observations are so far from the expected relationship.
 
@@
 
-We are considering adding a random forest model to further examine the variance in our model.
+We also want to explore a random forest model with all 5 main predictors since the previous cross-validation process revealed high overfitting of the multivariate linear regression model.
+We will use RMSE, MAE, and R squared as performance metrics to compare it to the previous LASSO regression model.
+```{r}
+#| message: false
+# set seed for reproducibility
+set.seed(2468)
+
+# create workflow for random forest model
+rf_mod <- rand_forest(mode = "regression") %>% # specify the model type
+            set_engine("ranger", seed = 2468) #  set seed for reproducibility internally
+
+random_forest <- rf_mod %>% 
+                    fit(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg +
+                    rate_matdeaths_abs, data = mydata) # include all predictors
+
+# create predictions using the random forest model
+RFpred <- predict(random_forest, new_data = mydata) %>% 
+                  select(.pred)
+
+# create a data frame with the predictions and true values 
+RFdata <- bind_cols(allpred, mydata$pct_upreg_sarc) %>%
+            rename(pct_upreg_sarc = "...2")
+
+# find performance metrics to determine model fit
+rmse10 <- rmse(RFdata, truth = pct_upreg_sarc, estimate = .pred)
+mae10 <- mae(RFdata, truth = pct_upreg_sarc, estimate = .pred)
+rsq10 <- rsq(RFdata, truth = pct_upreg_sarc, estimate = .pred)
+
+# print performance metric results 
+print(rmse10)
+print(mae10)
+print(rsq10)
+
+# visualize random forest predictions verus observed values 
+figure4 <-ggplot(RFdata, aes(x=.pred, y= pct_upreg_sarc)) + 
+                geom_point() +
+                geom_abline(linetype = "dashed") + 
+                labs(x="Predicted Values", y="Observed Values", title="Random Forest Model Predictions and Observations")
+figure4
+
+# save the random forest figure 
+figurefile4 <- here("results", "figures", "final-figures", "randomforestpred.png")
+ggsave(filename = figurefile4, plot = figure4)
+```
+The random forest model has an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.8369754.
+Compared to the previous model with an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.820600, the random forest model performs very similarly.
+The RMSE and MAE values are exactly the same beause the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
+Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
+The plot of the predicted versus observed values shows strong correlation for predicted values less than 15, but the deviation from the expected values increases as the predicted values increase.
+
+It is crucial to tune the parameters associated with the random forest model, so we will do that before concluding that it outperforms the LASO regression model.
+We updated the workflow to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
+We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so.
+```{r}
+#| message: false
+# set seed for reproducibility
+set.seed(2468)
+
+# define the recipe
+recipe <- recipe(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg +
+                    rate_matdeaths_abs, data = mydata)
+
+# update workflow accordingly 
+rf_mod2 <- rand_forest(mode = "regression",
+                      mtry = tune(),  # allow for tuning
+                      min_n = tune(), # allow for tuning
+                      trees = 300) %>% 
+                          set_engine("ranger", seed = 2468)
+
+# use updated workflow
+RFworkflow <- workflow() %>%
+  add_recipe(recipe) %>%
+  add_model(rf_mod2)
+
+# define a 5 by 5 parameter grid
+RFparameter_grid <- grid_regular(mtry(range = c(1,5)),
+                            min_n(range = c(1,15)),
+                            levels = 5)
+
+# create the real samples cross-validation 
+real_samples <- vfold_cv(mydata, v = 5, repeats = 5)
+
+# tune the random forest model 
+tuned_RFmod <- RFworkflow %>% 
+                    tune_grid(resamples = real_samples,
+                          grid = RFparameter_grid,
+                          metrics = metric_set(rmse))
+
+# observe the tuning process
+rf_tune_plot <- tuned_RFmod %>% 
+              autoplot()
+rf_tune_plot
+
+# save plot 
+figure_file5 = here("results", "figures", "final-figures", "randomforesttune.png")
+ggsave(filename = figure_file5, plot=rf_tune_plot) 
+```
+Based on the red line that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
+This begs us to ask the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
+
+Since both of these models are overfitting the data, we think it is wise to examine an information criterion to determine how well our model might perform on unseen data.
+Even though cross-validation is a better approach for answering this question because it actually resamples the data, the concern of overfitting justifies finding the AIC value in this case.
+This measure still evaluates the trade-off between good fit to the data, shown by a low cost function, and model complexity based on the number of parameters.
+We will compare the single predictor model with the standarized `sarc_standard` variable and the full model because the models have already been shown to have the best performance metrics.
+```{r}
+#| message: false
+# define single predictor model in format conducive to finding AIC
+singlemod <- glm(pct_upreg_sarc ~ sarc_standard, data = mydata)
+
+# find AIC value 
+summary(singlemod)
+
+# define full predictor model in format conducive to finding AIC 
+fullmod <- glm(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs, data = mydata)
+
+# find AIC value
+summary(fullmod)
+```
+
+The single predictor model has an AIC of 812.95, and the full predictor model has an AIC of 803.21.
+The full predictor model has a lower AIC score, and the difference between the two models is 9.74, which means the difference between the two model is meaningful given that it is so close to 10.
+A lower AIC indicates a better performance, and the rule of thumb is that a difference of 10 between 2 models is meaningful.
+However, information criterion are not limited by an arbitrary cutoff point, so we feel confident that there is a difference between these two models.

### 2024-04-15 code/eda-code/eda.qmd
Part 5 Updates
Added random forest model and tuned the parameters. Found AIC criteria and analyzed. Added explanation of outcome predictor to clear up confusion.

@@
 ```{r}
-
 # explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods

### 2024-04-15 products/manuscript/Manuscript.qmd
Part 5 Updates
Added random forest model and tuned the parameters. Found AIC criteria and analyzed. Added explanation of outcome predictor to clear up confusion.

@@
 
-This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-
 ```{r, echo=FALSE, message=FALSE}
-# load a few R packages
+# load a few R packages for rendering 
 library(here)
@@
 We complete the tuning process with a 5-fold cross-validation repeated 5 times.
-We are still considering adding a random forest procedure to examine the variance in the model.
+We added a random forest model to attempt to improve model performance.
+We used another parameter tuning grid to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
+We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so.
 Each of these processes are crucial in determining which model performs best for our dataset.
 Result tables for each model and test are stored in the results > tables folder.
-Figures created from cross-validation, boostrapping, and LASSO regression are stored in the results > figure > final-figures folder.
+Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.
+Lastly, we examined the AIC scores for the two best performing models since there is strong evidence for overfitting.
+Even though cross-validation is a better approach for answering the question of how well a model predicted values in unseen data because it actually resamples the data, the concern of overfitting justifies finding the AIC value in this case.
+This measure still evaluates the trade-off between good fit to the data, shown by a low cost function, and model complexity based on the number of parameters.
 
@@
 Lastly, we used cross-validation to determine how well our multiple linear regression model performs on data it has not seen before.
+We also used LASSO regression and random forest modeling as forms of subset selection by applying a regularization penalty and generating a tree structure, respectively.
+Lastly, we examine the AIC values associated with our two best performing models.
 
@@
 There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected.
+Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception.
+This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place.
+This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users.
 This association is displayed in @exploratory-result2.
@@
 We completed a simple linear regression analysis with each of the 7 predictors shown in @statistical-analysis1.
-The predictor with the strongest relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
+The predictor with the strongest positive relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
+This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place.
 
@@
 
+After creating the random forest model, we found that the RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
+Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
+The plot of the predicted versus observed values shown in @statistical-analysis11 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
+```{r}
+#| label: statistical-analysis11
+#| fig-cap: "Predicted vs Observed Values in Random Forest Model"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "randomforestpred.png"))
+```
+
+Again, it is crucial to tune hyperparameters associate with machine learning models.
+Based on the red line displayed in @statistical-analysis12 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
+This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
+However, there is a minimal increase in RMSE when all five models are included in the model with a node size of 1.
+```{r}
+#| label: statistical-analysis12
+#| fig-cap: "Tuning Parameters for Random Forest Model"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "final-figures", "randomforesttune.png"))
+```
+
 {{< pagebreak >}}

### 2024-04-15 code/eda-code/eda.qmd
Outcome Description
Updated our outcome variable description in eda and manuscript files to better explain our results.

--- 2024-04-15 code/eda-code/eda.qmd
@@
 ```{r}
-
 # explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
-plot_3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
+plot_3 <- mydata %>% ggplot(aes(x=sarc_standard, y=pct_upreg_sarc)) + 
   geom_point() + 
   geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
 plot(plot_3)
@@
 
-Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.
+Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARCs, but there is considerably more variation in this relationship.
+The outcome of interest should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place.
+This percentage does not reflect the effectiveness of short-acting reversible contraception, but it does classify the reason for an unintended pregnancy.
+As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control.
+It makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users.
+This association is a little challenging to contextualize since we could not analyze all 4 classifications for the percentage of unintended pregnancies.
 

### 2024-04-15 products/manuscript/Manuscript.qmd
Outcome Description
Updated our outcome variable description in eda and manuscript files to better explain our results.

@@
 This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users.
-This association is displayed in @exploratory-result2.
+As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control.
+Each of these percentages are positively associated with number of people in its respective classification group that it represents.
+The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @exploratory-result2.
 ```{r}

### 2024-04-16 code/analysis-code/statistical-analysis.qmd
Project 5 Updates
Added message:false to make sure there is no raw code output in the manuscript. Added correct figure labels for rendering.

@@
 
-Since both of these models are overfitting the data, we think it is wise to examine an information criterion to determine how well our model might perform on unseen data.
-Even though cross-validation is a better approach for answering this question because it actually resamples the data, the concern of overfitting justifies finding the AIC value in this case.
-This measure still evaluates the trade-off between good fit to the data, shown by a low cost function, and model complexity based on the number of parameters.
-We will compare the single predictor model with the standarized `sarc_standard` variable and the full model because the models have already been shown to have the best performance metrics.
-```{r}
-#| message: false
-# define single predictor model in format conducive to finding AIC
-singlemod <- glm(pct_upreg_sarc ~ sarc_standard, data = mydata)
-
-# find AIC value 
-summary(singlemod)
-
-# define full predictor model in format conducive to finding AIC 
-fullmod <- glm(pct_upreg_sarc ~ sarc_standard + region + pct_currentlymarried + rate_matdeaths_upreg + rate_matdeaths_abs, data = mydata)
-
-# find AIC value
-summary(fullmod)
-```
-
-The single predictor model has an AIC of 812.95, and the full predictor model has an AIC of 803.21.
-The full predictor model has a lower AIC score, and the difference between the two models is 9.74, which means the difference between the two model is meaningful given that it is so close to 10.
-A lower AIC indicates a better performance, and the rule of thumb is that a difference of 10 between 2 models is meaningful.
-However, information criterion are not limited by an arbitrary cutoff point, so we feel confident that there is a difference between these two models.
+Choose final model with justification

### 2024-04-16 products/manuscript/Manuscript.qmd
Project 5 Updates
Added message:false to make sure there is no raw code output in the manuscript. Added correct figure labels for rendering.

@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @tbl-summarytable.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @table1.
 ```{r}
-#| label: tbl-summarytable
+#| message: false
+#| label: table1
 #| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
@@
 We created a multitude of exploratory plots to understand our data better.
-The first @exploratory-result shows the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
+The first exploratory result shown in @figure1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
 The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
 ```{r}
-#| label: exploratory-result
+#| message: false
+#| label: figure1
 #| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
@@
 Each of these percentages are positively associated with number of people in its respective classification group that it represents.
-The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @exploratory-result2.
+The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @figure2.
 ```{r}
-#| label: exploratory-result2
+#| message: false
+#| label: figure2
 #| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
@@
 
-We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @exploratory-result3 .
+We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @figure3 .
 The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
 ```{r}
-#| label: exploratory-result3
+#| message: false
+#| label: figure3
 #| fig-cap: "Figure 2: Graph Depicting Percent Unintended Pregnancies Using SARC by region"
@@
 
-We also created violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4.
+We also created violin plots using the categorical usage variables for each birth control method as shown in @figure4.
 It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
@@
 ```{r}
-#| label: exploratory-result4
+#| message: false
+#| label: figure4
 #| fig-cap: "Plot 5: Violin Plot Depicitng Usage of SARc Per Region"
@@
 
-Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @exploratory-result5.
+Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @figure5 - @figure10.
 The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
@@
 ```{r}
-#| label: exploratory-result5
+#| message: false
+#| label: figure5-figure10
 #| fig-cap: "Predictor Variable Distributions"
 #| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
@@
 ## Basic statistical analysis
-We completed a simple linear regression analysis with each of the 7 predictors shown in @statistical-analysis1.
+We completed a simple linear regression analysis with each of the 7 predictors shown in @table2.
 The predictor with the strongest positive relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
@@
 ```{r}
-#| label: statistical-analysis1
+#| message: false
+#| label: table2
 #| fig-cap: "Table 2: Strongest Predictor: sarc_standard"
@@
 The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
-These associations can be seen in @statistical-analysis2.
+These associations can be seen in @table3.
 ```{r}
-#| label: statistical-analysis2
+#| message: false
+#| label: table3
 #| fig-cap: "Table 3: Regional Predictor"
@@
 
-We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4.
+We created one model with all 5 predictors using multiple linear regression, shown in @table4.
 This model shows the strongest predictor is the number of women using short acting reversible contraception.
@@
 ```{r}
-#| label: statistical-analysis4
+#| message: false
+#| label: table4
 #| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
@@
 We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion.
-@statistical-analysis5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
+@table5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
 The model with the population size using SARCs is the only single predictor model that performs better than the null.
@@
 ```{r}
-#| label: statistical-analysis5
+#| message: false
+#| label: table5
 #| fig-cap: "Table 5: Performance Metric Comparisons"
@@
 We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces.
-The cross-validated results, shown in @statistical-analysis6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
+The cross-validated results, shown in @table6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
 The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319.
@@
 ```{r}
-#| label: statistical-analysis6
+#| message: false
+#| label: table6
 #| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
@@
 Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance.
-As seen in @statistical-analysis7, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
+As seen in @figure11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
 The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
@@
 ```{r}
-#| label: statistical-analysis7
+#| message: false
+#| label: figure11
 #| fig-cap: "Figure 3: Evaluating Final Model Performance on Test Data"
@@
 We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data.
-The figure shown in @statistical-analysis8 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
+The figure shown in @figure12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
 The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction.
@@
 ```{r}
-#| label: statistical-analysis8
+#| message: false
+#| label: figure12
 #| fig-cap: "Figure 4: Evaluation of Uncertainty in Full Model Predictions"
@@
 This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting.
-As seen in @statistical-analysis9, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
+As seen in @table7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
 The `sarc_standard` predictor still has the strongest relationship with the outcome.
@@
 ```{r}
-#| label: statistical-analysis9
+#| message: false
+#| label: table7
 #| fig-cap: "Table 7: LASSO Regression Results"
@@
 ```
-It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @statistical-analysis10.
+It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @figure13.
 As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
@@
 ```{r}
-#| label: statistical-analysis10
+#| message: false
+#| label: figure13
 #| fig-cap: "Figure 5: Parameter Tuning for LASSO Regression"
@@
 Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
-The plot of the predicted versus observed values shown in @statistical-analysis11 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
+The plot of the predicted versus observed values shown in @figure14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
 ```{r}
-#| label: statistical-analysis11
+#| message: false
+#| label: figure14
 #| fig-cap: "Predicted vs Observed Values in Random Forest Model"
@@
 Again, it is crucial to tune hyperparameters associate with machine learning models.
-Based on the red line displayed in @statistical-analysis12 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
+Based on the red line displayed in @figure15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
 This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
@@
 ```{r}
-#| label: statistical-analysis12
+#| message: false
+#| label: figure15
 #| fig-cap: "Tuning Parameters for Random Forest Model"
@@
 _Summarize what you did, what you found and what it means._
+critically think about each model (same number of predictors so we recommend random forest) all not great
 

### 2024-04-16 products/manuscript/Manuscript.qmd
Project 5 Updates
Figure caption update still looked incorrect when rendered, so I tried again.

@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @table1.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @table-1.
 ```{r}
 #| message: false
-#| label: table1
+#| label: table-1
 #| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
@@
 We created a multitude of exploratory plots to understand our data better.
-The first exploratory result shown in @figure1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
+The first exploratory result shown in @figure-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
 The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
@@
 #| message: false
-#| label: figure1
+#| label: figure-1
 #| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"

### 2024-04-16 code/analysis-code/statistical-analysis.qmd
Part 5 Updates
Critically analyzed each model and chose the best one. Added a summary of findings with interpretation to the manuscript. Attempted to fix the table and figure captions again.

@@
 The predictors used in our models include total number of women using short acting reversible methods standardized by population size, region, percent of women currently married, percent of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy, and rate of maternal deaths due to abortion.
-We are considering adding another predictor using a political freedom index for each country.
 
@@
 It is important to consider appropriate metrics to measure model performance.
-We will use RMSE and MAE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent never married predictor, and one using the rate of maternal deaths due to abortion predictor.
+We will use RMSE and MAE at this point to compare 3 models: one with the number of women using short acting reversible contraception predictor, one using the percent currently married predictor, and one using the rate of maternal deaths due to abortion predictor.
 We chose these predictors because they have the strongest associations with the outcomes based on the single linear regression models shown above.
@@
 
-Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression.
+Since we have evidence that our model overfits the data, we want to explore our full model using Lasso regression.
 This regression analysis shrinks the coefficients towards the mean, which allows it to perform subset selection.
@@
 Compared to the previous model with an RMSE of 4.361363, an MAE of 3.386193, and an R-squared of 0.820600, the random forest model performs very similarly.
-The RMSE and MAE values are exactly the same beause the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
+The RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
 Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
@@
 ggsave(filename = figure_file5, plot=rf_tune_plot)
+
 ```
@@
 
-Choose final model with justification
+## Choosing a final model 
+After considering each of the 10 models we created throughout the statistical analysis process, the random forest model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods.
+The collinearity between the total rate of maternal deaths, rate of maternal deaths due to abortion, and rate of maternal deaths due to unintended pregnancies justified why they should not be included in the same model.
+We chose to remove the total maternal mortality predictor to avoid the issue of collinearity because the other two predictors in this category had lower RMSE and MAE performance metrics.
+The same reasoning explains why the percent of currently married women and the percent of never married women predictors should not be included in the same model.
+We chose to include the percent of currently married women in the final model because it had the strongest relationship with the outcome.
+The four most promising models were the single predictor models with population size using short acting reversible contraception, percent of currently married women, and rate of maternal deaths due to abortion, respectively, along with the full model with all five predictors.
+The single model with the region variable only performed poorly with an RMSE of 18.11857, so it was not considered in the final running for the best model.
+The models with percent of currently married women and rate of maternal deaths due to abortion had comparable RMSE scores of 10.175909 and 10.141275, respectively.
+Large improvement in performance metrics was observed when considering the model with population size using short acting reversible contraception based on an RMSE of 5.143319.
+The full model with all five predictors, including short acting reversible contraception usage, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies, performed the best with an RMSE of 4.361363.
+We felt it was important to include the region predictor in the final model even though it did not perform well in a model on its own because of the large differences in the coefficients for the region predictor in a linear model with the outcome.
+Latin America/the Caribbean and Asia have much larger effects on the outcome than Europe and Oceania, so we argue for the inclusion of this variable in the final model to capture this information.
+
+Once we decided that the full model with 5 predictors was the best choice, we used cross-validation with LASSO regression and random forest modeling to explore subset selection.
+The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the sarc_standard predictor.
+The smallest coefficient was assigned to the percent of currently married women predictor.
+The parameters were tuned for the LASSO regression, which increased our confidence in using a penalty of 0.1.
+The random forest model generated the same RMSE of 4.361363 as the LASSO regression model, but it has a higher R-squared value of 0.8369754.
+Since the random forest model explains 83.69% of the variation between the predicted and observed outcomes compared to 82.06% explained by the LASSO regression model, it is slightly better at predicting the outcome.
+After tuning the parameters for this model, the number of randomly selected predictors that produces the lowest RMSE is 4, not 5.
+While we cannot be sure which predictor the model is suggesting to remove, we hypothesize that it is the percent of currently married women based on the information from the LASSO regression.
+However, the improvement in RMSE when using 4 predictors instead of 5 is minimal, and the minimal node size variable is much more impactful on the RMSE of the model.
+Based on all of this information, we suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model.
+While this model has a slightly lower R-squared value compared to the random forest model, the evidence of overfitting is so strong that this difference is negligible.
+Cross-validation of the model with all 5 predictors showed that predictions from testing data deviated from the expected relationship with larger variance than the predictions from the training data.
+Since the model did not perform well when used on test data, we conclude that there is overfitting in our model.
+Unfortunately, it is likely that any of these models are not optimal choices for predicting our outcome based on the poor performance of our model on testing data.
+While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.

### 2024-04-16 products/manuscript/Manuscript.qmd
Part 5 Updates
Critically analyzed each model and chose the best one. Added a summary of findings with interpretation to the manuscript. Attempted to fix the table and figure captions again.

--- 2024-04-16 products/manuscript/Manuscript.qmd
@@
 The Oceania and Europe region had weaker associations.
-Marriage status (both currently and never married) was inversely associated with unintended pregnancy rates.
-This means that as the percent of women who are currently married or never married increases, the unintended preganacy rate decreases.
-Maternal mortality rates showed a positive correlation with the outcoe.
+Marriage status, both currently and never married, was inversely associated with unintended pregnancy rates.
+This means that as the percent of women who are currently married or never married increases, the unintended pregnancy rate decreases.
+Maternal mortality rates showed a positive correlation with the outcome.
 
@@
 
-Overall, this study highlights the influence of regional variations, SARC use, and marital status on unintended pregnancy rates in low- and middle-income countries.
+Overall, this study highlights the influence of regional variations, short acting reversible contraception use, and marital status on unintended pregnancy rates in low- and middle-income countries.
 The findings suggest a need for further investigation into the factors influencing maternal mortality rates and unintended pregnancies.
@@
 Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.
-Lastly, we examined the AIC scores for the two best performing models since there is strong evidence for overfitting.
-Even though cross-validation is a better approach for answering the question of how well a model predicted values in unseen data because it actually resamples the data, the concern of overfitting justifies finding the AIC value in this case.
-This measure still evaluates the trade-off between good fit to the data, shown by a low cost function, and model complexity based on the number of parameters.
 
@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @table1.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
 ```{r}
 #| message: false
-#| label: table1
+#| label: tbl-1
 #| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
@@
 We created a multitude of exploratory plots to understand our data better.
-The first exploratory result shown in @figure1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
+The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods.
 The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
@@
 #| message: false
-#| label: figure1
+#| label: fig-1
 #| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
@@
 Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @figure5 - @figure10.
-The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes.
-about 35 countries.
+The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries.
 Latin America/the Caribbean accounts for about 25 countries.
@@
 The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers.
-We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.
+We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region.
 The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
@@
 The distribution of maternal mortality rate from abortions is also highly skewed to the right.
-The outliers are accounted for by Gambia and Mauritania in this predictor.
+The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
 ```{r}
@@
 knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))
+
 ```
@@
 The model with the population size using SARCs is the only single predictor model that performs better than the null.
-The full model with all 5 predictors peforms the best with the lowest RMSE and MAE values.
+The full model with all 5 predictors performs the best with the lowest RMSE and MAE values.
 ```{r}
@@
 
+As seen in the statistical analysis file, we created 10 different model options with unique combinations of our 7 predictors.
+After considering effect sizes and performance metrics for each model, we have decided that the LASSO regression model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods.
+After considering each of the 10 models we created throughout the statistical analysis process, the random forest model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods.
+We used cross-validation with LASSO regression and random forest modeling to explore subset selection of our full model with all 5 predictors.
+The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the sarc_standard predictor.
+The smallest coefficient was assigned to the percent of currently married women predictor.
+We suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model.
+While this model has a slightly lower R-squared value compared to the random forest model, the evidence of overfitting is so strong that this difference is negligible.
+Cross-validation of the model with all 5 predictors showed that predictions from testing data deviated from the expected relationship with larger variance than the predictions from the training data.
+Since the model did not perform well when used on test data, we conclude that there is overfitting in our model.
+Unfortunately, it is likely that any of these models are not optimal choices for predicting our outcome based on the poor performance of our model on testing data.
+While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.
+
 {{< pagebreak >}}
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
-critically think about each model (same number of predictors so we recommend random forest) all not great
+Our exploratory data analysis allowed us to examine the distribution of our outcome of interest, percentage of unintended pregnancies due to short acting reversible contraception, and the predictors used to describe the outcome.
+Most low- and middle-income countries were from the African region, followed by Asia and Latin America/the Caribbean.
+The Latin America/the Caribbean region experienced the highest percentage of unintended pregnancies followed by Europe and Asia.
+This association is likely explained by the fact that women in these regions have better access to short acting reversible contraception methods because we would have expected the African region to experience the highest rate of our outcome based on the number of countries included from that region.
+After standardizing the number of women who use short acting reversible contraception based on total population size, it is clear the women in the Latin America/the Caribbean region use this contraception method at the highest rates, followed my the Asian region.
+Further analysis of the region predictor in a linear model with the outcome confirmed this finding by showing that Latin America/the Caribbean has the strongest association with the outcome, followed by Asia.
+Asia accounted for the highest percentage of currently married women.
+The highest rate of maternal mortality due to abortion and highest rate of maternal mortality due to unintended pregnancies are both accounted for by the African region.
+This cannot be explained by having the highest number of reproductive age women because the Asian region has 6 times the number of reproductive age women compared to the African region.
+This discrepancy is likely due to poorer access to maternal healthcare in Africa compared to Asia.
+
+Our full statistical analysis of the final model with 5 predictors included the standardized number of women using short acting reversible methods, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors.
+Cross-validation using LASSO regression and random forest modeling techniques did not result in an increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data.
+Based on the model using LASSO regression, the number of women using short acting reversible contraception is the strongest predictor of the outcome, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place.
+The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome.
+The percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors are all negatively associated with the outcome in this model.
+This implies that the percentage of unintended pregnancies decreases when a higher percentage of women are currently married and when the country experiences a higher maternal mortality rate based on abortion and unintended pregnancies.
+However, these effects are so small that it is likely these predictors do not have a real effect on the outcome.
+This analysis was successful in determining that region does have a significant effect on the percentage of unintended pregnancies due to short acting reversible methods with the Latin America/the Caribbean region experiencing the outcome at the highest rates.
 

### 2024-04-16 products/manuscript/Manuscript.qmd
Figure Labels
Fixed all figure labels to make sure they render correctly now.

@@
 #| label: fig-1
-#| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
+#| fig-cap: "Number of Countries With Percent Unintended Pregnancy Using SARC"
 #| echo: FALSE
@@
 Each of these percentages are positively associated with number of people in its respective classification group that it represents.
-The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @figure2.
+The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
 ```{r}
 #| message: false
-#| label: figure2
-#| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
+#| label: fig-2
+#| fig-cap: "Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
 #| echo: FALSE
@@
 
-We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @figure3 .
+We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 .
 The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
@@
 #| message: false
-#| label: figure3
-#| fig-cap: "Figure 2: Graph Depicting Percent Unintended Pregnancies Using SARC by region"
+#| label: fig-3
+#| fig-cap: "Graph Depicting Percent Unintended Pregnancies Using SARC by region"
 #| echo: FALSE
@@
 
-We also created violin plots using the categorical usage variables for each birth control method as shown in @figure4.
+We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4.
 It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
@@
 #| message: false
-#| label: figure4
-#| fig-cap: "Plot 5: Violin Plot Depicitng Usage of SARc Per Region"
+#| label: fig-4
+#| fig-cap: "Violin Plot Depicitng Usage of SARc Per Region"
 #| echo: FALSE
@@
 
-Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @figure5 - @figure10.
+Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10.
 The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries.
@@
 #| message: false
-#| label: figure5-figure10
+#| label: fig-5 - fig 10
 #| fig-cap: "Predictor Variable Distributions"
@@
 ## Basic statistical analysis
-We completed a simple linear regression analysis with each of the 7 predictors shown in @table2.
+We completed a simple linear regression analysis with each of the 7 predictors shown in @tbl-2.
 The predictor with the strongest positive relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).
@@
 #| message: false
-#| label: table2
-#| fig-cap: "Table 2: Strongest Predictor: sarc_standard"
+#| label: tbl-2
+#| fig-cap: "Strongest Predictor: sarc_standard"
 #| echo: FALSE
@@
 The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
-These associations can be seen in @table3.
+These associations can be seen in @tbl-3.
 ```{r}
 #| message: false
-#| label: table3
-#| fig-cap: "Table 3: Regional Predictor"
+#| label: tbl-3
+#| fig-cap: "Regional Predictor"
 #| echo: FALSE
@@
 ```
-Details about the 9 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.
+Details about the 10 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.
 
@@
 
-We created one model with all 5 predictors using multiple linear regression, shown in @table4.
+We created one model with all 5 predictors using multiple linear regression, shown in @tbl-4.
 This model shows the strongest predictor is the number of women using short acting reversible contraception.
@@
 #| message: false
-#| label: table4
-#| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
+#| label: tbl-4
+#| fig-cap: "Multiple Linear Regression with 5 Predictors"
 #| echo: FALSE
@@
 We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion.
-@table5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
+@tbl-5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
 The model with the population size using SARCs is the only single predictor model that performs better than the null.
@@
 #| message: false
-#| label: table5
-#| fig-cap: "Table 5: Performance Metric Comparisons"
+#| label: tbl-5
+#| fig-cap: "Performance Metric Comparisons"
 #| echo: FALSE
@@
 We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces.
-The cross-validated results, shown in @table6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
+The cross-validated results, shown in @tbl-6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
 The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319.
@@
 #| message: false
-#| label: table6
-#| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
+#| label: tbl-6
+#| fig-cap: "Cross Validation Performance Metric Comparisons"
 #| echo: FALSE
@@
 Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance.
-As seen in @figure11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
+As seen in @fig-11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
 The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
@@
 #| message: false
-#| label: figure11
-#| fig-cap: "Figure 3: Evaluating Final Model Performance on Test Data"
+#| label: fig-11
+#| fig-cap: "Evaluating Final Model Performance on Test Data"
 #| echo: FALSE
@@
 We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data.
-The figure shown in @figure12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
+The figure shown in @fig-12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
 The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction.
@@
 #| message: false
-#| label: figure12
-#| fig-cap: "Figure 4: Evaluation of Uncertainty in Full Model Predictions"
+#| label: fig-12
+#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
 #| echo: FALSE
@@
 This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting.
-As seen in @table7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
+As seen in @tbl-7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
 The `sarc_standard` predictor still has the strongest relationship with the outcome.
@@
 #| message: false
-#| label: table7
-#| fig-cap: "Table 7: LASSO Regression Results"
+#| label: tbl-7
+#| fig-cap: "LASSO Regression Results"
 #| echo: FALSE
@@
 ```
-It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @figure13.
+It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-13.
 As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
@@
 #| message: false
-#| label: figure13
-#| fig-cap: "Figure 5: Parameter Tuning for LASSO Regression"
+#| label: fig-13
+#| fig-cap: "Parameter Tuning for LASSO Regression"
 #| echo: FALSE
@@
 Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
-The plot of the predicted versus observed values shown in @figure14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
+The plot of the predicted versus observed values shown in @fig-14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
 ```{r}
 #| message: false
-#| label: figure14
+#| label: fig-14
 #| fig-cap: "Predicted vs Observed Values in Random Forest Model"
@@
 Again, it is crucial to tune hyperparameters associate with machine learning models.
-Based on the red line displayed in @figure15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
+Based on the red line displayed in @fig-15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
 This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
@@
 #| message: false
-#| label: figure15
+#| label: fig-15
 #| fig-cap: "Tuning Parameters for Random Forest Model"

### 2024-04-17 code/analysis-code/statistical-analysis.qmd
Part 5 Updates
Added figure captions for both statistical analysis sections. Converted tibbles to organized tables for the rendered manuscript. Added a legend to the bootstrapping graph.

@@
 
+# create a professional format of the table to add to the manuscript 
+kable_lmtable1 <- kable(lmtable1)
+
 # save fit results table  
 table_file1 = here("results", "tables", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+saveRDS(kable_lmtable1, file = table_file1)
 ```
@@
 
+# create a professional format of the table to add to the manuscript 
+kable_lmtable2 <- kable(lmtable2)
+
 # save fit results table  
 table_file2 = here("results", "tables", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
+saveRDS(kable_lmtable2, file = table_file2)
 
@@
 
+# create a professional format of the table to add to the manuscript 
+kable_lmtable8 <- kable(lmtable8)
+
 # save fit results table  
 table_file8 = here("results", "tables", "resulttable8.rds")
-saveRDS(lmtable8, file = table_file8)
+saveRDS(kable_lmtable8, file = table_file8)
 ```
@@
 figure2 <- ggplot(finalbootstrap, aes(x = observed, y = point_estimates)) +
-              geom_point(color = "black") +  
-              geom_point(aes(y = medians), color = "coral") + 
+              geom_point(aes(shape = "Point Estimates", color = "black")) +  
+              geom_point(aes(y = medians, shape = "Medians",color = "coral")) + 
               geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.1, color = "blue") + 
@@
               ylim(0, 50) + # limit the y-axis values
-              labs(x = "Observed Values", y = "Predicted Values", title = "Uncertainty in Predictions with Final Model")
+              labs(x = "Observed Values", y = "Predicted Values", title = "Uncertainty in Predictions with Final Model") +
+              scale_color_manual(values = c("black" = "black", "coral" = "coral", "blue" = "blue"),
+                                 labels = c("Point Estimates", "Medians", "Error Bars"))
 figure2
@@
 This figure shows that 95% confidence intervals, shown as dark blue lines, for the predictions that do not fall on the expected 45 degree line do not intersect with the line either.
+The point estimates, represented as black triangles, have decent overlap with the medians, represented as coral circles, which is a sign of good agreement between predicted and observed values.
 The dark blue lines show that the confidence intervals are narrow, which increases the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions.
+Unfortunately, the model is inaccurate, but it is inaccurate without much variation.
 
@@
 
+# create a professional format of the table to add to the manuscript 
+kable_lmtable11 <- kable(lmtable11)
+
 # save fit results table  
 table_file16 = here("results", "tables", "resulttable16.rds")
-saveRDS(lmtable11, file = table_file16)
+saveRDS(kable_lmtable11, file = table_file16)
 ```
@@
 While we cannot be sure which predictor the model is suggesting to remove, we hypothesize that it is the percent of currently married women based on the information from the LASSO regression.
-However, the improvement in RMSE when using 4 predictors instead of 5 is minimal, and the minimal node size variable is much more impactful on the RMSE of the model.
+However, the improvement in RMSE when using 4 predictors instead of 5 is minimal, but it is interesting that the graph suggests using a minimal node size of 4 if the model contains all 5 predictors.
 Based on all of this information, we suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model.

### 2024-04-17 products/manuscript/Manuscript.qmd
Part 5 Updates
Added figure captions for both statistical analysis sections. Converted tibbles to organized tables for the rendered manuscript. Added a legend to the bootstrapping graph.

@@
 To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions.
-This process is similar to cross-validation because it relies on resamling the training data.
+This process is similar to cross-validation because it relies on resampling the training data.
 Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance.
@@
 We added a random forest model to attempt to improve model performance.
-We used another parameter tuning grid to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
+We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
 We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so.
@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1. [insert explanation here]
 ```{r}
@@
 #| label: tbl-1
-#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
+#| tbl-cap: "Summary Table Of Numeric Variables"
 #| echo: FALSE
@@
 #| label: fig-1
-#| fig-cap: "Number of Countries With Percent Unintended Pregnancy Using SARC"
+#| fig-cap: "[insert explanation here]"
 #| echo: FALSE
@@
 #| label: fig-2
-#| fig-cap: "Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
+#| fig-cap: "insert explanation here"
 #| echo: FALSE
@@
 #| label: fig-3
-#| fig-cap: "Graph Depicting Percent Unintended Pregnancies Using SARC by region"
+#| fig-cap: "insert explanation here"
 #| echo: FALSE
@@
 #| label: fig-4
-#| fig-cap: "Violin Plot Depicitng Usage of SARc Per Region"
+#| fig-cap: "insert explanation here"
 #| echo: FALSE
@@
 #| label: fig-5 - fig 10
-#| fig-cap: "Predictor Variable Distributions"
+#| fig-cap: "insert explanation here"
 #| echo: FALSE
@@
 #| label: tbl-2
-#| fig-cap: "Strongest Predictor: sarc_standard"
+#| fig-cap: "This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception.
+This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome.
+The coefficient of 87.65998 is the largest single coefficient of any predictor that we studied, which increases our confidence in including it as a predictor in our final model."
 #| echo: FALSE
@@
 #| label: tbl-3
-#| fig-cap: "Regional Predictor"
+#| fig-cap: "In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia.
+Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome.
+This model chose the African region as the reference group, which means that it has the weakest association with the outcome compared to the other 4 regions."
 #| echo: FALSE
@@
 #| label: tbl-4
-#| fig-cap: "Multiple Linear Regression with 5 Predictors"
+#| fig-cap: "The multiple linear regression model with all 5 predictors confirms that the `sarc_standard` predictor has the strongest effect on the outcome, but the effect size has decreased slightly.
+The region variable is the second most impactful predictor, but there is a shift in which region has the strongest effect on the outcome.
+Africa still serves as the control region, but Europe has the largest effect size in this model followed by Latin America/the Caribbean.
+Oceania still has the weakest effect, and the effect is negative in this model compared to the positive effect seen in the simple linear regression model with the region predictor.
+The percent of currently married women, rate of maternal deaths due to uninteded pregnancies, and rate of maternal deaths due to abortion are very weak predictors with miniscule coefficients."
 #| echo: FALSE
@@
 #| label: tbl-5
-#| fig-cap: "Performance Metric Comparisons"
+#| fig-cap: "The performance metric comparisons clearly highlight the full model with 5 predictors as the best model with the lowest RMSE of 4.361363 and lowest MAE of 3.386193.
+The model with percent of currently married women and the model with the rate of maternal mortality due to abortion only perform slightly better than the null model, which just predicts the average outcome.
+These models are extremely poor.
+The best single predictor model is the one with the standard population size using SARCs with an RMSE of 5.143319 and an MAE of 3.953767."
 #| echo: FALSE
@@
 #| label: tbl-6
-#| fig-cap: "Cross Validation Performance Metric Comparisons"
+#| fig-cap: "Cross-validated (CV) performance metric comparisons were only completed on two models: the single predictor model with the `sarc_standard` variable and the full model with all five predictors because these two models performed the best in the first analysis.
+The RMSE metrics are closer using the CV technique with a differencec of 0.42 between the two models.
+The standard error of the full predictor model is slightly higher, but the R-squared value is also slightly higher for the full predictor model.
+These metrics confirm that the full predictor model is the best choice compared to single predictor models."
 #| echo: FALSE
@@
 #| label: fig-11
-#| fig-cap: "Evaluating Final Model Performance on Test Data"
+#| fig-cap: "Evaluating the final model performance on test data reveals that the model is strongly overfitting the data.
+The testing model predictions are shown in coral, and the variation in these observations is much higher compared to the training model predictions, shown in blue.
+Ideally, there would be complete overlap between the testing and training data predictions following the black dotted line that represent the ideal relationship between observed and predicted values.
+The majority of testing data predictions do not overlap with the training data prediction in this case, and the testing data predictions are father away from the preferred relationship shown by the black dotted line."
 #| echo: FALSE
@@
 #| label: fig-12
-#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
+#| fig-cap: "Evaluation of uncertainty in the full model predictions reveals decent overlap between the point estimates, shown in black triangles, and the median prediction observation, shown in coral circles.
+The error bars, shown in blue, represent 95% confidence intervals of the median estimates, and most of the confidence intervals contain the point estimates.
+This provides confidence in the certainty of the model's predictions, but this is not very meaninful considering how the full model overfits the data."
 #| echo: FALSE
@@
 #| label: tbl-7
-#| fig-cap: "LASSO Regression Results"
+#| fig-cap: "LASSO regression results confirm the findings of the multiple linear regression model, but each of the coefficients has been shrunk as a result of the penalty parameter of 0.1.
+The `sarc_standard` predictor is still the most strongly related to the outcome, followed by the European region.
+The LASSO penaltty decreased the effect size of the Oceania region even further, and the smallest coefficient is now assigned to the percent of currently married women, which is a shift from the multiple linear regression model.
+The rates of maternal mortality due to unintended pregnancy and abortion are very weak predictors in this model as well."
 #| echo: FALSE
@@
 #| label: fig-13
-#| fig-cap: "Parameter Tuning for LASSO Regression"
+#| fig-cap: "Parameter tuning for the LASSO regression reveals that the parameter should be smaller than 1 to ensure the lowest RMSE performance metric value.
+The RMSE value increases exponentially when the penalty parameter exceeds 1, so this provides confidence in our choice of using 0.1 as the penalty parameter in the prevoiusly displayed LASSO regression model.
+There is no evidence to use a smaller penalty parameter than 0.1 because it does not result in a lower performance metric."
 #| echo: FALSE
@@
 #| label: fig-14
-#| fig-cap: "Predicted vs Observed Values in Random Forest Model"
+#| fig-cap: "The graph of predicted versus observed values in the random forest model shows less deviation from the expected relationship visualized by the black dotted line in comparison to the multiple linear regression model displayed previously.
+While there are still several outliers as the predicted values increase past a value of 20, the observations hover around the expected relationship with conistency for values less than 10.
+While it is still likely that the random forest model is overfitting because it uses the same predictors, it appears that this model has higher predictive accuracy."
 #| echo: FALSE
@@
 This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
-However, there is a minimal increase in RMSE when all five models are included in the model with a node size of 1.
+However, there is a minimal increase in RMSE when all five predictors are included in the model, but the results suggest that a full model should have a node size of 4 to achieve the lowest RMSE.
 ```{r}
@@
 #| label: fig-15
-#| fig-cap: "Tuning Parameters for Random Forest Model"
+#| fig-cap: "After tuning two parameters for the random forest model, it is clear that the best model based on the lowest RMSE performance metric value has a minimal node size of 1 for 4 out of the 5 models.
+This graph suggests that the final model should contain 4 predictors to achieve the best performance metric, but it does not provide information on which parameter should be dropped from the model.
+Interestingly, the best model with all 5 predictors has a minimal node size of 4."
 #| echo: FALSE
@@
 The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome.
+This finding implies that these three regions have the best access to SARCs because they also have the highest standardized population size using this form of contraception in the first place.
+Since the African region was always selected as the control for the model, it is likely that this region has the lowest access to SARCs.
 The percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors are all negatively associated with the outcome in this model.

### 2024-04-18 code/eda-code/eda.qmd
Added to conclusion.

@@
   geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
+  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
 plot(plot_4)

### 2024-04-18 products/manuscript/Manuscript.qmd
Added to conclusion.

@@
 _Discuss what you perceive as strengths and limitations of your analysis._
+This study relied on secondary data from the Guttmacher Institute's Adding It Up project.
+This means that we were limited by the data that was available and unable to collect our own data.
+Also, the study was cross-sectional, which means that it only looked at data from one point in time.
+This makes it difficult to establish causal relationships between the variables.
+
+Despite these limitations, our study provides valuable insights into the factors influencing unintended pregnancy rates in low- and middle-income countries.
+The findings suggest that a multifaceted approach is needed to address unintended pregnancy.
+This approach should include increasing access to SARC and LARC methods, improving education and counseling on contraception use, and addressing regional disparities in access to healthcare services and family planning programs.
 
@@
 _What are the main take-home messages?_
+This study investigated the association between unintended pregnancy rates and short-acting reversible contraception (SARC) use across low- and middle-income countries.
+The findings suggest that SARC use, regional variations, and marital status are all important factors influencing unintended pregnancy rates.
+
+The study found that SARC use was the strongest predictor of unintended pregnancy rates.
+This suggests that women who rely on SARC methods are more likely to experience unintended pregnancies compared to those who use other contraceptive methods.
+There are several possible explanations for this finding.
+First, SARC methods, such as pills and patches require more frequent user adherence than long-acting reversible contraceptives (LARCs) like implants.
+Perfect use of SARC methods is essential for their effectiveness, and even small lapses in use can increase the risk of unintended pregnancy.
+Second, SARC methods may not be as effective for women who have difficulty remembering to take their medication daily or who experience side effects from hormonal contraception.
+
+The study also found that regional variations were significantly associated with unintended pregnancy rates.
+The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia.
+The Oceania and Europe region had weaker associations.
+These findings suggest that there are significant contextual factors that influence unintended pregnancy rates across different regions.
+These factors may include access to healthcare services, legal barriers to access, and cultural attitudes towards contraception.
+Additionally, regions with cultural or religious restrictions on contraception use may also have higher rates of unintended pregnancy.
+
+The study found that marital status was inversely associated with unintended pregnancy rates.
+We found that as the percent of women who are currently married increases, the unintended pregnancy rate decreases.
+There are some reasonable explanations for this finding.
+First, married women may have better access to healthcare services and family planning programs than unmarried women.
+Secondly, married couples may be more likely to plan their pregnancies than unmarried couples.
+Third, unmarried women may be more likely to have multiple sexual partners who could be opposed to using contraception.
+
+Future research directions could include conducting qualitative studies to better understand the reasons why women choose SARC methods and the challenges they face in using them effectively.
+Additionally, we suggest that a longitudinal study should be conducted to track changes in unintended pregnancy rates over time and to evaluate the effectiveness of different interventions aimed at reducing unintended pregnancy.
 

### 2024-04-19 code/eda-code/eda.qmd
Removed distribution graphs from manuscript. Edited region graph

@@
 # explore percent unintended pregnancies using short acting reversible methods by region
-plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
+plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
   geom_point() + 
+  geom_jitter(aes(colour = region)) +
   geom_smooth(method='lm') + 
@@
 figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
-ggsave(filename = figure_file, plot=p23)
+ggsave(filename = figure_file, plot=plot_7)
 

### 2024-04-19 products/manuscript/Manuscript.qmd
Removed distribution graphs from manuscript. Edited region graph

@@
 The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
-```{r}
-#| message: false
-#| label: fig-5 - fig 10
-#| fig-cap: "insert explanation here"
-#| echo: FALSE
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
-knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))
-
-```
 

### 2024-04-19 products/manuscript/supplement/Supplementary-Material.qmd
Added labels to figures in manuscript, switched axis labels.
Removed distribution graphs from manuscript. Edited region graph

@@
 This information provides guidance on our thought processes behind the scenes and explains why we chose a single outcome to study.
+The plots below are numbered based off of order of creation, they are not numbered in order.
 
@@
 P5: This histogram shows that more countries report a higher percentage of unintended pregnancies among women who do not use any contraception, but it varies widely by country.
+
+We also needed to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`.
+Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers.
+```{r}
+# explore the region predictor
+plot_6 <- mydata %>% ggplot(aes(x=region)) + 
+  geom_bar() +
+  ggtitle("Countries in Each Region") + 
+  xlab("Region") + ylab("Number of Countries")
+plot(plot_6)
+figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
+ggsave(filename = figure_file, plot=plot_6)
+
+# explore the current marriage predictor
+plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Currently Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(plot_7)
+figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
+ggsave(filename = figure_file, plot=plot_7)
+
+# explore the never married predictor
+plot_8 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
+  geom_histogram() +
+  ggtitle("Percent of Women Never Married") + 
+  xlab("Percentage") + ylab("Number of Countries")
+plot(plot_8)
+figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
+ggsave(filename = figure_file, plot=plot_8)
+
+# explore the total maternal mortality rate predictor
+plot_9 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
+  geom_histogram() +
+  ggtitle("Total Rate of Maternal Deaths") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(plot_9)
+figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
+ggsave(filename = figure_file, plot=plot_9)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths > 1000)
+
+# explore the maternal mortality rate from unintended pregnancies predictor
+plot_10 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(plot_10)
+figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
+ggsave(filename = figure_file, plot=plot_10)
+
+# explore outliers in maternal mortality from unintended pregnancies
+mydata %>% filter(rate_matdeaths_upreg > 6000)
+
+# explore the maternal mortality rate from abortions
+plot_11 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
+  geom_histogram() +
+  ggtitle("Rate of Maternal Deaths from Abortions") + 
+  xlab("Maternal Deaths") + ylab("Number of Countries")
+plot(plot_11)
+figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
+ggsave(filename = figure_file, plot=plot_11)
+
+# explore outliers in maternal mortality from abortions
+mydata %>% filter(rate_matdeaths_abs > 500)
+
+```
 

### 2024-04-19 code/eda-code/eda.qmd
Part 5 Updates
Contributed to strengths/limitations and added labels to the summary tables so there are no variable names.

@@
 
+# Modify region variable type for proper labeling 
+selected_variables$region <- factor(selected_variables$region, labels = c("Africa", "Asia", "Europe", "Latin America/the Caribbean", "Oceania"))
+
 # Create table 
@@
     by = region,
-    label = list(wra ~ "Number of women of reproductive age, 15-49"),
+    label = list(wra ~ "Number of women of reproductive age, 15-49",
+                 starts_with("pct_upreg_sarc" ) ~ "Unintended Pregnancies due to SARC Usage (%)",
+                 starts_with("pct_currentlymarried" ) ~ "Currently Married Women (%)",
+                 starts_with("pct_nevermarried" ) ~ "Never Married Women (%)",
+                 starts_with("rate_matdeaths") ~ "Rate of Maternal Deaths (per 100,000 live births)",
+                 starts_with("rate_matdeaths_upreg") ~ "Rate of Maternal Deaths due to Unintended Pregnancies",
+                 starts_with("rate_matdeaths_abs") ~ "Rate of Maternal Deaths due to Abortions"
+                 ),
     type = all_continuous() ~ "continuous2",
@@
 modify_header(label ~ "**Predictor Variable**") %>%
-modify_caption("**Table 1. Predictor Variables**") %>%
+  modify_header(
+    starts_with("region") ~ "**Region (1-Africa, 2-Asia, 3-Europe, 4-Latin America/the Caribbean, 5-Oceania)**"
+  ) %>%
+modify_caption("**Table 1. Summary Table**") %>%
 modify_spanning_header(all_stat_cols() ~ "Region")

### 2024-04-19 products/manuscript/Manuscript.qmd
Part 5 Updates
Contributed to strengths/limitations and added labels to the summary tables so there are no variable names.

@@
 #| label: tbl-1
-#| tbl-cap: "Summary Table Of Numeric Variables"
+#| tbl-cap: "Summary Table Of Continuous Variables by Region"
 #| echo: FALSE
@@
 
-Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10.
+Lastly, we explored the predictor variables that will be used in our models.
 The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries.
@@
 Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance.
-As seen in @fig-11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
+As seen in @fig-6, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line.
 The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line.
@@
 #| message: false
-#| label: fig-11
+#| label: fig-6
 #| fig-cap: "Evaluating the final model performance on test data reveals that the model is strongly overfitting the data.
@@
 We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data.
-The figure shown in @fig-12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
+The figure shown in @fig-7 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model.
 The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction.
@@
 #| message: false
-#| label: fig-12
+#| label: fig-7
 #| fig-cap: "Evaluation of uncertainty in the full model predictions reveals decent overlap between the point estimates, shown in black triangles, and the median prediction observation, shown in coral circles.
@@
 ```
-It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-13.
+It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-8.
 As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
@@
 #| message: false
-#| label: fig-13
+#| label: fig-8
 #| fig-cap: "Parameter tuning for the LASSO regression reveals that the parameter should be smaller than 1 to ensure the lowest RMSE performance metric value.
@@
 Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
-The plot of the predicted versus observed values shown in @fig-14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
+The plot of the predicted versus observed values shown in @fig-9 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.
 ```{r}
 #| message: false
-#| label: fig-14
+#| label: fig-9
 #| fig-cap: "The graph of predicted versus observed values in the random forest model shows less deviation from the expected relationship visualized by the black dotted line in comparison to the multiple linear regression model displayed previously.
@@
 Again, it is crucial to tune hyperparameters associate with machine learning models.
-Based on the red line displayed in @fig-15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
+Based on the red line displayed in @fig-10 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
 This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model.
@@
 #| message: false
-#| label: fig-15
+#| label: fig-10
 #| fig-cap: "After tuning two parameters for the random forest model, it is clear that the best model based on the lowest RMSE performance metric value has a minimal node size of 1 for 4 out of the 5 models.
@@
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
-This study relied on secondary data from the Guttmacher Institute's Adding It Up project.
-This means that we were limited by the data that was available and unable to collect our own data.
-Also, the study was cross-sectional, which means that it only looked at data from one point in time.
-This makes it difficult to establish causal relationships between the variables.
+There are several limitations for this analysis that must be addressed.
+First, this study relied on secondary data from the Guttmacher Institute's Adding It Up project.
+We were limited by the data that was available to us because we were unable to collect our own data.
+The study was cross-sectional, which means that it only looked at data from one point in time.
+This makes it difficult to establish causal relationships between the variables based on lack of information about temporality.
+The largest limitation of the study is the evidence of overfitting for our final model.
+While we had solid evidence from performance metrics for choosing this model, it does a fairly poor job of representing the data based on the cross-validated testing and training data predictions.
+Our recommendation is to explore more variables from the Adding It Up project to determine if there are better predictors for this outcome.
+Since we chose our predictors based on the literature, it is unlikely that there are other predictors that will more accurately predict the percentage of unintended pregnancies due to SARCs.
+Further research into the other contraceptive methods studied in the Adding It Up project, such as long acting reversible contraception, sterilization, and traditional methods, could provide clarity on the patterns that we discovered in this analysis.
+Due to limited time and experience, we were not able to examine all 4 of these outcomes in this project.
 
 Despite these limitations, our study provides valuable insights into the factors influencing unintended pregnancy rates in low- and middle-income countries.
-The findings suggest that a multifaceted approach is needed to address unintended pregnancy.
-This approach should include increasing access to SARC and LARC methods, improving education and counseling on contraception use, and addressing regional disparities in access to healthcare services and family planning programs.
+We utilize cutting edge machine learning models to test 10 different model options, cross-validate 3 of the best performing models, and tune parameters for LASSO regression and random forest modeling of our final model.
+These robust techniques allowed for a thorough investigation of predictors and outcomes.
+Our findings suggest that a multifaceted approach is needed to address unintended pregnancy.
+This approach should include increasing access to SARC methods, improving education and counseling on contraception use, and addressing regional disparities in access to healthcare services and family planning programs.
 
 ## Conclusions
-_What are the main take-home messages?_
 This study investigated the association between unintended pregnancy rates and short-acting reversible contraception (SARC) use across low- and middle-income countries.
@@
 # References
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
 

### 2024-04-19 products/manuscript/supplement/Supplementary-Material.qmd
Part 5 Updates
Contributed to strengths/limitations and added labels to the summary tables so there are no variable names.

@@
 
-Any additional discussion regarding the supplementary material/findings.
-
-{{< pagebreak >}}
-
-# References
+The most helpful part of this supplementary material is the information about the other potential outcomes that can be studied including the percentage of unintended pregnancies due to long-acting reversible methods, sterilization, traditional methods, and no birth control method.
+The initial exploration into these variables will provide future researchers with an initial idea of the distribution of these variables.
+The additional exploratory analysis included above provides insight on how we standardized our population predictor and dealt with missing data, which increases the validity of our study methods.
+The details about the methods we chose add confidence that the analysis performed in the statistical-analysis.qmd was the correct approach for our data.
+Choosing robust techniques like the LASSO regression and bootstrapping add evidence to our thorough investigation of the outcome and 7 predictors.

### 2024-04-19 products/manuscript/Manuscript.qmd
Part 5 Updates
Minor manuscript update.

@@
 ## Statistical analysis
-We began with a simple linear regression model with each of the 7 predictor variables to determine the association betwen each variable and the outcome.
+We began with a simple linear regression model with each of the 7 predictor variables to determine the association between each variable and the outcome.
 We explored the region variable further because our main goal is to understand the difference in outcome by region.
-We used ANOVA to determine the difference between regions, and we added Tukey's Honeslty Significant Difference test to examine pairwise associations.
 This led to discovering correlations between some of the predictors, which we examined through correlation coefficients.

### 2024-04-19 products/manuscript/Manuscript.qmd
Hopefully works! Pushed manuscript figure captions and worked on tbl err

@@
 #| label: fig-1
-#| fig-cap: "[insert explanation here]"
+#| fig-cap: "[A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods.
+The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs.]"
 #| echo: FALSE
@@
 #| label: fig-2
-#| fig-cap: "insert explanation here"
+#| fig-cap: "[A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies.
+In this case, a positive correlation is shown.
+It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)]"
 #| echo: FALSE
@@
 #| label: fig-3
-#| fig-cap: "insert explanation here"
+#| fig-cap: "[Scatter plots are used to see if there is a relationship between two sets of continuous data.
+In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies.
+The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa.]"
 #| echo: FALSE
@@
 #| label: fig-4
-#| fig-cap: "insert explanation here"
+#| fig-cap: "[Violin plots are used to visualize the distribution of data across different categories.
+In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage.
+This means that there is more variation in unintended pregnancy rates in counties with high SARC usage.]"
 #| echo: FALSE

### 2024-04-19 products/manuscript/Manuscript.qmd
summary table caption

@@
 #| label: tbl-1
-#| tbl-cap: "Summary Table Of Continuous Variables by Region"
+#| tbl-cap: "[A summary table was done to present data about the seven predictor variables chosen in a structured format.
+Summary tables enable us to compare between different categories or groups within the data set.
+This can reveal trends or relationships.
+From the summary table below, it would appear that] Latin America/the Caribbean has the highest percentage of unintended pregancies due to SARC usage."
 #| echo: FALSE

### 2024-04-20 products/manuscript/Manuscript.qmd
Added to  summary table caption and rendered again

@@
 This can reveal trends or relationships.
-From the summary table below, it would appear that] Latin America/the Caribbean has the highest percentage of unintended pregancies due to SARC usage."
+From the summary table below, it would appear that the region with the highest rate of unintended pregnancies due to SARC usage is Latin America/Caribbean at 24%, followed by Europe at 23%.
+Africa and Asia have a mean of 13% and 14% respectively, while Oceania appears to have the lowest rate at 6%.
+It appears that a higher percentage of women are currently married in Asia (66%) and Africa (60%) compared to Latin America/Caribbean (54%) and Oceania (60%).The average percentage of never married women across the 5 regions is 31%.
+Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.]"
 #| echo: FALSE

### 2024-04-25 reviews/Roth-project-review.qmd

---
title: Project Review Template 
author: Cassia Roth
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Global Distribution of Unintended Pregnancies and Contraceptive Usage

Name of project author(s): Taylor Glass and Arlyn Santiago

Name of project reviewer: Cassia Roth

# Specific project content evaluation

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

You do an excellent job in your general background information and your motivation for the project. The first half of the paragraph is a tiny bit vague in that the studies you outline don't include where they were conducted, which seems to matter for the motivation of your study (and the differences among regions that you pinpoint in your findings). You also combine all contraceptives with emergency contraceptives in the second half of the paragraph. These are very different things, unless your definition of SARC is ONLY emergency contraceptives. You might want to clarify this, as it undermines your overall discussion. Are you including EC in SARCs? What else is a SARC? You might have seen the greatest variation in Latin American regions because that region also has a large variation between countries in abortion legality (from completely illegal in all circumstances, Nicaragua, El Salvador, Honduras to fully legal until 24 weeks LMP in Colombia). Some discussion of abortion might be interesting here.

In fact, you might want to provide more detailed definitions and examples for all of the three categories of contraceptives (long-acting reversible and sterilization; SARC; and traditional). I would suggest your opening sentence omit the word "higher" since you don't know the relationship at the outset.

### Summary assessment

-   **strong contextualization and motivation**

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

In the question section, you very clearly and expertly outline your main question: Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region? You might want to clarify what you mean by significant. Do you mean statistically significant at the 0.05 level? You might also want to clarify what you mean by "among region." Do you mean between regions? Which regions?

### Summary assessment

-   **question/hypotheses somewhat explained**

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

You do a superb job of describing your overall data. The codebook is clear as is the data source. You also outline where you got it from.

### Summary assessment

-   **source and overall structure of data well explained**

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Wow, you've done a great job here! I was very easily able to follow all of your wrangling and exploratory methods. You consider many different alternatives and provide all meaningful results in an easy-to-follow manner. I think some of the methods section could be removed from the final manuscript and put in the supplemental file, for example the description of all performance metrics you performed (just state the final one and refer to supplemental for the background analysis you did to get there).

### Summary assessment

-   **essentially no weaknesses in wrangling and exploratory component**

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Your descriptions of your models is very clear and informative. From my understanding of these models, they make sense. I would have found it helpful if you wrote out your model in mathematical form. I'm also wondering what the literature has included for covariates in similar studies? Why you chose the model you did and why you included/excluded specific covariates given model performance is very clear, but sometimes we want to know more about relationships even if that reduces the model's predictive power. What have other studies included as covariates? Does your study differ?

### Summary assessment

-   **strong and reasonable analysis**

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The Tables and figures are great, although some of the tables were hard to read because they were not centered. I also had this problem with flextable, and I don't know how to fix it! There are some misspellings of standardized in Figures 2 and 3. The references to table 2 and 3 came out as unformatted in the word document I rendered (For these tables, you might want to use flextable). You also have multiple tables with the same numbers.

### Summary assessment

-   **results are very well presented**

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

You do a great job of discussing your findings. I was interested to see a bit more about the "so-what" findings. Women who use SARC have higher probability of experiencing unintended pregnancies than women who don't use SARC? Even women who use traditional or no methods? I wasn't clear on this because in your discussion you only mention LARC. But, more importantly, what does this mean for policy? Should we focus on LARCs instead of SARCs? Or should we legalize abortion?

### Summary assessment

-   **strong, complete and clear discussion**

## Further comments

I'm only interested in knowing what you think this means for policy and contraception, since I am interested in this topic!

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Everything was well structured and cleaned up. Great job. My only question was the extra file in the eda-code folder (called libs). What is that for?

### Summary assessment

-   **well structured**

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Everything was very well documented and easy to follow. However, I think the master Readme could use some clarification. First, it did not mention the `eda.qmd` file and when it should be run. I ran it between the processing and analysis files. Second, I think you could have bullet points (or numbered points) easily showing which code to run in which order. This would make it easier for someone who is not familiar with your project to run it.

### Summary assessment

-   **fully and well documented**

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was able to reproduce the entire project without any problem.

### Summary assessment

-   **fully reproducible without issues**

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

You did a fantastic job with thinking through the performances of your model. I was very impressed. I thought you could have included a few more references on this topic in your literature review.

### Summary assessment

-   **strong level of thoroughness**

## Further comments

Great job Arlyn and Tessa! I would make sure to proofread the manuscript document, as I found some spelling errors.

### 2024-04-26 products/manuscript/Manuscript.qmd

@@
 
-We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional.
+We will be looking for patterns in contraceptive methods and unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional.
 Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022].
@@
 The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.
+It is important to note the types of contraception included in each category defined by the Adding It Up project.
+Permanent methods include female and male sterilization, and long acting methods include IUDs and implants.
+These methods are combined for several variables and include a 'larcster' prefix or suffix to identify them throughout the dataset.
+Short-acting methods include hormonal pills, injectables, male and female condoms, emergency contraceptive pills, patches, rings, diaphragms, vaginal spermicide, and other supply methods, which is denoted with a 'sarc' prefix or suffix in the variable name.
+Traditional methods include periodic abstinence, withdrawal, abstinence, and breast-feeding, which is denoted with a 'trad' prefix or suffix in the variable name.
 
@@
 
-Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region?
+Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users between the five regions included in the Adding It Up project?
 

### 2024-04-26 KosewickPart5Review_files/KosewickPart5Review.qmd

---
title: Project Review Template 
author: Kevin Kosewick
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Global Distribution of Unintended Pregnancies and Contraceptive Usage

Name of project author(s): Taylor Glass and Arlyn Santiago

Name of project reviewer: Kevin Kosewick

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background/summary was very thorough and clearly described the aims/context of the project. I found no flaws with it. The second page of the manuscript is completely blank and could just be removed.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question is clearly identified but the hypothesis isn't stated. The questions are clearly connected to the data after reading the description of data and background sections. Clearly stating your hypothesis would give this section a 3/3.

There is a typo in section 2.3 ("people women") (minor issue, I just wanted to point it out).

### Summary assessment

* question/hypotheses somewhat explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data is well-described and a codebook is provided. Meta information included in the manuscript makes it clear what the data is.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data is very thoroughly cleaned in the processing-file.qmd. All steps are explained and justified. The eda.qmd file thoroughly explores the data, making almost 20 figures and a few tables. Each step in the EDA process is explained and interpreted. The supplementary file in the products folder shows plots of interest from the EDA and further connects their importance to the analysis and future analyses.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

In the first set of models (the linear models), the analysis was done well and explored many alternatives. Performance was evaluated and compared between the models to determine the most useful set of predictors for the outcome of interest. CV and bootstrapping are used to make sure the models work on unseen data. Overfitting was identified, so LASSO was used to fix this issue and then compared to a Random Forest model. This is all good practice and explores several ways of doing the same thing. Everything is thoroughly explained, justified, and interpreted. The analysis seems appropriate to me and was done with a lot of care.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results are presented very thoroughly. Each figure has one or multipe paragraphs interpreting what we're seeing. Important ploratory and model results are both included here. Each plot is visually appealing but I believe that the tables could be improved using some more customization from different packages (they're a bit basic). More importantly, these figures and tables are easy to read/interpret. Sections 4.2 and 4.3 have links to tables that doesn't seem to be working in the manuscript word document; I'm not sure if this is normal or not.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The findings are thoroughly and properly discussed. Each model's results are interpreted properly and linked back to the background/context of the project nicely. Strengths and limitations are clearly acknowledged in their own section of the manuscript. Further research questions are identified and the supplementary material provides helpful figures from the EDA to guide other researchers in directions that they could take the project for further studies. The conclusion is also thorough and wraps up the project nicely.

### Summary assessment

* strong, complete and clear discussion

## Further comments

This is a very well-done project. Everything is evaluated and justified thoroughly and several alternatives were explored in the model fitting process. Each of your documents is well-written and all code is heavily annotated and explained. I would read through your manuscript one more time just to catch any typos; there were several throughout. This is relatively unimportant though compared to the great job you both did analyzing this dataset.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is neatly organized and structured with reasonable names. No junk files are left. Some of the ReadMe folders (e.g. in the statistical analysis folder) still have template text left though; it would improve the project to remove the template text and write things more applicable to the design of your current repository.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is extremely well documented and dense with text. I am able to easily follow each step, decision, and line of code thanks to the annotations and the extra clarification in the supplement quarto file (included in the products folder).

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was able to completely reproduce everything except for the supplement quarto file. At line 76, I get an error saying that the .rds file being loaded doesn't exist. I'm not sure why though; I can clearly see it in the folder and your code uses the correct path as far as I can tell. Maybe it's something on my end? I would double check on your end just to be sure.

### Summary assessment

* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study was very thorough. Several models were considered in the analysis and the processing took several iterations before reaching a justifiable final result. The questions posed were thoroughly addressed. There wasn't a hypothesis posed to be addressed though.

### Summary assessment

* strong level of thorougness

## Further comments

A very solid project with only minor issues. As previously mentioned, I would double check line 76 in your supplementary file, but other than that all I could find were small typos in the manuscript. Great job doing such a thorough analysis!

### 2024-04-29 code/analysis-code/statistical-analysis.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

@@
 library(tidymodels) # for model building 
+library(flextable) # for creating tables for the manuscript
 library(yardstick) # for finding performance metrics 
@@
 # create a professional format of the table to add to the manuscript 
-kable_lmtable1 <- kable(lmtable1)
+lmtable1 <- as_flextable(lmtable1)
 
@@
 table_file1 = here("results", "tables", "resulttable1.rds")
-saveRDS(kable_lmtable1, file = table_file1)
+saveRDS(lmtable1, file = table_file1)
 ```
@@
 # create a professional format of the table to add to the manuscript 
-kable_lmtable2 <- kable(lmtable2)
+lmtable2 <- as_flextable(lmtable2)
 
@@
 table_file2 = here("results", "tables", "resulttable2.rds")
-saveRDS(kable_lmtable2, file = table_file2)
+saveRDS(lmtable2, file = table_file2)
 
@@
 # create a professional format of the table to add to the manuscript 
-kable_lmtable8 <- kable(lmtable8)
+lmtable8 <- as_flextable(lmtable8)
 
@@
 table_file8 = here("results", "tables", "resulttable8.rds")
-saveRDS(kable_lmtable8, file = table_file8)
+saveRDS(lmtable8, file = table_file8)
 ```

### 2024-04-29 code/eda-code/eda.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

@@
   geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
+  ggtitle("Percent of Unintended Pregnancies vs Standardized Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
 plot(plot_3)
@@
   geom_smooth(method='lm') + 
-  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
+  ggtitle("Percent of Unintended Pregnancies vs Standardized Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
 plot(plot_4)

### 2024-04-29 products/manuscript/Manuscript.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

@@
 # load a few R packages for rendering 
-library(here)
-library(readr)
-library(knitr)
-library(dplyr)
-library(naniar)
-library(ggplot2)
+library(here) # for locating tables and figures
+library(readr) # for loading tables and figures
+library(knitr) # for displaying tables and figures 
+library(dplyr) # for basic syntax
 ```
@@
 We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion.
-We also explore the differences in the outcome between regions using an ANOVA test and Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA.
 Result tables for each model and test are stored in the results > tables folder.
@@
 We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE).
-We chose these performance metrics because they are commonly used for linear regression models like the ones we have created.
-RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values.
-We include both metrics to increase the validity of our results.
-Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before.
+Once we decided on the final model based on RMSE and MAE, we use cross-validation to determine how well it will perform on data it has not seen before.
 This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets.
 To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions.
-This process is similar to cross-validation because it relies on resampling the training data.
-Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance.
+Next, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance.
 It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best.
-We use  a grid search to ensure that we try all combinations of the parameters in our grid.
-We complete the tuning process with a 5-fold cross-validation repeated 5 times.
-We added a random forest model to attempt to improve model performance.
+We used a grid search to ensure that we try all combinations of the parameters in our grid.
+Lastly, we added a random forest model to attempt to improve model performance.
 We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
-We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so.
-Each of these processes are crucial in determining which model performs best for our dataset.
+Each of these processes are crucial in determining which model performs best for our data.
 Result tables for each model and test are stored in the results > tables folder.
-Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.
+Figures created from cross-validation, bootstrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.
 
@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1. [insert explanation here]
-```{r}
-#| message: false
-#| label: tbl-1
-#| tbl-cap: "[A summary table was done to present data about the seven predictor variables chosen in a structured format.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
+A summary table was done to present data about the seven predictor variables chosen in a structured format.
 Summary tables enable us to compare between different categories or groups within the data set.
@@
 It appears that a higher percentage of women are currently married in Asia (66%) and Africa (60%) compared to Latin America/Caribbean (54%) and Oceania (60%).The average percentage of never married women across the 5 regions is 31%.
-Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.]"
+Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.
+```{r}
+#| message: false
+#| label: tbl-1
+#| tbl-cap: "Table 1: Summary of All Explored Variables"
 #| echo: FALSE
@@
 #| label: fig-1
-#| fig-cap: "[A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods.
-The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs.]"
+#| fig-cap: "A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods.
+The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs considering more than half of the countries experience less than 20% of unintended pregnancies when using SARCs."
 #| echo: FALSE
@@
 #| label: fig-2
-#| fig-cap: "[A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies.
+#| fig-cap: "A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies.
 In this case, a positive correlation is shown.
-It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)]"
+It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)"
 #| echo: FALSE
@@
 #| label: fig-3
-#| fig-cap: "[Scatter plots are used to see if there is a relationship between two sets of continuous data.
+#| fig-cap: "Scatter plots are used to see if there is a relationship between two sets of continuous data.
 In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies.
-The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa.]"
+The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa."
 #| echo: FALSE
@@
 #| label: fig-4
-#| fig-cap: "[Violin plots are used to visualize the distribution of data across different categories.
+#| fig-cap: "Violin plots are used to visualize the distribution of data across different categories.
 In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage.
-This means that there is more variation in unintended pregnancy rates in counties with high SARC usage.]"
+This means that there is more variation in unintended pregnancy rates in counties with high SARC usage."
 #| echo: FALSE
@@
 #| label: tbl-2
-#| fig-cap: "This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception.
+#| fig-cap: "Table 2: This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception.
 This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome.
@@
 #| label: tbl-3
-#| fig-cap: "In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia.
+#| fig-cap: " Table 3: In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia.
 Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome.

### 2024-04-29 products/manuscript/supplement/Supplementary-Material.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

@@
 
-We only briefly mentioned that we used cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript.
+We chose two performance metrics, root mean square error (RMSE) and mean absolute error (MAE) because they are commonly used for linear regression models like the ones we have created.
+RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values.
+We include both metrics to increase the validity of our results.
+We only briefly mentioned that the next step is to use cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript.
 Cross-validation was chosen as a method for determining how well the model performs on data it has not seen before because that is the main point of building models.
@@
 We chose a parameter of 0.1 first because that is standard for LASSO regression, but we also completed a parameter tuning process to validate that choice.
+We complete the tuning process with a 5-fold cross-validation repeated 5 times.
 As shown in @supplement-3, the RMSE performance metric increases exponentially when the penalty parameter increases past 1, which validates our previous selection of 0.1 for the penalty parameter because that will produce the lowest RMSE possible.
@@
 
-# Additional data cleaning explanation
-The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
-We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
-We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information.
-We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence.
-```{r}
-#| label: tbl-summarytable2
-#| tbl-cap: "Summary table to discover missingness"
-#| echo: FALSE
-resulttable2 <- readRDS("../../results/tables/summarytable.rds")
-kable(resulttable2)
-```
-
 {{< pagebreak >}}

### 2024-04-29 code/analysis-code/statistical-analysis.qmd
Part 6 Updates
Removed the slides folder. Responded to Kevin's feedback to fix supplement error and update READMe files throughout.

@@
 # create a professional format of the table to add to the manuscript 
-kable_lmtable11 <- kable(lmtable11)
+lmtable11 <- as_flextable(lmtable11)
 
@@
 table_file16 = here("results", "tables", "resulttable16.rds")
-saveRDS(kable_lmtable11, file = table_file16)
+saveRDS(lmtable11, file = table_file16)
 ```

### 2024-04-29 products/manuscript/Manuscript.qmd
Part 6 Updates
Removed the slides folder. Responded to Kevin's feedback to fix supplement error and update READMe files throughout.

@@
 
-The predictor variables will include the total number of people women using each type of birth control standardized by population size, region, percentage of women currently married, percentage of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due to abortion.
+The predictor variables will include the total number of women using each type of birth control standardized by population size, region, percentage of women currently married, percentage of women never married, total rate of maternal deaths, rate of maternal deaths due to unintended pregnancy and rate of maternal deaths due to abortion.
 
 A full list of variables can be found in the processing-file in the code > processing-code folder.
+
+We hypothesize that the percentage of unintended preggnancies among SARC users will vary between the five regions based on differing access to this type of contraception.
 

### 2024-04-29 products/manuscript/supplement/Supplementary-Material.qmd
Part 6 Updates
Removed the slides folder. Responded to Kevin's feedback to fix supplement error and update READMe files throughout.

@@
 # Additional data cleaning explanation
-The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable.
+The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable2.
 We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained.
@@
 #| echo: FALSE
-resulttable2 <- readRDS("../../results/tables/summarytable.rds")
+resulttable2 <- readRDS(here("results", "tables", "summarytable.rds"))
 kable(resulttable2)

### 2024-05-01 products/manuscript/Manuscript.qmd
Part 6 Updates
Final proofreading of the manuscript

@@
 Data from the Guttmacher Institute's Adding It Up project was used, which included information on reproductive-aged women (15-49) and various demographic factors.
-Single and multivariate linear regression models were employed to examine relationships between predictor variables and the outcome variable (unintended pregnancy rate).
+Single and multivariate linear regression models were employed to examine relationships between predictor variables and the outcome variable, percentage of unintended pregnancies among SARC users.
 
 The exploratory data analysis revealed a less skewed distribution of unintended pregnancies among SARC users compared to other contraceptive methods.
-This led researchers to focus on SARC use for the remainder of the study.
-The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia.
+This led the researchers to focus on SARC use for the remainder of the study.
+Exploratory graphs with the region predictor showed that the Latin America/Caribbean region displayed the strongest association with percentage of unintended pregnancies, followed by Asia.
 The Oceania and Europe region had weaker associations.
-Marriage status, both currently and never married, was inversely associated with unintended pregnancy rates.
-This means that as the percent of women who are currently married or never married increases, the unintended pregnancy rate decreases.
+Marriage status, both currently and never married, was inversely associated with percentage of unintended pregnancies.
+As the percent of women who are currently married or never married increases, the percentage of unintended pregnancies decreases.
 Maternal mortality rates showed a positive correlation with the outcome.
 
-A full model incorporating five predictors (total SARC users, region, marital status, and maternal mortality rates due to unintended pregnancy and abortion) was built.
+A full model incorporating five predictors, total SARC users, region, marital status, and maternal mortality rates due to unintended pregnancy and abortion, was chosen as the ideal model for this data.
 Despite performing the best among the evaluated models and being validated through cross-validation and the bootstrap method, the full model exhibited tendencies towards overfitting the data.
@@
 Marital status and maternal mortality rates displayed weaker negative associations with unintended pregnancy rates.
-While the direction of the association between abortion-related maternal mortality and unintended pregnancy rates flipped in the Lasso model, the coefficient (effect size) remained minimal.
+While the direction of the association between abortion-related maternal mortality and unintended pregnancy rates flipped in the Lasso model, the coefficient that represents the effect size remained minimal.
 
@@
 Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022].
-According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group .
+According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group.
 Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021].
@@
 This could be due to the current legal status of contraception within Latin American countries.
-Free emergency contraception access exists, but regulations differ.
-Chile, Colombia, and Ecuador legally recognize access, while Nicaragua and Bolivia rely on Ministerial protocols.
+Free emergency contraception access exists in this region, but differing regulations limit access.
+Chile, Colombia, and Ecuador legally recognize access to emergency contraception, while Nicaragua and Bolivia rely on Ministerial protocols.
 Argentina and Brazil lack legal recognition, but offer protocols and guides.
-Mexico requires provision to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012].
+Mexico requires provision of emergency contraception to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012].
 Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].
@@
 Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted.
-Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others.
+Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others.
 The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.
+
 It is important to note the types of contraception included in each category defined by the Adding It Up project.
-Permanent methods include female and male sterilization, and long acting methods include IUDs and implants.
+Permanent methods include female and male sterilization, and long acting methods include intrauterine devices (IUDs) and implants.
 These methods are combined for several variables and include a 'larcster' prefix or suffix to identify them throughout the dataset.
 Short-acting methods include hormonal pills, injectables, male and female condoms, emergency contraceptive pills, patches, rings, diaphragms, vaginal spermicide, and other supply methods, which is denoted with a 'sarc' prefix or suffix in the variable name.
-Traditional methods include periodic abstinence, withdrawal, abstinence, and breast-feeding, which is denoted with a 'trad' prefix or suffix in the variable name.
+Traditional methods include periodic abstinence, withdrawal, abstinence, and breast-feeding, which is denoted with a 'trad' prefix or suffix.
 
-For the purpose of this project, we chose several numeric variables from the data source: percentages of pregnancy outcomes, rates of safe and unsafe abortions, percentage of various contraceptive usages, percentages of care received for various pregnancy complications, and costs for many variables including abortions and STIs.
+For the purpose of this project, we chose several numeric variables from the data source: percentages of pregnancy outcomes, rates of safe and unsafe abortions, and percentage of various contraceptive usages.
 There are only three character variables that can be used to group the numeric variables: country name, region, and sub region.
@@
 
-A full list of variables can be found in the processing-file in the code > processing-code folder.
+A full list of variables can be found in the `processing-file.qmd` in the code > processing-code folder.
 
-We hypothesize that the percentage of unintended preggnancies among SARC users will vary between the five regions based on differing access to this type of contraception.
+We hypothesize that the percentage of unintended pregnancies among SARC users will vary between the five regions based on differing access to this type of contraception.
 
@@
 We chose 30 variables to clean to find the percentage of unintended pregnancies based on each type of contraceptive methods: long acting reversible and sterilization methods, short acting reversible methods, traditional methods, and no contraceptive usage.
-After the missing data was explored, we chose to use the percentage of unintended pregnancies variables related to each method of birth control because the total variable had unexplained missingness.
-The processed data is saved in the data > processed-data folder as processeddata.rds.
+After the missing data was explored, we chose to use the percentage of unintended pregnancies variables related to each method of birth control because the total variable had unexplained missing data.
+The processed data is saved in the data > processed-data folder as `processeddata.rds`.
 
@@
 These graphs revealed outliers, so we standardized by population size for each country.
-The next sets of plots are line graphs showing total usage of contraceptive methods to percentage of unintended pregnancies associated with that method.
-The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing.
-All of these figures are displayed and summarized in the Supplementary-Material.qmd file in the products > manuscript > supplement folder.
-This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods as our single outcome because this method is becoming increasingly popular, so we feel that results from this contraception method will be the most relevant.
+The next sets of plots are scatter plots showing total usage of contraceptive methods and percentage of unintended pregnancies associated with that method.
+The last two sets of plots contain scatter plots of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing.
+All of these figures are displayed and summarized in the `Supplementary-Material.qmd` file in the products > manuscript > supplement folder.
+This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods users as our single outcome because this method covers the most commonly used types of contraception, so we feel that results from this contraception method will be the most relevant.
 Additionally, the exploratory data analysis process showed that the distribution of this outcome variable is less skewed than the other options, and the stratification by region is clearer, which allows for better understanding.
@@
 
-Step three is the statistical analysis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder.
-We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion.
+Step three is the statistical analysis and model fitting portion found in the `statistical-analysis.qmd` file in the code > analysis-code folder.
+We explore the single outcome `pct_upreg_sarc` based on 7 predictors: total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion.
+We utilize simple linear regression and multiple linear regression with various combinations of the 7 predictors.
 Result tables for each model and test are stored in the results > tables folder.
 
-Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder.
+Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the `statistical-analysis.qmd` file in the code > analysis-code folder.
 We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE).
-Once we decided on the final model based on RMSE and MAE, we use cross-validation to determine how well it will perform on data it has not seen before.
-This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets.
-To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions.
+Once we decided on the final model based on RMSE and MAE, we use cross-validation to determine how well it will performs on data it has not seen before.
+This involves splitting the data into testing and training pieces, which allows for comparison of the predictions created using both datasets.
+To examine the uncertainty in the model, we used the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions.
 Next, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance.
 It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best.
-We used a grid search to ensure that we try all combinations of the parameters in our grid.
-Lastly, we added a random forest model to attempt to improve model performance.
+We used a grid search to ensure that all combinations of the parameters in our grid are considered.
+Lastly, we added a random forest model to attempt to improve model performance further.
 We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size.
-Each of these processes are crucial in determining which model performs best for our data.
+Each of these processes are crucial in determining which model performed best for our data.
 Result tables for each model and test are stored in the results > tables folder.
@@
 
-The Manuscript.qmd file for this document can be found in the products > manuscript folder.
+The `Manuscript.qmd` file for this document can be found in the products > manuscript folder.
 
@@
 This research aims to analyze the factors influencing unintended pregnancies and contraception use in low-and-middle-income countries.
-The Guttmacher Institute's Adding It Up project provides the foundation, offering data on sexual and reproductive health services in these regions.Focusing on women of reproductive age (15-49), the study usese information from nationally representative surveys and population projections to estimate unintended pregnancy rates.
+The Guttmacher Institute's Adding It Up project provides the foundation, offering data on sexual and reproductive health services in these regions.
+Focusing on women of reproductive age (15-49), the study uses information from nationally representative surveys and population projections to estimate percentage of unintended pregnancies.
 
-While data on the overall percentage of unintended pregnancies was unavailable, we strategically utilized data on rates related to specific birth control methods.
-This approach allows us to analyze the relationship between these rates and various social and health factors.
-Factors like the number of women utilizing different birth control methods, maternal death rates, and marital status will be meticulously examined.
-All 7 of these predictor variables were found within the study's codebook, which can can be found in the `Data` folder of this project.
+While data on the overall percentage of unintended pregnancies was unavailable, we strategically utilized data on percentages related to specific birth control methods.
+This approach allows us to analyze the relationship between these percentages and various social and health factors.
+For example, the number of women utilizing different birth control methods, maternal death rates, and marital status are all meticulously examined.
+All 7 of these predictor variables were found within the study's codebook, which can can be found in the `data` folder of this project.
 By discovering these relationships, the research aims to shed light on the complex landscape of contraception use and unintended pregnancy in low-and -middle-income countries.
@@
 ## Data import and cleaning
-As shown in the processing-file.qmd in the code > processing-code folder, we import the data from a .csv file that was downloaded directly from the Guttmacher AIU project website.
-We also included the codebook in this file.
-After taking a glimpse at the data, we identified the variables we need to clean and created a new dataset with those variables only.
-We explored missingness in the data, which is further explained in the supplementary-material.qmd file found in the products > manuscript > supplement folder.
+As shown in the `processing-file.qmd` in the code > processing-code folder, the data is imported from a .csv file that was downloaded directly from the Guttmacher AIU project website.
+The codebook is also included in this file.
+After taking a glimpse at the data, we identified the variables that need to be cleaned and created a new dataset with those variables only.
+We explored missingness in the data, which is further explained in the `supplementary-material.qmd` file found in the products > manuscript > supplement folder.
 Simple processing tasks include converting variable types, converting decimal value into true percentages, creating variables to find total usage of each birth control method across age groups, exploring outliers, and creating new standardized total usage values based on population size.
@@
 We began with a simple linear regression model with each of the 7 predictor variables to determine the association between each variable and the outcome.
-We explored the region variable further because our main goal is to understand the difference in outcome by region.
+We explored the `region` variable further because our main research question is to understand the difference in outcome by region.
 This led to discovering correlations between some of the predictors, which we examined through correlation coefficients.
@@
 We included a multiple linear regression model with all 5 of these predictors.
-We used the RMSE and MAE performance metrics to evaluate all 9 models that we created to determine which one fit best.
+We used the RMSE and MAE performance metrics to evaluate all 10 models that we created to determine which one fit best, including the null model with no predictors to serve as a comparison.
 Lastly, we used cross-validation to determine how well our multiple linear regression model performs on data it has not seen before.
@@
 
-After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
-A summary table was done to present data about the seven predictor variables chosen in a structured format.
-Summary tables enable us to compare between different categories or groups within the data set.
-This can reveal trends or relationships.
-From the summary table below, it would appear that the region with the highest rate of unintended pregnancies due to SARC usage is Latin America/Caribbean at 24%, followed by Europe at 23%.
+After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we used to predict it, which can be seen in @tbl-1.
+Based on the summary table below, it appears that the region with the highest rate of unintended pregnancies due to SARC usage is Latin America/Caribbean at 24%, followed by Europe at 23%.
 Africa and Asia have a mean of 13% and 14% respectively, while Oceania appears to have the lowest rate at 6%.
 It appears that a higher percentage of women are currently married in Asia (66%) and Africa (60%) compared to Latin America/Caribbean (54%) and Oceania (60%).The average percentage of never married women across the 5 regions is 31%.
-Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.
+Latin America/Caribbean region has the highest percentage of never married women at 35%, and the Asia region has the lowest percentage of never married women at 29%.
+Rates of maternal death are much higher in Africa with an average total rate of maternal deaths due to unintended pregnancies of 1,656 per 100,000 live births.
+The Asia region experiences the second highest rate in this category with 701 maternal deaths due to unintended pregnancies per 100,000 live births.
 ```{r}
@@
 #| label: tbl-1
-#| tbl-cap: "Table 1: Summary of All Explored Variables"
+#| tbl-cap: "Summary of All Explored Variables"
 #| echo: FALSE
@@
 #| label: fig-1
-#| fig-cap: "A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods.
+#| fig-cap: "This bar chart showcases the frequencies of the number of countries and the percentage of unintended pregnancies using SARC methods.
 The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs considering more than half of the countries experience less than 20% of unintended pregnancies when using SARCs."
@@
 Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception.
-This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place.
-This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users.
+This percentage should increase as the number of women using SARCs increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place.
+This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as SARC users.
 As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control.
-Each of these percentages are positively associated with number of people in its respective classification group that it represents.
-The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
+Each of these percentages are positively associated with the number of people in its respective classification group that it represents.
+The association for the contraception that we chose to focus on for this analysis, SARCs, is displayed in @fig-2.
 ```{r}
@@
 #| label: fig-2
-#| fig-cap: "A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies.
-In this case, a positive correlation is shown.
+#| fig-cap: "A scatter plot was contructed to examine if there is a correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies among SARC users.
+In this case, a clear positive correlation is shown.
 It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)"
@@
 
-We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 .
-The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
+The previous scatter plot is stratified by region to examine how the relationship differs across each of the regions which is shown in @fig-3 .
+The graph shows clear differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa experiences the sharpest increase in percentage of unintended pregnancies as the total population increases.
 ```{r}
@@
 #| label: fig-3
-#| fig-cap: "Scatter plots are used to see if there is a relationship between two sets of continuous data.
-In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies.
-The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa."
+#| fig-cap: "In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is examined by region.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies regardless of region, but there is some variation between regions.
+The regions with the highest percentage of unintended pregnacies are Latin America & the Caribbean, followed by Africa."
 #| echo: FALSE
@@
 
-We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4.
-It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
-The high usage countries appear to have a bottleneck in percentage of unintended pregnancies.
-The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies.
+Violin plots were created using the categorical SARC usage variable as shown in @fig-4.
+It appears that the distribution of SAR usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.
+The high-usage countries appear to have a bottleneck in percentage of unintended pregnancies.
+The low-usage countries appear to be skewed towards lower levels of unintended pregnancies among SARC users, but the medium-usage countries appear to have slightly higher percentages of unintended pregnancies.
 ```{r}
@@
 #| fig-cap: "Violin plots are used to visualize the distribution of data across different categories.
-In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage.
+In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).
+This violin plot shows that the distribution of percentage of unintended pregnancies is wider for countries with high SARC usage compared to counties with medium or low SARC usage.
 This means that there is more variation in unintended pregnancy rates in counties with high SARC usage."
@@
 We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region.
-The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram.
+The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right, which is to be expected based on the trend shown in the previous histogram.
 Chad and South Sudan also account for the highest rates of maternal mortality in this category.
@@
 #| label: tbl-2
-#| fig-cap: "Table 2: This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception.
+#| fig-cap: "Table 2: This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to SARCs.
 This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome.
-The coefficient of 87.65998 is the largest single coefficient of any predictor that we studied, which increases our confidence in including it as a predictor in our final model."
+The coefficient of 87.65998 is the largest single coefficient of any predictor studied in this analysis, which increases our confidence in including it as a predictor in our final model."
 #| echo: FALSE
@@
 
-The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region.
+The second model we created used the `region` predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable, followed closely by the Asia region.
 The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
@@
 #| label: tbl-3
-#| fig-cap: " Table 3: In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia.
-Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome.
+#| fig-cap: " Table 3: In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6 followed closely by Asia.
+Europe has a smaller efffect on the outcome based on the coefficient of 0.3, but Oceania has the smallest effect on the outcome.
 This model chose the African region as the reference group, which means that it has the weakest association with the outcome compared to the other 4 regions."
@@
 We created one model with all 5 predictors using multiple linear regression, shown in @tbl-4.
-This model shows the strongest predictor is the number of women using short acting reversible contraception.
-The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor.
+This model shows the strongest predictor is the number of women using SARCs.
+The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular `region` predictor.
 The Oceania region now has a negative association with the outcome.
@@
 #| label: tbl-4
-#| fig-cap: "The multiple linear regression model with all 5 predictors confirms that the `sarc_standard` predictor has the strongest effect on the outcome, but the effect size has decreased slightly.
+#| fig-cap: "Table 4: The multiple linear regression model with all 5 predictors confirms that the `sarc_standard` predictor has the strongest effect on the outcome, but the effect size has decreased slightly.
 The region variable is the second most impactful predictor, but there is a shift in which region has the strongest effect on the outcome.
@@
 To determine which model performed the best, we started by examining the RMSE and MAE performance metrics.
-We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion.
+We found that the model with the standardized total count of women using SARCs performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion.
 @tbl-5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors.
@@
 #| label: tbl-5
-#| fig-cap: "The performance metric comparisons clearly highlight the full model with 5 predictors as the best model with the lowest RMSE of 4.361363 and lowest MAE of 3.386193.
+#| fig-cap: "Table 5: The performance metric comparisons clearly highlight the full model with 5 predictors as the best model with the lowest RMSE of 4.361363 and lowest MAE of 3.386193.
 The model with percent of currently married women and the model with the rate of maternal mortality due to abortion only perform slightly better than the null model, which just predicts the average outcome.
@@
 We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces.
-The cross-validated results, shown in @tbl-6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
+The cross-validated results, shown in @tbl-6, lead to the same conclusion: the multiple linear regression model with 5 predictors performs better.
 The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319.
@@
 #| label: tbl-6
-#| fig-cap: "Cross-validated (CV) performance metric comparisons were only completed on two models: the single predictor model with the `sarc_standard` variable and the full model with all five predictors because these two models performed the best in the first analysis.
+#| fig-cap: "Table 6: Cross-validated (CV) performance metric comparisons were only completed on two models: the single predictor model with the `sarc_standard` variable and the full model with all five predictors because these two models performed the best in the first analysis.
 The RMSE metrics are closer using the CV technique with a differencec of 0.42 between the two models.
@@
 #| fig-cap: "Evaluating the final model performance on test data reveals that the model is strongly overfitting the data.
-The testing model predictions are shown in coral, and the variation in these observations is much higher compared to the training model predictions, shown in blue.
-Ideally, there would be complete overlap between the testing and training data predictions following the black dotted line that represent the ideal relationship between observed and predicted values.
+The testing model predictions are shown in coral, and the variation in these observations is much higher compared to the training model predictions, shown in teal.
+Ideally, there would be complete overlap between the testing and training data predictions following the black dotted line that represents the ideal relationship between observed and predicted values.
 The majority of testing data predictions do not overlap with the training data prediction in this case, and the testing data predictions are father away from the preferred relationship shown by the black dotted line."
@@
 
-Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression.
-This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting.
+Since we have evidence that our model overfits the data, we explored our full model with Lasso regression to determine if its performance could be improved.
+This technique could reveal predictors that should not be included in the final model, which could reduce the risk of overfitting.
 As seen in @tbl-7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model.
@@
 The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome.
-This process confirms that our final model should contain all 5 predictors.
+This process confirms that the final model should contain all 5 predictors.
 ```{r}
@@
 #| label: tbl-7
-#| fig-cap: "LASSO regression results confirm the findings of the multiple linear regression model, but each of the coefficients has been shrunk as a result of the penalty parameter of 0.1.
+#| fig-cap: "Table 7: LASSO regression results confirm the findings of the multiple linear regression model, but each of the coefficients has been shrunk as a result of the penalty parameter of 0.1.
 The `sarc_standard` predictor is still the most strongly related to the outcome, followed by the European region.
@@
 It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-8.
-As the regularization penalty increases past 1, we see that the RMSE value increases exponentially.
+As the regularization penalty increases past 1, the RMSE value increases exponentially.
 The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1.
@@
 
-After creating the random forest model, we found that the RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
+After creating the random forest model as another attempt to immprove model performance, we found that the RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better.
 Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy.
@@
 
-Again, it is crucial to tune hyperparameters associate with machine learning models.
+Again, it is crucial to tune hyperparameters associated with machine learning models.
 Based on the red line displayed in @fig-10 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors.
@@
 
-As seen in the statistical analysis file, we created 10 different model options with unique combinations of our 7 predictors.
-After considering effect sizes and performance metrics for each model, we have decided that the LASSO regression model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods.
-After considering each of the 10 models we created throughout the statistical analysis process, the random forest model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to short acting reversible methods.
+As seen in the `statistical-analysis.qmd` file, we created 10 different model options with unique combinations of our 7 predictors.
+After considering effect sizes and performance metrics for each model, we have decided that the LASSO regression model with 5 predictors is the best overall choice for predicting our outcome of percentage of unintended pregnancies due to SARCs.
 We used cross-validation with LASSO regression and random forest modeling to explore subset selection of our full model with all 5 predictors.
-The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the sarc_standard predictor.
+The LASSO regression kept all 5 predictors, and the coefficients for the Asia, Europe, and Latin America/the Caribbean predictors were the largest behind the `sarc_standard` predictor.
 The smallest coefficient was assigned to the percent of currently married women predictor.
+
 We suggest using the full model with all 5 predictors with a LASSO regression because it accurately portrays the importance of each predictor while providing a penalty to predictors that do not meaningfully contribute to the model.
-While this model has a slightly lower R-squared value compared to the random forest model, the evidence of overfitting is so strong that this difference is negligible.
+While this model has a slightly lower R-squared value than the random forest model, the evidence of overfitting is so strong that this difference is negligible.
 Cross-validation of the model with all 5 predictors showed that predictions from testing data deviated from the expected relationship with larger variance than the predictions from the training data.
@@
 Unfortunately, it is likely that any of these models are not optimal choices for predicting our outcome based on the poor performance of our model on testing data.
-While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.
+While we suggest using the LASSO regression with all 5 predictors to predict our outcome because it is slightly easier to interpret than the random forest model, future researchers who are willing to sacrifice interpretation for predictive power should use the random forest model with 4 predictors based on the slightly higher predictive performance shown by the R-squared performance metric.
 
@@
 ## Summary and Interpretation
-Our exploratory data analysis allowed us to examine the distribution of our outcome of interest, percentage of unintended pregnancies due to short acting reversible contraception, and the predictors used to describe the outcome.
+Our exploratory data analysis allowed for examination of the distribution of the outcome of interest, percentage of unintended pregnancies due to short acting reversible contraception and the predictors used to describe the outcome.
 Most low- and middle-income countries were from the African region, followed by Asia and Latin America/the Caribbean.
@@
 This association is likely explained by the fact that women in these regions have better access to short acting reversible contraception methods because we would have expected the African region to experience the highest rate of our outcome based on the number of countries included from that region.
-After standardizing the number of women who use short acting reversible contraception based on total population size, it is clear the women in the Latin America/the Caribbean region use this contraception method at the highest rates, followed my the Asian region.
+After standardizing the number of women who use SARCs based on total population size, it is clear the women in the Latin America/the Caribbean region use this contraception method at the highest rates, followed by the Asian region.
 Further analysis of the region predictor in a linear model with the outcome confirmed this finding by showing that Latin America/the Caribbean has the strongest association with the outcome, followed by Asia.
@@
 
-Our full statistical analysis of the final model with 5 predictors included the standardized number of women using short acting reversible methods, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors.
-Cross-validation using LASSO regression and random forest modeling techniques did not result in an increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data.
-Based on the model using LASSO regression, the number of women using short acting reversible contraception is the strongest predictor of the outcome, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place.
-The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome.
+Our full statistical analysis of the final model with 5 predictors included the standardized number of women using SARCs, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors.
+Cross-validation using LASSO regression and random forest modeling techniques did not result in a meaningful increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data.
+Based on the model using LASSO regression, the number of women using SARCs is the strongest predictor, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place.
+The most influential factor in this model is the `region` variable with European, Asian, and Latin American regions having the strongest effects on the outcome.
 This finding implies that these three regions have the best access to SARCs because they also have the highest standardized population size using this form of contraception in the first place.
@@
 First, this study relied on secondary data from the Guttmacher Institute's Adding It Up project.
-We were limited by the data that was available to us because we were unable to collect our own data.
+We were limited by the data that was available because we were unable to collect our own data.
 The study was cross-sectional, which means that it only looked at data from one point in time.
@@
 While we had solid evidence from performance metrics for choosing this model, it does a fairly poor job of representing the data based on the cross-validated testing and training data predictions.
-Our recommendation is to explore more variables from the Adding It Up project to determine if there are better predictors for this outcome.
+Our recommendation is to explore more variables from the Adding It Up project or additional sources to determine if there are better predictors for this outcome.
 Since we chose our predictors based on the literature, it is unlikely that there are other predictors that will more accurately predict the percentage of unintended pregnancies due to SARCs.
@@
 There are several possible explanations for this finding.
-First, SARC methods, such as pills and patches require more frequent user adherence than long-acting reversible contraceptives (LARCs) like implants.
+First, SARC methods, such as pills and patches, require more frequent user adherence than long-acting reversible contraceptives (LARCs) like implants.
 Perfect use of SARC methods is essential for their effectiveness, and even small lapses in use can increase the risk of unintended pregnancy.
@@
 
-The study also found that regional variations were significantly associated with unintended pregnancy rates.
-The Latin America/Caribbean region displayed the strongest association with unintended pregnancy rates, followed by Asia.
+The study also found that regional variations were significantly associated with percentage of unintended pregnancies.
+The Latin America/Caribbean region displayed the strongest association with the outcome, followed by Asia.
 The Oceania and Europe region had weaker associations.
-These findings suggest that there are significant contextual factors that influence unintended pregnancy rates across different regions.
+These findings suggest that there are significant contextual factors that influence the percentage of unintended pregnancies across different regions.
 These factors may include access to healthcare services, legal barriers to access, and cultural attitudes towards contraception.
-Additionally, regions with cultural or religious restrictions on contraception use may also have higher rates of unintended pregnancy.
+Additionally, regions with cultural or religious restrictions on contraception use may also have higher percetanges of unintended pregnancy.
 
-The study found that marital status was inversely associated with unintended pregnancy rates.
-We found that as the percent of women who are currently married increases, the unintended pregnancy rate decreases.
+The study found that marital status was inversely associated with the percentage of unintended pregnancies.
+As the percent of women who are currently married increases, the percentage of unintended pregnancies decreases.
 There are some reasonable explanations for this finding.
@@
 Future research directions could include conducting qualitative studies to better understand the reasons why women choose SARC methods and the challenges they face in using them effectively.
-Additionally, we suggest that a longitudinal study should be conducted to track changes in unintended pregnancy rates over time and to evaluate the effectiveness of different interventions aimed at reducing unintended pregnancy.
+Additionally, we suggest that a longitudinal study should be conducted to track changes in the percentage of unintended pregnancies over time and to evaluate the effectiveness of different interventions aimed at reducing unintended pregnancy.
 

### 2024-05-02 products/manuscript/supplement/Supplementary-Material.qmd
Fixed error message on line 346 in supplementary.qmd

@@
 #| echo: FALSE
-resulttable1 <- readRDS("../../results/tables/summarytablecat.rds")
+resulttable1 <- here("results","tables","summarytable.cat")
 kable(resulttable1)

### 2024-05-02 products/manuscript/Manuscript.qmd
Added to Background Info and proofread manuscript

@@
 This could be due to the current legal status of contraception within Latin American countries.
-Free emergency contraception access exists in this region, but differing regulations limit access.
-Chile, Colombia, and Ecuador legally recognize access to emergency contraception, while Nicaragua and Bolivia rely on Ministerial protocols.
-Argentina and Brazil lack legal recognition, but offer protocols and guides.
-Mexico requires provision of emergency contraception to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012].
-Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].
+All Latin American have ratified international agreements that recognize a woman's right to privacy, health, and control over her reproductive choices.
+However, the legal status of contraception varies greatly by country.
+Chile, Colombia, and Ecuador legally recognize access to emergency contraception, while Nicaragua and Bolivia rely on legal protocols.
+In 2008, the Chilean Constitutional Court argued against the legal principle of “giving preference to an interpretation that favors the right of ‘a person’ to life above any other interpretation that may annul such a right.”[@heviaLegalStatusEmergency2012] This decision permitted the use of emergency contraception.
+Argentina and Brazil lack legal recognition, but offer protocols and guides as part of the “Obligatory Medical Program”.
+Mexico requires provision of emergency contraception to victims of sexual and domestic violence.
+Contrastingly, in 2009, Honduras issued a complete ban on the distribution of emergency contraception.
+[@heviaLegalStatusEmergency2012].
+It is important to note that judges are not responsible for examining scientific evidence before making a legal decision.
+It has been observed that stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging.
+[@kavanaughContraceptiveMethodUse2018].
 

### 2024-05-02 products/manuscript/supplement/Supplementary-Material.qmd
Part 6 Final Updates
Checked ReadMe files and made one small update in the supplement.

@@
 #| echo: FALSE
-resulttable1 <- here("results","tables","summarytable.cat")
+resulttable1 <- readRDS(here("results","tables","summarytablecat.rds"))
 kable(resulttable1)
